---
title: "2025-26 Q4 Data Report"
author: "Molly Wilson"
date: "2026-07-01"
output: 
  pdf_document:
    toc: true
  # html_document:
  #   toc: true
  #   toc_float: true
  #   code_folding: hide
  #   theme: cerulean
  #   highlight: haddock
---
# Setup

```{r, message = F, warning = F, echo = F}
library(tidyverse)
library(here) 
library(readxl) 
library(janitor)
library(scales)
library(lubridate)
library(knitr)
library(kableExtra) # do I need this with DT?
library(DT)

if (knitr::is_html_output()) {
  knitr::opts_chunk$set(
    message = FALSE,
    warning = FALSE
  )
} else {
  knitr::opts_chunk$set(
    message = FALSE,
    warning = FALSE,
    echo = FALSE
  )
}

print_table <- function(x, options = list(pageLength = 10)) {
  
  if (knitr::is_html_output()) {
    
    DT::datatable(
      x,
      options = options,
      rownames = FALSE
    )
    
  } else {
    
    knitr::kable(
      x,
      format = "latex",
      booktabs = TRUE,
      longtable = TRUE,
      row.names = FALSE
    )
    
  }
}
```

```{r}
# quarterly settings
q_label <- "Q4 2025-26"
q_start <- "2026-06-01"
q_end <- "2026-08-15" # extended for nursery work

# biannual settings
b_start <- "2025-02-01"
b_end <- "2026-08-15" # extended for nursery work

# fig directory
fig_dir <- here("restoration_monitoring", "figs", "25-26_Q4")
```


# Nurseries

## Nursery stats

```{r}
nursery_mon <- read_excel(here("restoration_monitoring", "data_raw", "Genotype tracking.xlsx"), sheet = "nursery monitoring") %>% 
  clean_names() %>%
  mutate(date = ymd(as.character(date)),
         nursery = if_else(nursery == "TPB", "Ten Pound Bay", nursery)) %>%
  filter(date >= ymd(q_start) & date <= ymd(q_end)) %>%
  mutate(species = substr(genotype, 1, 4),
         across(starts_with("n_"), ~ replace_na(.x, 0)),
         n_tot = n_healthy + n_bleached + n_pale + n_diseased + n_other)

nurs_mon_last <- nursery_mon %>%
  group_by(nursery) %>%
  slice_max(date, n = 1) %>% # keep row(s) with max date per nursery, assumes all monitoring per nursery done in one day
  ungroup() %>%
  select(date, nursery, location, species, genotype, starts_with("n_"))
```

### By nursery

```{r}
frags_nursery <- nurs_mon_last %>%
  group_by(nursery) %>%
  summarize(n_tot = sum(n_tot),
            n_healthy = sum(n_healthy),
            n_bleached = sum(n_bleached),
            n_pale = sum(n_pale),
            n_diseased = sum(n_diseased)) %>%
  adorn_totals("row") %>%
  mutate(p_healthy = percent(n_healthy / n_tot, accuracy = 0.1),
         p_stressed = percent((n_bleached + n_pale) / n_tot, accuracy = 0.1),
         p_diseased = percent(n_diseased / n_tot, accuracy = 0.1)) %>%
  left_join(nurs_mon_last %>%
              select(nursery, species) %>%
              distinct() %>%
              group_by(nursery) %>%
              summarize(n_spp = n())) %>%
  left_join(nurs_mon_last %>%
              select(nursery, genotype) %>%
              distinct() %>%
              filter(!str_detect(genotype, "XX")) %>% # don't want unknown genotypes to double count
              group_by(nursery) %>%
              summarize(n_genotypes = n())) %>%
  select(nursery, n_frags = n_tot, p_healthy, p_stressed, p_diseased, n_spp, n_genotypes)

kable(frags_nursery)
```

### By species

```{r}
frags_spp <- nurs_mon_last %>%
  select(species, genotype) %>%
  distinct() %>%
  filter(!str_detect(genotype, "XX")) %>% # don't want unknown genotypes to double count
  group_by(species) %>%
  summarize(n_genotypes = n()) %>%
  left_join(nurs_mon_last %>%
              group_by(species) %>%
              summarize(n_tot = sum(n_tot),
                        n_healthy = sum(n_healthy),
                        n_bleached = sum(n_bleached),
                        n_pale = sum(n_pale),
                        n_diseased = sum(n_diseased))) %>%
  adorn_totals("row") %>%
  mutate(p_healthy = percent(n_healthy / n_tot, accuracy = 0.1),
         p_stressed = percent((n_bleached + n_pale) / n_tot, accuracy = 0.1),
         p_diseased = percent(n_diseased / n_tot, accuracy = 0.1)) %>%
  select(species, n_genotypes, n_frags = n_tot, p_healthy, p_stressed, p_diseased)

print_table(frags_spp)
```
### By genotype

*See genotype tracking for consolidated summary*

```{r}
frags_genotype <- nurs_mon_last %>%
  group_by(genotype) %>%
  summarize(n_tot = sum(n_tot)) %>%
  select(genotype, n_frags = n_tot)
```

### By structure

- Could eventually add structure type, but not sure how to automate this based on current nursery map formatting
- Completely empty structures are currently omitted in map code - adjust this to include here?
```{r}
frags_structure <- nurs_mon_last %>%
  group_by(nursery, location) %>%
  summarize(n_tot = sum(n_tot)) %>%
  select(nursery, location, n_frags = n_tot)

print_table(frags_structure)
```


## Nursery locations

### Map locations

```{r}
letter_names <- function(nms) LETTERS[seq_along(nms)]

map_york <- read_excel(here("restoration_monitoring", "data_raw", "Nursery map.xlsx"), sheet = "York", 
                   range = "A1:G9", col_names = FALSE, .name_repair = letter_names) %>%
  mutate(row = row_number()) %>%
  pivot_longer(cols = -row, 
               names_to = "col", 
               values_to = "genotype", 
               values_drop_na = TRUE) %>%
  unite(location, col, row, sep = "", remove = TRUE, na.rm = TRUE) %>%
  mutate(nursery = "York")

map_tpb <- read_excel(here("restoration_monitoring", "data_raw", "Nursery map.xlsx"), sheet = "TPB", 
                   range = "A1:C13", col_names = FALSE, .name_repair = letter_names) %>%
  pivot_longer(cols = everything(),
               names_to = NULL,
               values_to = "genotype",
               values_drop_na = TRUE) %>%
  mutate(
    location = str_extract(genotype, "^[A-Z]\\d+"),
    genotype = str_remove(genotype, "^[A-Z]\\d+\\s*")
  ) %>%
  # separate_rows(genotype, sep = "\\s*/\\s*")
  filter(genotype != "") %>% # remove empty structures
  bind_rows(read_excel(here("restoration_monitoring", "data_raw", "Nursery map.xlsx"), sheet = "TPB", 
                       range = "A16:H23", col_names = FALSE) %>% 
              pivot_longer(
                cols = everything(),
                names_to = NULL,
                values_to = "genotype",
                values_drop_na = TRUE) %>%
              distinct() %>%
              mutate(location = "ST")
  ) %>%
  mutate(nursery = "Ten Pound Bay")

map_cades_sculptures <- read_excel(here("restoration_monitoring", "data_raw", "Nursery map.xlsx"), sheet = "Cades Sculptures",
                   range = "A1:B8", col_names = FALSE, .name_repair = letter_names) %>%
  pivot_longer(cols = everything(),
               names_to = NULL,
               values_to = "genotype",
               values_drop_na = TRUE) %>%
  mutate(
    location = str_extract(genotype, "^[A-Z]\\d+"),
    genotype = str_remove(genotype, "^[A-Z]\\d+\\s*"),
    nursery = "Cades Sculptures")

map_jumby <- read_excel(here("restoration_monitoring", "data_raw", "Nursery map.xlsx"), sheet = "Jumby",
                   range = "A1:C8", col_names = FALSE, .name_repair = letter_names) %>%
  pivot_longer(cols = everything(),
               names_to = NULL,
               values_to = "genotype",
               values_drop_na = TRUE) %>%
  mutate(
    location = str_extract(genotype, "^[A-Z]\\d+"),
    genotype = str_remove(genotype, "^[A-Z]\\d+\\s*"),
    nursery = "Jumby")

nursery_maps <- bind_rows(map_york, map_tpb, map_cades_sculptures, map_jumby) %>%
  # mutate(genotype = strsplit(as.character(genotype), " / ")) %>% 
  mutate(
    genotype = strsplit(as.character(genotype), "/"),
    genotype = lapply(genotype, function(x) {
      str_trim(str_replace_all(x, "\\\\n|\\n", ""))
    })
  ) %>%
  unnest(genotype) %>%
  filter(!str_detect(genotype, "GB")) %>% # removing genebanks for now
  mutate(geno_check = nchar(genotype) == 6)
```

### Monitoring locations

```{r}
nursery_locations <- nurs_mon_last %>%
              select(genotype, nursery, location) %>%
              group_by(genotype, nursery) %>%
              summarise(location = paste(location, collapse=", ")) %>%
              pivot_wider(names_from = nursery, values_from = location) %>%
              rename(york_locations = "York",
                     tpb_locations = "Ten Pound Bay",
                     cades_locations = "Cades Sculptures",
                     jumby_locations = "Jumby")
```



### Consistency check

*Inconsistencies in location records between nursery maps and nursery monitoring*
```{r}
diff_nmap <- anti_join(nursery_maps %>% select(nursery, location, genotype), 
                       nurs_mon_last %>% select(nursery, location, genotype), 
                       by = c("nursery", "location", "genotype"))

diff_nmon <- anti_join(nurs_mon_last %>% select(nursery, location, genotype), 
                       nursery_maps %>% select(nursery, location, genotype), 
                       by = c("nursery", "location", "genotype")) %>%
  anti_join( # removing GB locations because those are not in maps by genotype
   bind_rows(map_york, map_tpb, map_cades_sculptures, map_jumby) %>%
     filter(str_detect(genotype, "GB")) %>% select(nursery, location),
    by = c("nursery", "location")
  )

diff_locations <- bind_rows(
  mutate(diff_nmap, source = "nursery maps"),
  mutate(diff_nmon, source = "nursery monitoring")) 

kable(diff_locations)
```

# Collections

- Merge DCYL02 and DCYL03?

```{r}
collection <- read_excel(here("restoration_monitoring", "data_raw", "Genotype tracking.xlsx"), sheet = "collection") %>%
  clean_names() %>%
  mutate(across(everything(), 
                ~ ifelse(grepl("^(na|n/a)$", ., ignore.case = TRUE), NA, .)),
         collection_date = as.Date(as.numeric(date), origin = "1899-12-30"),
         site = str_to_title(site),
         species = substr(genotype, 1, 4)) %>%
  rename(collection_site = site,
         collection_lat = lat,
         collection_lon = lon) %>%
  select(genotype, species, collection_date, collection_site, collection_lat, collection_lon) %>%
  group_by(genotype) %>%
  slice_min(collection_date, n = 1)
```

## By species
```{r}
collection_spp <- collection %>%
  group_by(species) %>%
  summarize(colonies = n()) %>%
  adorn_totals("row")

print_table(collection_spp)
```

## By site
```{r}
collection_sites <- collection %>%
  group_by(collection_site) %>%
  summarize(colonies = n()) %>%
  adorn_totals("row")

print_table(collection_sites)
```

# Outplants

## Outplant numbers

```{r}
outplants_general <- read_excel(here("restoration_monitoring", "data_raw", "Genotype tracking.xlsx"), sheet = "outplants") %>% 
  clean_names() %>%
  filter(!is.na(date)) %>%
  rename(id = genotype) %>%
  mutate(genotype = substr(id, 1, 6),
         species = substr(id, 1, 4)) %>%
  select(date, site, species, genotype, n_frags)

outplants_pmplots <- read_excel(here("restoration_monitoring", "data_raw", "Genotype tracking.xlsx"), sheet = "photomosaic plots") %>% 
  clean_names() %>%
  filter(!is.na(date)) %>%
  rename(date = date_planted,
         site = plot) %>%
  mutate(species = substr(genotype, 1, 4)) %>%
  select(date, site, species, genotype, n_frags)

outplants <- bind_rows(outplants_general, outplants_pmplots)
```

### By genotype
```{r}
outplant_genotypes <- outplants %>%
  filter(year(date) >= 2024) %>% # not including outplants prior to bleaching
  group_by(genotype, site) %>%
  summarize(n_frags = sum(n_frags)) %>%
  pivot_wider(names_from = site, values_from = n_frags) %>%
  mutate(across(everything(), ~ replace_na(., 0))) %>%
  mutate(n_outplanted = rowSums(across(where(is.numeric)), na.rm = TRUE)) %>%
  adorn_totals("row")

print_table(outplant_genotypes)
```

# Genotype tracking

Notes:

- Need to integrate genotype-level genebank performance over time

```{r}
# integrate genotype location(s) (nursery map), # frags per genotype (nursery monitoring) and genotype performance into one output

genotype_tracking <- collection %>%
  full_join(nursery_locations, by = "genotype") %>%
  full_join(frags_genotype, by = "genotype") %>%
  full_join(outplant_genotypes, by = "genotype") %>%
  mutate(updated = Sys.Date(),
         across(contains("_locations"), ~replace_na(.,"")),
         across(starts_with("n_"), ~replace_na(.,0)),
         active = if_else(if_all(contains("_locations"), ~ . == "" | is.na(.)),
                           "N", "Y")) %>%  # genotypes with frags in nurseries
  arrange(genotype) %>%
  select(updated, genotype, active, collection_date, collection_site, collection_lat, collection_lon, york_locations, tpb_locations, cades_locations, jumby_locations, n_frags, n_outplanted)

print_table(genotype_tracking)

write.csv(genotype_tracking, here("restoration_monitoring", "data_outputs", "genotype_tracking.csv"), row.names=FALSE)
```

# Temp loggers

```{r}
cades <- read_excel(here("restoration_monitoring", "data_raw", "hobo", "Hobo logger data.xlsx"), sheet = "Cades") %>%
  clean_names() %>%
  slice(-1) %>% # remove first row 
  mutate(logger = "Cades")

tpb <- read_excel(here("restoration_monitoring", "data_raw", "hobo", "Hobo logger data.xlsx"), sheet = "Ten Pound Bay") %>%
  clean_names() %>%
  slice(-1) %>%
  mutate(logger = "Ten Pound Bay")

jumby <- read_excel(here("restoration_monitoring", "data_raw", "hobo", "Hobo logger data.xlsx"), sheet = "Jumby") %>%
  clean_names() %>%
  slice(-1) %>%
  mutate(logger = "Jumby")

# create a filter to remove times during surface intervals
surf_int <- read_excel(here("restoration_monitoring", "data_raw", "hobo", "Hobo logger data.xlsx"), sheet = "Surface intervals") %>%
  clean_names() %>%
  select(-notes)

temp_data <- bind_rows(cades, tpb, jumby) %>%
  select(-notes) %>%
  filter(between(temp_c, 25, 35)) %>% # removing rogue outliers
  # filter out any readings taken during surface intervals
  anti_join(surf_int,
            by = join_by(
              logger == logger,
              between(date_time, surface_start, surface_end)
            )) %>%
  # drop any readings more than 2C above previous hour's reading
  arrange(logger, date_time) %>%
  group_by(logger) %>%
  mutate(temp_diff = temp_c - lag(temp_c)) %>% # create helper column
  filter(is.na(temp_diff) | temp_diff <= 2) %>%  # keep if first row or <= 2 °C jump
  ungroup() %>%
  select(-temp_diff) # drop helper column

# define references
lab_temp <- expression("Temperature ("*~degree*C*")")
ref_temp = data.frame(xmin = -Inf, xmax = Inf, ymin = 23, ymax = 29.6)
lab_ref_temp <- "Target range (Coral Reef Alliance)"
ref_temp_bleaching = 30.63
lab_ref_temp_bleaching <- "Coral bleaching threshold (NOAA)"
ylims_temp <- c(22, (max(temp_data$temp_c, na.rm = T) + 1))

# plots
ggplot() +
  geom_rect(data = ref_temp, 
              aes(xmin = as.POSIXct(-Inf), xmax = as.POSIXct(Inf), 
                  ymin = ymin, ymax = ymax, 
                  fill = lab_ref_temp),
              alpha = 0.3) +
  scale_fill_manual(values = "gray") +
  geom_line(data = temp_data,
            aes(x = date_time, y = temp_c, group = logger, color = logger),
            alpha = 0.9) +
  scale_color_manual(values = c("cadetblue", "coral", "gold")) +
  geom_hline(yintercept = ref_temp_bleaching, linetype = 'dashed') +
  scale_y_continuous(sec.axis = sec_axis(~.*9/5+32, name = "Temperature (°F)")) +
  labs(x = "", y = "Temperature (°C)", color = "Nursery location", fill = "") +
  theme_bw() +
  theme(legend.position = "top")
ggsave(here(fig_dir, "hobo_summary.png"), width = 6, height = 4)

ggplot(temp_data,
       aes(x = date_time, y = temp_c)) +
  geom_line(alpha = 0.9, size = 0.3, color = "cadetblue") +
  geom_hline(yintercept = ref_temp_bleaching, linetype = 'dashed', color = "coral") +
  scale_y_continuous(sec.axis = sec_axis(~.*9/5+32, name = "Temperature (°F)")) +
  facet_grid(. ~ logger) +
  labs(x = "", y = "Temperature (°C)", color = "Nursery") +
  theme_bw()
ggsave(here(fig_dir, "hobo_nursery.png"), width = 6, height = 4)
```