---
title: "Ecological monitoring at restoration sites"
author: "Molly Wilson"
date: "2023-06-04"
output: 
  html_document:
    code_folding: hide
---
# {.tabset}

```{r, message = F, warning = F, echo = F}
library(tidyverse)
library(here) 
library(readxl) 
library(janitor)
library(snakecase) # for adjusting capitalization of text within data (e.g., species names)
library(knitr) # for including tables
library(lubridate)
library(forcats) # for fct_relevel

knitr::opts_chunk$set(message = FALSE, warning = FALSE)
```

## Benthic

```{r}
# import and clean data
## note benthic data currently includes only transects 1-2 of original TPB assessment data (because 3-4 were collected on east side)

rf_benthic <- read_excel(here("restoration_monitoring", "data_raw", "ecological monitoring.xlsx"), sheet = "benthic") %>% 
  clean_names() %>%
  rename(transect_og = transect) %>%
  unite(transect, date, transect_og, sep = "_", remove = FALSE) %>% # making sure transects don't get combined if they were surveyed on different dates (for combining TPB baseline with MMA data)
  select(-time) %>%
  filter(!is.na(site)) %>% # remove any incomplete rows at the end of the data
  # separate(species_code, c("species_code", "indicator")) %>% # separating variables with an underscore into two columns
  mutate(date = ymd(date),
         # date_mon = format_ISO8601(ymd(date), precision = "ym"),
         monitoring = case_when(year(date) <= 2022 ~ "baseline",
                                date %within% interval(ymd("2023-01-01"), ymd("2023-06-01")) ~ "6month"),
         species = to_any_case(species, case = "sentence"), # this makes sure species names are correctly capitalized
         genus = word(species, 1), # eventually should only put species (not turf, etc.) in species column
         site = to_any_case(site, case = "title"),
         # consolidating categories here for graphs:
         cat_c = case_when(category_code %in% c("LC", "SLC") ~ "Hard corals",
                                  category_code %in% c("MA") ~ "Macroalgae",
                                  category_code %in% c("TA") ~ "Turf algae",
                                  category_code %in% c("CCA", "CCA_ND") ~ "CCA",
                                  grepl("CYAN", category_code) ~ "Cyanobacteria", # currently sites with CYAN and others are considered CYAN
                                  category_code %in% c("OINV", "SPON", "AINV", "PEY") ~ "Other competitors",
                                  category_code %in% c("DC", "SAND", "HOLE", "SG", "PAVE") ~ "Other substratum"
                                    ),
         # adding algal type for graphs about palatability:
         algal_type = case_when(type_code %in% c("BFMA", "GFMA", "RFMA") ~ "Fleshy macroalgae",
                                type_code %in% c("GCMA", "RCMA") ~ "Calcareous macroalgae",
                                type_code %in% c("TA", "TAS", "STA") ~ "Turf algae"),
         # certain substrates are not suitable for coral or algal growth, so should not detract from percent cover
         av_sub_yn = if_else(category_code %in% c("SAND", "HOLE", "SG", "PAVE"), "no", "yes")
         )

n_transects <- rf_benthic %>%
  select(monitoring, site, transect) %>%
  distinct() %>%
  group_by(monitoring, site) %>%
  summarize(n = n())
```

### Percent cover by site and category

```{r}
rf_pc_cat_m <- rf_benthic %>%
  filter(av_sub_yn == "yes") %>% # percent cover is relative to available substrate
  expand(nesting(monitoring, site, transect, meter), cat_c) %>%
  # this is where we add in our actual data to this expanded template:
  left_join(rf_benthic %>% 
              filter(av_sub_yn == "yes") %>% # only looking at what is considered available substrate (no sand, etc.)
              group_by(monitoring, site, transect, meter) %>%
              mutate(n_pts = n()) %>% # showing total number of points per meter that are considered available substrate
              ungroup() %>%
              group_by(monitoring, site, transect, meter, n_pts, cat_c) %>%
              summarize(pc_m = 100*n()/n_pts) %>% # n() counts the number of entries within a given group
              ungroup() %>%
              distinct() %>%
              select(-n_pts), 
            by = c("monitoring", "site", "transect", "meter", "cat_c")) %>%
  mutate(pc_m = if_else(is.na(pc_m), 0, pc_m))

# average these meter-level results within transect, then within sites
rf_pc_cat_site <- rf_pc_cat_m %>%
  group_by(monitoring, site, transect, cat_c) %>%
  summarize(pc_t = mean(pc_m)) %>%
  ungroup() %>%
  group_by(monitoring, site, cat_c) %>%
  summarize(n_test = n(), # should add up to n transects
            pc_mean = mean(pc_t),
            pc_se = sd(pc_t)/sqrt(n())
            ) %>%
  filter(cat_c != "Other substratum") %>% 
  mutate(cat_c = factor(cat_c, levels = c("Hard corals", "CCA", "Macroalgae", "Turf algae", "Cyanobacteria", "Other competitors")))

cat_palette <- c("coral2", "pink", "darkolivegreen", "darkkhaki", "slategray3", "gray50")
```

```{r, eval = F}
ggplot(rf_pc_cat_site, 
       aes(x = "cat_c", y = pc_mean, fill = cat_c)) + 
  geom_bar(width = 1, stat = "identity", color = "black") +
  coord_polar("y", start=0) +
  scale_fill_manual(values = cat_palette) +
  facet_wrap(vars(relevel(monitoring, "baseline"), site), nrow = 2) +
  theme_void() +
  theme(legend.position = "bottom",
        legend.title = element_blank(),
        panel.spacing = unit(1, "lines"))
# ggsave(here("restoration_monitoring", "figs", "selection_benthic_pc.png"))
```

```{r}
ggplot(rf_pc_cat_site, 
       aes(x = relevel(factor(monitoring), "baseline"), y = pc_mean, fill = cat_c)) +
  geom_col(color = "black") +
  scale_fill_manual(values = cat_palette) +
  geom_errorbar(aes(ymin = pc_mean - pc_se, ymax = pc_mean + pc_se), width = .2,
                 position = position_dodge(.9)) +
  facet_grid(site ~ cat_c) +
  labs(y = "Percent cover", x = "") +
  scale_x_discrete(labels = c('Baseline','6 mo.')) +
  theme_bw() +
  theme(legend.position = "none")

ggplot(rf_pc_cat_site, 
       aes(x = relevel(factor(monitoring), "baseline"), y = pc_mean, color = cat_c, shape = cat_c, group = cat_c)) +
  geom_point() +
  geom_line() +
  scale_color_manual(values = cat_palette) +
  facet_grid(. ~ site) +
  labs(y = "Percent cover", x = "", color = "Category", shape = "Category") +
  scale_x_discrete(labels = c('Baseline','6 mo.')) +
  theme_bw()
```
```{r}
# calculating percent cover by species at each meter
rf_pc_spp_m <- rf_benthic %>%
  filter(av_sub_yn == "yes") %>%
  expand(nesting(monitoring, site, transect, meter), nesting(species, cat_c)) %>%
  left_join(rf_benthic %>% 
              filter(av_sub_yn == "yes") %>% 
              group_by(monitoring, site, transect, meter) %>%
              mutate(n_pts = n()) %>% 
              ungroup() %>%
              group_by(monitoring, site, transect, meter, n_pts, species, cat_c) %>%
              summarize(pc_m = 100*n()/n_pts) %>%
              ungroup() %>%
              distinct() %>%
              select(-n_pts), 
            by = c("monitoring", "site", "transect", "meter", "species", "cat_c")) %>%
  mutate(pc_m = if_else(is.na(pc_m), 0, pc_m))

# calculating mean percent cover by transect -> site
rf_pc_spp_site <- rf_pc_spp_m %>%
  group_by(monitoring, site, transect, species, cat_c) %>%
  summarize(pc_t = mean(pc_m)) %>%
  ungroup() %>%
  group_by(monitoring, site, species, cat_c) %>%
  summarize(pc_mean = mean(pc_t),
            pc_se = sd(pc_t)/sqrt(n())
            ) %>%
  ungroup() %>%
  distinct()

# calculating mean percent cover by species across all sites
rf_pc_spp <- rf_pc_spp_site %>%
  group_by(monitoring, species, cat_c) %>%
  summarize(pc = mean(pc_mean),
            pc_se = sd(pc_mean)/sqrt(n())) %>%
  rename(pc_mean = pc) 

# testing that total percent cover per transect and site add up to 100
test1 <- rf_pc_spp_m %>%
  group_by(monitoring, site, transect, meter) %>%
  summarize(total = sum(pc_m))
test2 <- rf_pc_spp_site %>%
  group_by(monitoring, site) %>%
  summarize(total = sum(pc_mean))
test3 <- sum(rf_pc_spp$pc_mean)
```

### Percent cover of hard corals by genus and site

```{r}
# calculating percent cover by genus at each meter
rf_pc_gen_m <- rf_benthic %>%
  filter(av_sub_yn == "yes") %>%
  expand(nesting(monitoring, site, transect, meter), nesting(genus, cat_c)) %>%
  left_join(rf_benthic %>% 
              filter(av_sub_yn == "yes") %>% 
              group_by(monitoring, site, transect, meter) %>%
              mutate(n_pts = n()) %>% 
              ungroup() %>%
              group_by(monitoring, site, transect, meter, n_pts, genus, cat_c) %>%
              summarize(pc_m = 100*n()/n_pts) %>%
              ungroup() %>%
              distinct() %>%
              select(-n_pts), 
            by = c("monitoring", "site", "transect", "meter", "genus", "cat_c")) %>%
  mutate(pc_m = if_else(is.na(pc_m), 0, pc_m))

# calculating mean percent cover by transect -> site
rf_pc_gen_site <- rf_pc_gen_m %>%
  group_by(monitoring, site, transect, genus, cat_c) %>%
  summarize(pc_t = mean(pc_m)) %>%
  ungroup() %>%
  group_by(monitoring, site, genus, cat_c) %>%
  summarize(pc_mean = mean(pc_t),
            pc_se = sd(pc_t)/sqrt(n())
            ) %>%
  ungroup() %>%
  distinct()

# calculating mean percent cover by species across all sites
rf_pc_gen <- rf_pc_gen_site %>%
  group_by(monitoring, genus, cat_c) %>%
  summarize(pc = mean(pc_mean),
            pc_se = sd(pc_mean)/sqrt(n())) %>%
  rename(pc_mean = pc) 

# testing that total percent cover per transect and site add up to 100
test1 <- rf_pc_gen_m %>%
  group_by(monitoring, site, transect, meter) %>%
  summarize(total = sum(pc_m))
test2 <- rf_pc_gen_site %>%
  group_by(monitoring, site) %>%
  summarize(total = sum(pc_mean))
test3 <- sum(rf_pc_gen$pc_mean)
```
(Could put acroporids in a different color to show restored species? Maybe vs species susceptible to SCTLD?)
```{r}
ggplot(rf_pc_gen_site %>%
         filter(cat_c == "Hard corals") %>%
         mutate(restoration = case_when(genus == "Acropora" ~ "Outplanted",
                                         genus != "Acropora" ~ "Not outplanted"))
         ,
       aes(x = relevel(factor(monitoring), "baseline"), 
           y = pc_mean, 
           fill = factor(restoration, levels = c("Outplanted", "Not outplanted")))) +
  geom_col(color = "black", alpha = 0.9) +
  scale_fill_manual(values = c("goldenrod1", "coral")) +
  geom_errorbar(aes(ymin = pc_mean - pc_se, ymax = pc_mean + pc_se), width = .2,
                 position = position_dodge(.9)) +
  facet_grid(site ~ genus) +
  labs(y = "Mean percent cover", x = "", fill = "") +
  scale_x_discrete(labels = c('Base.','6 mo.')) +
  theme_bw() + # need to make genus italics/+ spp.
  theme(legend.position = "bottom")
ggsave(here("restoration_monitoring", "figs", "6mo_corals by genotype.png"), width = 8, height = 4)

ggplot(rf_pc_gen_site %>%
         filter(cat_c == "Hard corals" & genus == "Acropora"),
       aes(x = relevel(factor(monitoring), "baseline"), 
           y = pc_mean)) +
  geom_col(color = "black", fill = "goldenrod1", alpha = 0.9) +
  geom_errorbar(aes(ymin = pc_mean - pc_se, ymax = pc_mean + pc_se), width = .2,
                 position = position_dodge(.9)) +
  facet_grid(. ~ site) +
  labs(y = "Mean percent cover", x = "", fill = "") +
  scale_x_discrete(labels = c('Base.','6 mo.')) +
  theme_bw() + # need to make genus italics/+ spp.
  theme(legend.position = "bottom")
ggsave(here("restoration_monitoring", "figs", "6mo_acroporids.png"), width = 5, height = 2.5)
```

### Diadema
(Not essential, but maybe interesting?)

```{r}
diadema <- read_excel(here("restoration_monitoring", "data_raw", "ecological monitoring.xlsx"), sheet = "inverts") %>%
  clean_names() %>%
  rename(transect_og = transect) %>%
  unite(transect, date, transect_og, sep = "_", remove = FALSE) %>% # making sure transects don't get combined if they were surveyed on different dates (for combining TPB baseline with MMA data)
  mutate(date = ymd(date),
         # date_mon = format_ISO8601(ymd(date), precision = "ym"),
         monitoring = case_when(year(date) <= 2022 ~ "baseline",
                                date %within% interval(ymd("2023-01-01"), ymd("2023-06-01")) ~ "6month")) %>%
  select(monitoring, site, transect, diadema_adult, diadema_juvenile) %>%
  pivot_longer(cols = starts_with("diadema"),
               names_to = "phase",
               names_pattern = "diadema_(.*)",
               values_to = "count")

# average within sites
diadema_site <- diadema %>%
  group_by(monitoring, site) %>%
  summarize(density_mean = mean(count/10), # 1x10m belt
            density_se = sd(count/10)/sqrt(n())) 

ggplot(diadema_site,
       aes(x = relevel(factor(monitoring), "baseline"), y = density_mean)) +
  geom_col(fill = "black", position = "dodge") +
  scale_fill_manual(values = c("gray10", "gray90"), name = "Phase", labels = c("Adult", "Juvenile")) +
  geom_errorbar(aes(ymin = density_mean - density_se, ymax = density_mean + density_se), width = .2,
                 position = position_dodge(.9)) +
  facet_wrap(. ~ site) +
  labs(x = "", y = expression(Density~(indv.~m^-2))) +
  scale_x_discrete(labels = c('Baseline','6 mo.')) +
  theme_bw()
  #theme(axis.text.x = element_text(angle = 45, hjust = 1))

# ggsave(here("restoration_monitoring", "figs", "selection_diadema.png"), height = 4, width = 5)
```

## Fish

```{r}
# import and clean data

rf_fish <- read_excel(here("restoration_monitoring", "data_raw", "ecological monitoring.xlsx"), sheet = "fish") %>% 
  clean_names() %>%
  filter(!is.na(site)) %>% # remove any incomplete rows at end of data
  filter(species_code != "squid") %>% # removing non-fish
  rename(transect_og = transect) %>%
  unite(transect, date, transect_og, sep = "_", remove = FALSE) %>% # making sure transects don't get combined if they were surveyed on different dates (for combining TPB baseline with MMA data)
  mutate(date = ymd(date),
         # date_mon = format_ISO8601(ymd(date), precision = "ym"),
         monitoring = case_when(year(date) <= 2022 ~ "baseline",
                                date %within% interval(ymd("2023-01-01"), ymd("2023-06-01")) ~ "6month")) %>%
  mutate(number = if_else(is.na(number), 1, number)) %>% # all entries with no number specified were single observations
  uncount(number) %>% # expand to replicate rows if multiple fish were recorded to look at length distributions
  mutate(biomass = as.numeric(biomass),
         phase_code = tolower(phase))

n_transects <- rf_fish %>%
  select(monitoring, site, transect) %>%
  distinct() %>%
  group_by(monitoring, site) %>%
  summarize(n = n())

belt_area <- 30*2 # 30m x 2m belt transects
```

### Total biomass by site and monitoring event
(Personally think this may not be relevant yet as it would be a fairly short time frame to impact total biomass?)
```{r}
# calculating total biomass in each transect (sum) -> mean at each site
rf_fish_site <- rf_fish %>% 
  group_by(monitoring, site, transect) %>%
  summarize(bm_tot = sum(biomass)/1000/belt_area*10000) %>% #kg/ha
  group_by(monitoring, site) %>%
  summarize(bm_mean = mean(bm_tot),
            bm_se = sd(bm_tot)/sqrt(n()))

# graph
ggplot(rf_fish_site, aes(x = relevel(factor(monitoring), "baseline"), y = bm_mean)) +
  geom_col(fill = "cadetblue3", color = "black", alpha = 0.8) +
  geom_errorbar(aes(ymin = bm_mean - bm_se, ymax = bm_mean + bm_se), width = .2,
                 position = position_dodge(.9)) +
  facet_wrap(. ~ site) +
  labs(x = "", y = expression(Total~fish~biomass~(kg~ha^-1))) +
  scale_x_discrete(labels = c('Baseline','6 mo.')) +
  theme_bw()

# ggsave(here("restoration_monitoring", "figs", "selection_fish_bm.png"), width = 3, height = 2)
```

### Juvenile herbivore (scarid + acanthurid) densities
(This seems more relevant)
```{r}
rf_jvherb_dens_site <- rf_fish %>% 
  filter(family %in% c("Scaridae", "Acanthuridae") & length <= 8) %>%
  group_by(monitoring, site, transect) %>%
  summarize(dens_t = n()/belt_area) %>% #indv/m
  group_by(monitoring, site) %>%
  summarize(dens_mean = mean(dens_t),
            dens_se = sd(dens_t)/sqrt(n()))

ggplot(rf_jvherb_dens_site,
       aes(x = relevel(factor(monitoring), "baseline"), y = dens_mean)) +
  geom_col(color = "black", fill = "cadetblue3", alpha = 0.8, stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = dens_mean - dens_se, ymax = dens_mean + dens_se), width = .2, color = "gray40",
                 position = position_dodge(.9)) +
  facet_grid(. ~ site) +
  labs(x = "", y = expression(Juvenile~herbivore~density~(indv.~m^-2))) +
  scale_x_discrete(labels = c('Baseline','6 mo.')) +
  theme_bw()
ggsave(here("restoration_monitoring", "figs", "6mo_jv herb fish.png"), width = 5, height = 4)
```