---
title: "Nursery monitoring"
author: "Molly Wilson"
date: "2023-09-02"
output: html_document
---

```{r, message = F, warning = F, echo = F}
library(tidyverse)
library(here) 
library(readxl) 
library(janitor)
library(snakecase) # for adjusting capitalization of text within data (e.g., species names)
library(knitr) # for including tables
library(stringr)
library(lubridate)

knitr::opts_chunk$set(message = FALSE, warning = FALSE)


nursery_mon <- read_excel(here("restoration_monitoring", "data_raw", "Genotype tracking.xlsx"), sheet = "nursery monitoring") %>% 
  clean_names() %>%
  mutate(date = ymd(date),
         species = case_when(substr(genotype, 1, 4) == "ACER" ~ "A. cervicornis",
                             substr(genotype, 1, 4) == "APRO" ~ "A. prolifera",
                             substr(genotype, 1, 4) == "APAL" ~ "A. palmata",
                             substr(genotype, 1, 4) == "PFUR" ~ "P. furcata",
                             substr(genotype, 1, 4) == "OANN" ~ "O. annularis"),
         ) %>%
  filter(!is.na(date)) %>%
  mutate_at(c('n_healthy', 'n_dead', 'n_bleached', 'n_pale', 'n_diseased', 'n_other'), ~replace_na(.,0))
```

Summarizing results from the 2024 bleaching event
```{r}
n_tot <- nursery_mon %>% # total number of fragments on 09/25
  filter(date == "2023-09-25") %>%
  pivot_longer(cols = n_healthy:n_diseased, names_to = "condition", values_to = "n_frags") %>%
  summarize(n_tot = sum(n_frags))

therm_stress_summary <- nursery_mon %>%
  filter(date == max(date)) %>%
  select(n_healthy, n_bleached, n_pale, n_diseased) %>%
  summarize(p_healthy = sum(n_healthy)/n_tot$n_tot*100,
            p_diseased = sum(n_diseased)/n_tot$n_tot*100,
            p_thermstressed = (sum(n_bleached + n_pale))/n_tot$n_tot*100) %>%
  mutate(p_dead = 100 - (p_healthy + p_diseased + p_thermstressed))

therm_stress_time <- nursery_mon %>%
  select(date, n_healthy, n_bleached, n_pale, n_diseased) %>%
  group_by(date) %>%
  summarize(Healthy = sum(n_healthy)/n_tot$n_tot*100,
            Bleached = sum(n_bleached)/n_tot$n_tot*100,
            Pale = sum(n_pale)/n_tot$n_tot*100) %>%
  mutate(Dead = 100 - (Healthy + Bleached + Pale)) %>%
  pivot_longer(cols = Healthy:Dead, names_to = "condition", values_to = "percent")
factor(therm_stress_time$condition, levels = c("Dead", "Bleached", "Pale", "Healthy"))

ggplot(therm_stress_time,
       aes(x = as.factor(date), y = percent, group = condition)) +
  geom_bar(aes(fill = condition), position = "stack", stat = 'identity', color = "black") +
  scale_fill_manual(breaks = c('Dead', 'Bleached', 'Pale', 'Healthy'), values = c("slategray", "azure1", "tan", "goldenrod")) +
  labs(y = "Percent of fragments", x = "Date", fill = "Condition") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

Genotype-specific results of the 2024 bleaching event
- survivorship
- % therm tolerant (no bleaching or paling) out of living fragments

Notes for future improvement:
- currently doesn't distinguish between heat-related death vs. disease, etc.
- more checks and balances to ensure that no data collected doesn't equate to 0 corals
- how to account for times when structures are fluffled, etc. (i.e., number of fragments increase)
- disease monitoring can be ongoing - and relative to actual number of fragments, while bleaching monitoring would only occur in the fall (when structures are not refragged etc?)

```{r}
int_bleaching_2023 <- interval(ymd("2023-09-01"), ymd("2023-12-10")) # date range for investigation of 2023 bleaching

thermtol_geno_date <- nursery_mon %>%
  filter(date == "2023-09-25") %>%
  group_by(genotype) %>%
  summarize(n_og = sum(n_healthy + n_dead + n_bleached + n_pale + n_diseased + n_other)) %>%
  right_join(nursery_mon %>%
               filter(date %within% int_bleaching_2023) %>%
               group_by(genotype, date) %>%
               summarize(n_healthy = sum(n_healthy), n_bleached = sum(n_bleached), n_pale = sum(n_pale),  n_dead = sum(n_dead), n_diseased = sum(n_diseased))) %>%
  mutate(p_thermtolerant = n_healthy / n_og) 

thermtol_geno <- thermtol_geno_date %>%
  group_by(genotype) %>%
  summarize(min_thermtolerance = min(p_thermtolerant)) %>% # need to figure out PCLI01 issue
  left_join(thermtol_geno_date %>%
              filter(date == "2023-12-04") %>%
              mutate(survivorship = n_healthy / n_og) %>%
              select(genotype, survivorship)
            ) %>%
  mutate(survivorship = replace_na(survivorship, 0))
write.csv(thermtol_geno, here("restoration_monitoring", "data_outputs", "thermtol_2023.csv"), row.names=FALSE)
```












```{r}
# archive...
therm_stress_geno <- nursery_mon %>%
  filter(date == "2023-09-25" & !location %in% c("C1", "D1", "E1", "F1")) %>% # excluding genebanks
  mutate(n_og = n_healthy + n_dead + n_bleached + n_pale + n_diseased) %>%
  select(nursery, location, genotype, n_og) %>%
  left_join(nursery_mon %>%
              filter(date == "2023-12-04") %>% # comparing to December 2023 data collection
              select(nursery, location, genotype, n_healthy, n_bleached, n_pale, n_diseased), 
            by = c("nursery", "location", "genotype")) %>%
  mutate_at(c('n_healthy', 'n_bleached', 'n_pale', 'n_diseased'), ~replace_na(.,0)) %>%
  mutate(p_healthy = n_healthy/n_og*100,
         p_bleached = n_bleached/n_og*100,
         p_pale = n_pale/n_og*100,
         p_diseased = n_diseased/n_og*100,
         p_dead = (n_og - (n_healthy + n_bleached + n_pale + n_diseased))/n_og*100) %>%
  group_by(genotype) %>%
  summarise(dead_percent = mean(p_dead), # took averages across structures for each genotype
            dead_se = sd(p_dead)/sqrt(n()),
            healthy_percent = mean(p_healthy),
            healthy_se = sd(p_healthy)/sqrt(n()),
            bleached_percent = mean(p_bleached),
            bleached_se = sd(p_bleached)/sqrt(n()),
            pale_percent = mean(p_pale),
            pale_se = sd(p_pale)/sqrt(n()),
            diseased_percent = mean(p_diseased),
            diseased_se = sd(p_diseased)/sqrt(n()),
            thermstressed_percent = mean(p_bleached + p_pale),
            thermstressed_se = sd(p_bleached + p_pale)/sqrt(n()))
write.csv(therm_stress_geno, here("restoration_monitoring", "data_outputs", "bleaching_outcomes.csv"))

therm_stress_long_temp <- therm_stress_geno %>%
  pivot_longer(cols = -genotype,
                 names_to = c('variable', '.value'),
                 names_sep = '_') %>%
  rename(condition = variable) 

therm_stress_long <- therm_stress_long_temp %>%
  left_join(therm_stress_long_temp %>%
              filter(condition == "dead") %>%
              arrange(percent) %>%
              mutate(order = row_number()) %>%
              select(genotype, order))

ggplot(therm_stress_long %>%
         filter(condition %in% c("healthy", "dead", "thermstressed", "diseased")) %>%
         mutate(genotype = fct_reorder(genotype, order)),
       aes(x = genotype, y = percent, group = condition)) +
  geom_col(aes(fill = condition)) +
  scale_fill_manual(values = c("darkseagreen", "coral", "goldenrod", "azure1")) +
  labs(x = "Genotype", y = "Percent of fragments", fill = "Condition") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
ggsave(here("restoration_monitoring", "figs", "bleaching_outcomes_23.12.04.png"), width = 12, height = 5)
```

```{r}
# rtl - mean rate of infection (number of diseased frags/total frags) over 6 weeks and across structures

rtl <- nursery_mon %>%
  filter(species %in% c("A. cervicornis", "A. prolifera", "A. palmata")) %>%
  filter(date == "2023-09-25" & !location %in% c("C1", "D1", "E1", "F1")) %>%
  mutate(n_tot = n_healthy + n_dead + n_bleached + n_pale + n_diseased,
         p_diseased = n_diseased/n_tot) %>%
  group_by(nursery, location, genotype) %>%
  summarise(p_diseased = mean(p_diseased)) %>%
  group_by(genotype) %>%
  summarise(p_diseased_mean = mean(p_diseased),
            p_diseased_se = sd(p_diseased)/sqrt(n()))
  
```


```{r, eval = FALSE}
### OG code...

nursery_structures <- read.csv(here("restoration_monitoring", "data_outputs", "nursery locations.csv")) %>%
  mutate_at(c('n_trees', 'n_ropes'), ~replace_na(.,0)) %>%
  mutate(n_structures = n_trees + n_ropes,
         species = case_when(substr(genotype, 1, 4) == "ACER" ~ "A. cervicornis",
                             substr(genotype, 1, 4) == "APRO" ~ "A. prolifera",
                             substr(genotype, 1, 4) == "APAL" ~ "A. palmata",
                             substr(genotype, 1, 4) == "PFUR" ~ "P. furcata",
                             substr(genotype, 1, 4) == "OANN" ~ "O. annularis")) %>%
  select(genotype, species, n_structures)

disease_genotype <- nursery_mon %>%
  filter(!str_detect(structure_location, "^ST|^DT|^Cotton|^Deep") & condition == "disease" & level_affected == "high (>50%)") %>%
  filter(!str_detect(genotype, "/")) %>%
  left_join(nursery_structures, by = "genotype") %>%
  filter(!is.na(level_affected)) %>%
  group_by(genotype, n_structures) %>% # can add level_affected here
  summarize(outbreaks = n()) %>%
  mutate(outbreaks_per_structure = outbreaks/n_structures) %>%
  filter(!is.na(outbreaks_per_structure)) %>%
  right_join(nursery_structures, by = c("genotype", "n_structures")) %>%
  mutate_at('outbreaks_per_structure', ~replace_na(.,0)) %>%
  mutate(species = case_when(substr(genotype, 1, 4) == "ACER" ~ "A. cervicornis",
                             substr(genotype, 1, 4) == "APRO" ~ "A. prolifera",
                             substr(genotype, 1, 4) == "APAL" ~ "A. palmata",
                             substr(genotype, 1, 4) == "PFUR" ~ "P. furcata",
                             substr(genotype, 1, 4) == "OANN" ~ "O. annularis"),
         )

disease_spp <- nursery_mon %>%
  filter(!str_detect(structure_location, "^ST|^DT|^Cotton|^Deep") & condition == "disease" & level_affected == "high (>50%)") %>%
  group_by(species) %>%
  summarize(outbreaks = n()) %>%
  left_join(nursery_structures %>%
              group_by(species) %>%
              summarize(tot_structures = sum(n_structures))) %>%
  mutate(outbreaks_per_structure = outbreaks/tot_structures) %>%
  filter(!is.na(outbreaks_per_structure))


ggplot(data = disease_spp, 
       aes(x = species, y = outbreaks_per_structure)) +
  geom_col(color = "black", fill = "goldenrod", alpha = 0.8) +
  labs(y = "Outbreaks per nursery structure", x = "") +
  theme_bw() +
  theme(axis.text.x = element_text(face = "italic"))
ggsave(here("restoration_monitoring", "figs", "RTL_2023.png"), width = 5.5, height = 3)
  
```


