---
title: "Nursery monitoring"
author: "Molly Wilson"
date: "2023-09-02"
output: 
  html_document:
    code_folding: hide
---
To do:
- Update over-time graphs
- Update genotype-specific graphs
- Compare 2023 vs 2024 bleaching performance by genotype?
- Incorporate Hobo data here?

```{r, message = F, warning = F, echo = F}
library(tidyverse)
library(here) 
library(readxl) 
library(janitor)
library(snakecase) # for adjusting capitalization of text within data (e.g., species names)
library(knitr) # for including tables
library(stringr)
library(lubridate)

knitr::opts_chunk$set(message = FALSE, warning = FALSE)

nursery_mon <- read_excel(here("restoration_monitoring", "data_raw", "Genotype tracking.xlsx"), sheet = "nursery monitoring") %>% 
  clean_names() %>%
  mutate(date = ymd(date),
         date_floor = floor_date(ymd(date), unit = "month"),
         month = month(ymd(date), label = TRUE),
         year = year(ymd(date)),
         date_label = paste(as.character(month), as.character(year))
         ) %>%
    filter(!is.na(date)) %>% # removing any extra lines
  mutate_at(c('n_healthy', 'n_dead', 'n_bleached', 'n_pale', 'n_diseased', 'n_other'), ~replace_na(.,0))

# determine date references

date_check <- nursery_mon %>%
  select(date, date_floor, date_label, nursery) %>%
  distinct() # making sure nurseries were not monitored more than once per date floor/date label group

pre_bleaching_baseline_2024 <- ymd("2024-07-01")
post_bleaching_2024 <- last(nursery_mon$date_floor) # eventually cap this when bleaching subsides
last_mon_month <- last(nursery_mon$date_floor)
```

Summarizing current fragment status by nursery
```{r}
status_nursery <- nursery_mon %>%
  filter(date_floor == last_mon_month) %>%
  select(nursery, n_healthy, n_bleached, n_pale, n_diseased, n_other, n_dead) %>%
  group_by(nursery) %>%
  summarize(n_healthy = sum(n_healthy + n_other), # "other" frags consolidated here for simplicity
            n_pale = sum(n_pale),
            n_bleached = sum(n_bleached),
            n_diseased = sum(n_diseased),
            n_dead = sum(n_dead)
            ) %>% 
  mutate(n_tot = n_healthy + n_pale + n_bleached + n_diseased + n_dead) %>%
  adorn_totals("row") %>%
  mutate(p_healthy = n_healthy/n_tot*100, # relative to current total
         p_diseased = n_diseased/n_tot*100, 
         p_pale = n_pale/n_tot*100,
         p_bleached = n_bleached/n_tot*100,
         p_dead = n_dead/n_tot*100
         )

# plots

ggplot(status_nursery %>%
         select(nursery, n_healthy, n_diseased, n_pale, n_bleached, n_dead) %>%
         pivot_longer(cols = n_healthy:n_dead, 
                      names_to = "condition", 
                      names_prefix = 'n_',
                      values_to = "n_frags") %>%
         filter(nursery != "Total") %>%
         mutate(condition = fct_relevel(str_to_sentence(condition), c("Dead", "Diseased", "Bleached", "Pale", "Healthy"))) %>%
         arrange(nursery),
       aes(x = nursery, y = n_frags, fill = condition)) +
  geom_col(color = "black") +
  scale_fill_manual(values = c("slategray", "coral", "azure1", "tan", "goldenrod")) +
  labs(y = "Number of fragments", x = "Nursery", fill = "Condition", , title = paste0("Status of fragments in ", month(ymd(last_mon_month), label = TRUE), " ", year(ymd(last_mon_month)))) +
  theme_bw() +
  theme(plot.title = element_text(hjust=0.5))

ggsave(here("restoration_monitoring", "figs", "current_status_nursery.png"), width = 12, height = 5)
```


Summarizing 2024 bleaching outcomes by nursery
```{r}
# determine nursery-level pre-bleaching baseline numbers
status_nursery_pre_bleaching_2024 <- nursery_mon %>%
  filter(date_floor == pre_bleaching_baseline_2024) %>% # manually check here that there are only one monitoring event per nursery in this date group
  group_by(nursery) %>%
  summarize(n_tot_baseline = sum(n_healthy + n_other + n_pale)) # "other" frags consolidated here for simplicity, should not be significant numbers of bleached in baseline

# determining condition of fragments at post-bleaching monitoring dates
status_nursery_post_bleaching_2024 <- nursery_mon %>%
  filter(date_floor == post_bleaching_2024) %>%
  select(nursery, n_healthy, n_bleached, n_pale, n_diseased, n_other) %>%
  group_by(nursery) %>%
  summarize(n_healthy = sum(n_healthy + n_other), # "other" frags consolidated here for simplicity
            n_pale = sum(n_pale),
            n_bleached = sum(n_bleached),
            n_diseased = sum(n_diseased)
            ) %>% 
  mutate(n_tot_alive = n_healthy + n_pale + n_bleached + n_diseased) %>%
  left_join(status_nursery_pre_bleaching_2024 %>%
              select(nursery, n_tot_baseline)) %>%
  mutate(n_dead_rel_baseline = n_tot_baseline-n_tot_alive) %>%
  select(nursery, n_tot_baseline, n_tot_alive, n_healthy, n_diseased, n_pale, n_bleached, n_dead_rel_baseline) %>%
  adorn_totals("row") %>%
  mutate(p_healthy = n_healthy/n_tot_alive*100, # relative to current total
         p_diseased = n_diseased/n_tot_alive*100, 
         p_pale = n_pale/n_tot_alive*100,
         p_bleached = n_bleached/n_tot_alive*100,
         p_healthy_rel_baseline = n_healthy/n_tot_baseline*100, # relative to baseline total
         p_diseased_rel_baseline = n_diseased/n_tot_baseline*100,
         p_pale_rel_baseline = n_pale/n_tot_baseline*100,
         p_bleached_rel_baseline = n_bleached/n_tot_baseline*100,
         p_dead_rel_baseline = (n_tot_baseline-n_tot_alive)/n_tot_baseline*100,
         p_survival_rel_baseline = n_tot_alive/n_tot_baseline*100
         ) %>%
  select(nursery, n_tot_alive, n_healthy, n_diseased, n_pale, n_bleached, n_dead_rel_baseline, p_healthy, p_diseased, p_pale, p_bleached, p_healthy_rel_baseline, p_diseased_rel_baseline, p_pale_rel_baseline, p_bleached_rel_baseline, p_dead_rel_baseline, p_survival_rel_baseline)

# plots

ggplot(status_nursery_post_bleaching_2024 %>%
         select(nursery, healthy = p_healthy_rel_baseline, bleached = p_bleached_rel_baseline, pale = p_pale_rel_baseline, diseased = p_diseased_rel_baseline, dead = p_dead_rel_baseline) %>%
         pivot_longer(cols = healthy:dead, 
                      names_to = "condition",
                      values_to = "percent")  %>%
         mutate(nursery = fct_relevel(nursery, c('Total','York','TPB', 'DN', 'Cades')),
                condition = fct_relevel(str_to_sentence(condition), c("Dead", "Diseased", "Bleached", "Pale", "Healthy"))) %>%
         arrange(nursery),
       aes(x = nursery, y = percent, fill = condition)) +
  geom_col(color = "black") +
  scale_fill_manual(values = c("slategray", "coral", "azure1", "tan", "goldenrod")) +
  labs(y = "Percent of fragments", 
       x = "Nursery", fill = "Condition", 
       title = paste0("Fragment status in ",  
                      month(ymd(post_bleaching_2024), label = TRUE), " ", 
                      year(ymd(post_bleaching_2024)), " ",
                      "relative to ", 
                      month(ymd(pre_bleaching_baseline_2024), label = TRUE), " ", 
                      year(ymd(pre_bleaching_baseline_2024)))) +
  theme_bw() +
  theme(plot.title = element_text(hjust=0.5))

ggsave(here("restoration_monitoring", "figs", "bleaching_outcomes_2024_nursery.png"), width = 12, height = 5)
```

Summarizing 2024 bleaching by genotype
```{r}
# determine genotype-level pre-bleaching baseline numbers
status_genotype_baseline <- nursery_mon %>%
  filter(date_label == pre_bleaching_baseline_2024) %>%
  group_by(genotype) %>%
  summarize(n_healthy = sum(n_healthy),
            n_diseased = sum(n_diseased),
            n_pale = sum(n_pale),
            n_bleached = sum(n_bleached),
            n_other = sum(n_other)
            ) %>% 
  mutate(n_tot_baseline = n_healthy + n_diseased + n_pale + n_bleached + n_other)

# nursery snapshot at post-bleaching date, using baseline pre-bleaching data to calculate percent loss
status_genotype <- nursery_mon %>%
  filter(date_floor == post_bleaching_2024) %>% # most recent dates
  select(genotype, n_healthy, n_diseased, n_pale, n_bleached, n_other) %>%
  group_by(genotype) %>%
  summarize(n_healthy = sum(n_healthy),
            n_diseased = sum(n_diseased),
            n_pale = sum(n_pale),
            n_bleached = sum(n_bleached),
            n_other = sum(n_other)
            ) %>% 
  mutate(n_tot_alive = n_healthy + n_diseased + n_pale + n_bleached + n_other) %>%
  left_join(status_genotype_baseline %>%
              select(genotype, n_tot_baseline)) %>%
  mutate(n_dead_rel_baseline = n_tot_baseline-n_tot_alive) %>%
  select(genotype, n_tot_baseline, n_tot_alive, n_healthy, n_diseased, n_pale, n_bleached, n_dead_rel_baseline) %>%
  distinct() %>%
  mutate(p_healthy = n_healthy/n_tot_alive*100, # relative to current total
         p_diseased = n_diseased/n_tot_alive*100, 
         p_pale = n_pale/n_tot_alive*100,
         p_bleached = n_bleached/n_tot_alive*100,
         p_healthy_rel_baseline = n_healthy/n_tot_baseline*100, # relative to baseline total
         p_diseased_rel_baseline = n_diseased/n_tot_baseline*100,
         p_pale_rel_baseline = n_pale/n_tot_baseline*100,
         p_bleached_rel_baseline = n_bleached/n_tot_baseline*100,
         p_dead_rel_baseline = (n_tot_baseline-n_tot_alive)/n_tot_baseline*100,
         p_survival_rel_baseline = n_tot_alive/n_tot_baseline*100
         ) %>%
  select(genotype, n_tot_alive, n_healthy, n_diseased, n_pale, n_bleached, n_dead_rel_baseline, p_healthy, p_diseased, p_pale, p_bleached, p_healthy_rel_baseline, p_diseased_rel_baseline, p_pale_rel_baseline, p_bleached_rel_baseline, p_dead_rel_baseline, p_survival_rel_baseline)
### need to figure out negative numbers/NAs here ^

ggplot(status_genotype %>%
         select(genotype, healthy = p_healthy_rel_baseline, bleached = p_bleached_rel_baseline, pale = p_pale_rel_baseline, diseased = p_diseased_rel_baseline, dead = p_dead_rel_baseline) %>%
         pivot_longer(cols = -genotype,
                 names_to = "condition",
                 values_to = "percent"),
       aes(x = genotype, y = percent, group = condition)) +
  geom_col(aes(fill = condition)) +
  scale_fill_manual(breaks = c('dead', 'diseased', 'bleached', 'pale', 'healthy'), values = c("slategray", "coral", "azure1", "tan", "goldenrod")) +
  labs(x = "Genotype", 
       y = "Percent of fragments", 
       fill = "Condition",
       title = paste0("Fragment status in ",  
                      month(ymd(post_bleaching_2024), label = TRUE), " ", 
                      year(ymd(post_bleaching_2024)), " ",
                      "relative to ", 
                      month(ymd(pre_bleaching_baseline_2024), label = TRUE), " ", 
                      year(ymd(pre_bleaching_baseline_2024)))) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

ggsave(here("restoration_monitoring", "figs", "bleaching_outcomes_2024_genotype.png"), width = 12, height = 5)
```


Graphing nursery heat stress over time
```{r, eval = F}
### need to redo this by nursery or with mechanism for combining dates
therm_stress_time <- nursery_mon %>%
  select(date_floor, date_label, n_healthy, n_bleached, n_pale, n_diseased, n_dead) %>%
  mutate(n_tot_frags = n_healthy + n_bleached + n_pale + n_diseased + n_dead) %>%
  group_by(date_floor, date_label) %>%
  summarize(n_tot = sum(n_tot_frags),
            Healthy = sum(n_healthy)/n_tot*100,
            Bleached = sum(n_bleached)/n_tot*100,
            Pale = sum(n_pale)/n_tot*100,
            Diseased = sum(n_diseased)/n_tot*100,
            Dead = sum(n_dead)/n_tot*100
            ) %>%
  pivot_longer(cols = Healthy:Dead, names_to = "condition", values_to = "percent") %>%
  mutate(condition = fct_relevel(condition, c("Dead", "Diseased", "Bleached", "Pale", "Healthy")))

ggplot(therm_stress_time,
       aes(x = date_floor, y = percent, group = condition)) +
  geom_bar(aes(fill = condition),
               position = "stack", stat = 'identity', color = "black") +
  scale_fill_manual(values = c("slategray", "coral", "azure1", "tan", "goldenrod")) +
  scale_x_date(date_labels = "%Y-%b", date_breaks  = "1 month") +
  labs(y = "Percent of fragments", x = "Date", fill = "Condition") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

Genotype-specific results of 2023 bleaching
### Need to repeat for 2024
```{r}
int_bleaching_2023 <- interval(ymd("2023-09-01"), ymd("2023-12-10")) # date range for investigation of 2023 bleaching

thermtol_geno_date <- nursery_mon %>%
  filter(date == "2023-09-25") %>%
  group_by(genotype) %>%
  summarize(n_og = sum(n_healthy + n_dead + n_bleached + n_pale + n_diseased + n_other)) %>%
  right_join(nursery_mon %>%
               filter(date %within% int_bleaching_2023) %>%
               group_by(genotype, date) %>%
               summarize(n_healthy = sum(n_healthy), n_bleached = sum(n_bleached), n_pale = sum(n_pale),  n_dead = sum(n_dead), n_diseased = sum(n_diseased), n_other = sum(n_other))) %>%
  mutate(p_thermtolerant = n_healthy / n_og) 

thermtol_geno_2023 <- thermtol_geno_date %>%
  group_by(genotype) %>%
  summarize(min_thermtolerance = min(p_thermtolerant)) %>%
  left_join(thermtol_geno_date %>%
              filter(date == "2023-12-04") %>%
              mutate(survivorship = n_healthy / n_og) %>%
              select(genotype, survivorship)
            ) %>%
  mutate(survivorship = replace_na(survivorship, 0))

write.csv(thermtol_geno_2023, here("restoration_monitoring", "data_outputs", "thermtol_2023.csv"), row.names=FALSE)
```

Disease susceptibility
- rate of infection: number of infected frags / total frags across all structures

```{r}
### need to account for multiple dates - or combine dates
disease <- nursery_mon %>%
  group_by(genotype, date) %>%
  summarize(n_healthy = sum(n_healthy), n_bleached = sum(n_bleached), n_pale = sum(n_pale),  n_dead = sum(n_dead), n_diseased = sum(n_diseased), n_other = sum(n_other)) %>%
  mutate(p_diseased = n_diseased / (n_diseased + n_healthy + n_bleached + n_pale + n_other)) %>% # not relative to dead fragments
  mutate(across(p_diseased, ~ replace(., is.nan(.), 0))) %>%
  group_by(genotype) %>%
  summarize(mean_infection = mean(p_diseased)) # need to check APAL14

write.csv(disease, here("restoration_monitoring", "data_outputs", "disease.csv"), row.names=FALSE)
  
```

Broodstock survival
```{r}
#Cades: A4
#TPB: A4

broodstock <- nursery_mon %>%
  filter((nursery == "Cades" & location == "A4") | (nursery == "TPB" & location == "A4")) %>%
  mutate(n_tot = n_healthy + n_diseased + n_bleached + n_pale) %>%
  group_by(genotype) %>%
  summarize()
```












```{r, EVAL = F}
# archive...
# 
# 
# 
# # species codes
#          species = case_when(substr(genotype, 1, 4) == "ACER" ~ "A. cervicornis",
#                              substr(genotype, 1, 4) == "APRO" ~ "A. prolifera",
#                              substr(genotype, 1, 4) == "APAL" ~ "A. palmata",
#                              substr(genotype, 1, 4) == "PPOR" ~ "P. porites",
#                              substr(genotype, 1, 4) == "PDIV" ~ "P. divaricata",
#                              substr(genotype, 1, 4) == "OANN" ~ "O. annularis",
#                              substr(genotype, 1, 4) == "OFAV" ~ "O. faveolata",
#                              substr(genotype, 1, 4) == "OFRA" ~ "O. franksi",
#                              substr(genotype, 1, 4) == "PSTR" ~ "P. strigosa",
#                              substr(genotype, 1, 4) == "PCLI" ~ "P. clivosa",
#                              substr(genotype, 1, 4) == "CNAT" ~ "C. natans",
#                              substr(genotype, 1, 4) == "DLAB" ~ "D. labyrinthiformis",
#                              substr(genotype, 1, 4) == "DCYL" ~ "D. cylindrus",
#                              substr(genotype, 1, 4) == "MCAV" ~ "M. cavernosa"),
```







