---
title: "Nursery monitoring"
author: "Molly Wilson"
date: "2023-09-02"
output: 
  html_document:
    code_folding: hide
---
To do:

- Compare 2023 vs 2024 bleaching performance by genotype (need to fix data errors with MB)
- Update disease susceptibility with 2024 data


```{r, message = F, warning = F, echo = F}
library(tidyverse)
library(here) 
library(readxl) 
library(janitor)
library(snakecase) # for adjusting capitalization of text within data (e.g., species names)
library(lubridate)
library(knitr) # for including tables
library(stringr)
library(rstatix)
library(dunn.test)
library(ggpubr)
library(ggbeeswarm) # beeswarm plots

knitr::opts_chunk$set(message = FALSE, warning = FALSE)
```

## Nursery temperature status

```{r}
cades <- read_excel(here("restoration_monitoring", "data_raw", "hobo", "Cades.xlsx"), 
                  sheet = 1) %>%
  clean_names() %>%
  dplyr::rename(date = 2, temp_c = 3) %>%
  select(-number) %>%
  mutate(nursery = "Cades")

tpb <- read_excel(here("restoration_monitoring", "data_raw", "hobo", "Ten pound bay.xlsx"), 
                  sheet = 1) %>%
  clean_names() %>%
  dplyr::rename(date = 2, temp_c = 3) %>%
  select(-number) %>%
  mutate(nursery = "Ten Pound Bay")

deep <- read_excel(here("restoration_monitoring", "data_raw", "hobo", "Deep water.xlsx"), 
                  sheet = 1) %>%
  clean_names() %>%
  dplyr::rename(date = 2, temp_c = 3) %>%
  select(-number) %>%
  mutate(nursery = "Deep Nursery")
# deep nursery monitor was moved to York on 11/26/24, reformatted on 12/2/24, logger dysfunctional as of early 2025

temp_data <- rbind(cades, tpb) %>%
  filter(temp_c < 31.99 & temp_c > 20) %>% # removing outliers from loggers being on the boat and heating up - check this doesn't affect data during heat events
  filter(as.Date(date) < "2025-06-30") #for July 2025 report

# references
lab_temp <- expression("Temperature ("*~degree*C*")")
ref_temp = data.frame(xmin = -Inf, xmax = Inf, ymin = 23, ymax = 29.6)
lab_ref_temp <- "Target range (Coral Reef Alliance)"
ref_temp_bleaching = 30.63
lab_ref_temp_bleaching <- "Coral bleaching threshold (NOAA)"
ylims_temp <- c(22, (max(data_mon$temperature_c, na.rm = T) + 1))

# plot
ggplot() +
  geom_rect(data = ref_temp, 
              aes(xmin = as.POSIXct(-Inf), xmax = as.POSIXct(Inf), ymin = ymin, ymax = ymax, fill = lab_ref_temp),
              alpha = 0.5) +
  scale_fill_manual(values = "gray") +
  geom_line(data = temp_data,
            aes(x = date, y = temp_c, group = nursery, color = nursery),
            alpha = 0.9) +
  geom_hline(yintercept = ref_temp_bleaching, linetype = 'dashed') +
  scale_y_continuous(sec.axis = sec_axis(~.*9/5+32, name = "Temperature (째F)")) +
  labs(x = "", y = "Temperature (째C)", color = "Nursery location", fill = "") +
  theme_bw() +
  theme(legend.position = "top")
ggsave(here("restoration_monitoring", "figs", "hobo_summary.png"), width = 6, height = 4)

ggplot(temp_data,
       aes(x = date, y = temp_c)) +
  geom_line(alpha = 0.9, size = 0.3, color = "cadetblue") +
  geom_hline(yintercept = ref_temp_bleaching, linetype = 'dashed', color = "coral") +
  scale_y_continuous(sec.axis = sec_axis(~.*9/5+32, name = "Temperature (째F)")) +
  facet_grid(. ~ nursery) +
  labs(x = "", y = "Temperature (째C)", color = "Nursery") +
  theme_bw()
ggsave(here("restoration_monitoring", "figs", "hobo_nursery.png"), width = 6, height = 4)
  
```

```{r, message = F, warning = F, echo = F}
nursery_mon <- read_excel(here("restoration_monitoring", "data_raw", "Genotype tracking.xlsx"), sheet = "nursery monitoring") %>% 
  clean_names() %>%
  mutate(date = as.character(date),
         date = if_else(date == "2024-12-27", "2025-01-01", date),
         date = ymd(date),
         date_floor = floor_date(ymd(date), unit = "month"),
         month = month(ymd(date), label = TRUE),
         year = year(ymd(date)),
         date_label = paste(as.character(month), as.character(year))
         ) %>%
    filter(!is.na(date)) %>% # removing any extra lines
  mutate_at(c('n_healthy', 'n_dead', 'n_bleached', 'n_pale', 'n_diseased', 'n_other'), ~replace_na(.,0))

date_check <- nursery_mon %>%
  select(date, date_floor, date_label, nursery) %>%
  distinct() # making sure nurseries were not monitored more than once per date floor/date label group
```

## Summarizing current fragment status by nursery
```{r}
last_mon_month <- last(nursery_mon$date_floor)

status_nursery <- nursery_mon %>%
  filter(date_floor == last_mon_month) %>%
  select(nursery, n_healthy, n_bleached, n_pale, n_diseased, n_other, n_dead) %>%
  group_by(nursery) %>%
  summarize(n_healthy = sum(n_healthy + n_other), # "other" frags consolidated here for simplicity
            n_pale = sum(n_pale),
            n_bleached = sum(n_bleached),
            n_diseased = sum(n_diseased),
            n_dead = sum(n_dead)
            ) %>% 
  mutate(n_tot = n_healthy + n_pale + n_bleached + n_diseased + n_dead) %>%
  adorn_totals("row") %>%
  mutate(p_healthy = n_healthy/n_tot*100, # relative to current total
         p_diseased = n_diseased/n_tot*100, 
         p_pale = n_pale/n_tot*100,
         p_bleached = n_bleached/n_tot*100,
         p_dead = n_dead/n_tot*100
         )

write.csv(status_nursery, here("restoration_monitoring", "data_outputs", "status_nursery_2025Jan.csv"), row.names=FALSE)

# plots

ggplot(status_nursery %>%
         select(nursery, n_healthy, n_diseased, n_pale, n_bleached, n_dead) %>%
         mutate(nursery = case_when(nursery == "DN" ~ "Deep",
                                    nursery == "TPB" ~ "Ten Lb.",
                                    .default = nursery)) %>%
         pivot_longer(cols = n_healthy:n_dead, 
                      names_to = "condition", 
                      names_prefix = 'n_',
                      values_to = "n_frags") %>%
         filter(nursery != "Total") %>%
         mutate(condition = fct_relevel(str_to_sentence(condition), c("Dead", "Diseased", "Bleached", "Pale", "Healthy"))) %>%
         filter(condition != "Dead") %>% # removing dead for this graph
         arrange(nursery),
       aes(x = nursery, y = n_frags, fill = condition)) +
  geom_col(color = "black") +
  scale_fill_manual(values = c("coral", "azure1", "tan", "goldenrod")) + # dead color "slategray", 
  labs(y = "Number of fragments", x = "Nursery", fill = "Condition", , title = paste0("Status of fragments in ", month(ymd(last_mon_month), label = TRUE), ". ", year(ymd(last_mon_month)))) +
  theme_bw() +
  theme(plot.title = element_text(hjust=0.5))

ggsave(here("restoration_monitoring", "figs", "frag_status_nursery_Jan2025.png"), width = 8, height = 5)
```


## Summarizing 2024 bleaching outcomes by nursery

### Status of fragments by nursery, regardless of genotype

***make sure this is best way to incorporate adjustments - and that I don't have to split into 2 time periods?***

```{r}
# establish date references

pre_bleaching_baseline_2024 <- ymd("2024-07-01")
post_bleaching_2024 <- last(nursery_mon$date_floor) # eventually cap this when bleaching subsides

# import fragment adjustments based on nursery activities between July - Dec 2024

nursery_adjustments <- read_excel(here("restoration_monitoring", "data_raw", "Genotype tracking.xlsx"), sheet = "nursery adjustments") %>% 
  clean_names() %>%
  select(date, nursery, genotype, n_frags = change_in_frags)

outplant_adjustments <- read_excel(here("restoration_monitoring", "data_raw", "Genotype tracking.xlsx"), sheet = "outplants") %>% 
  clean_names() %>%
  filter(date == as.Date("2024-11-23") & site == "Cades Nursery") %>%
  mutate(genotype = substr(id, start = 1, stop = 6),
         date = as.Date("2024-12-01")) %>% # this will substitute December data
  group_by(date, nursery = site, genotype) %>%
  summarize(n_healthy = sum(n_frags)) %>%
  mutate(nursery = if_else(nursery == "Cades Nursery", "Cades", nursery),
         date_floor = floor_date(ymd(date), unit = "month"))

# import lab data for comparison

lab_frags <- read_excel(here("restoration_monitoring", "data_raw", "Nursery map.xlsx"), sheet = "Lab") %>% 
  clean_names() %>%
  select(genotype, n_frags_initial, n_frags_final) %>%
  pivot_longer(cols = n_frags_initial:n_frags_final,
               names_to = "date",
               values_to = "n_healthy") %>%
  mutate(date = case_when(date == "n_frags_initial" ~ as.Date("2024-07-01"),
                          date == "n_frags_final" ~ as.Date("2024-12-01")),
         nursery = "Lab",
         date_floor = floor_date(ymd(date), unit = "month"))

nursery_mon_bleaching_2024 <- nursery_mon %>%
  bind_rows(lab_frags, outplant_adjustments) %>%
  mutate_at(c('n_healthy', 'n_dead', 'n_bleached', 'n_pale', 'n_diseased', 'n_other'), ~replace_na(.,0))

# determine nursery-level pre-bleaching baseline numbers

status_nursery_pre_bleaching_2024 <- nursery_mon_bleaching_2024 %>%
  filter(date_floor == pre_bleaching_baseline_2024) %>%
  group_by(nursery) %>%
  summarize(n_tot_baseline_og = sum(n_healthy + n_other + n_pale)) %>% # "other" frags consolidated here for simplicity, should not be significant numbers of bleached in baseline
  left_join(nursery_adjustments %>% # incorporating frags that were removed/re-added for lab, DN, etc.
              group_by(nursery) %>%
              summarize(n_adj = sum(n_frags))) %>%
  mutate(n_adj = replace_na(n_adj, 0),
         n_tot_baseline = n_tot_baseline_og + n_adj)

# determining condition of fragments at post-bleaching monitoring dates

status_nursery_post_bleaching_2024 <- nursery_mon_bleaching_2024 %>%
  filter(date_floor == post_bleaching_2024 & 
           !(nursery == "York" & location %in% c("C1", "D1", "E1", "F1")) & # removing GBs because those would be double counted with deep nursery after moving it in Nov.
           !(nursery == "Cades" & location %in% c("A1", "B1"))) %>% # remove A1 and B1 from Cades as new ropes were added in Dec.
  select(nursery, n_healthy, n_bleached, n_pale, n_diseased, n_other) %>%
  group_by(nursery) %>%
  summarize(n_healthy = sum(n_healthy + n_other), # "other" frags consolidated here for simplicity
            n_pale = sum(n_pale),
            n_bleached = sum(n_bleached),
            n_diseased = sum(n_diseased)
            ) %>% 
  mutate(n_tot_alive = n_healthy + n_pale + n_bleached + n_diseased) %>%
  left_join(status_nursery_pre_bleaching_2024 %>%
              select(nursery, n_tot_baseline)) %>%
  mutate(n_dead_rel_baseline = n_tot_baseline - n_tot_alive) %>%
  select(nursery, n_tot_baseline, n_tot_alive, n_healthy, n_diseased, n_pale, n_bleached, n_dead_rel_baseline) %>%
  adorn_totals("row") %>%
  mutate(p_healthy = n_healthy/n_tot_alive*100, # relative to current total
         p_diseased = n_diseased/n_tot_alive*100, 
         p_pale = n_pale/n_tot_alive*100,
         p_bleached = n_bleached/n_tot_alive*100,
         p_healthy_rel_baseline = n_healthy/n_tot_baseline*100, # relative to baseline total
         p_diseased_rel_baseline = n_diseased/n_tot_baseline*100,
         p_pale_rel_baseline = n_pale/n_tot_baseline*100,
         p_bleached_rel_baseline = n_bleached/n_tot_baseline*100,
         p_dead_rel_baseline = (n_tot_baseline - n_tot_alive)/n_tot_baseline*100,
         p_survival_rel_baseline = n_tot_alive/n_tot_baseline*100
         ) %>%
  select(nursery, n_tot_baseline, n_tot_alive, n_healthy, n_diseased, n_pale, n_bleached, n_dead_rel_baseline, p_healthy, p_diseased, p_pale, p_bleached, p_healthy_rel_baseline, p_diseased_rel_baseline, p_pale_rel_baseline, p_bleached_rel_baseline, p_dead_rel_baseline, p_survival_rel_baseline)

# plots

ggplot(status_nursery_post_bleaching_2024 %>%
         select(nursery, healthy = p_healthy_rel_baseline, bleached = p_bleached_rel_baseline, pale = p_pale_rel_baseline, diseased = p_diseased_rel_baseline, dead = p_dead_rel_baseline) %>%
         pivot_longer(cols = healthy:dead, 
                      names_to = "condition",
                      values_to = "percent")  %>%
         mutate(nursery = fct_relevel(nursery, c('Total','York','TPB', 'DN', 'Cades')),
                condition = fct_relevel(str_to_sentence(condition), c("Dead", "Diseased", "Bleached", "Pale", "Healthy"))) %>%
         arrange(nursery),
       aes(x = nursery, y = percent, fill = condition)) +
  geom_col(color = "black") +
  scale_fill_manual(values = c("slategray", "coral", "azure1", "tan", "goldenrod")) +
  labs(y = "Percent of fragments", 
       x = "Nursery", fill = "Condition", 
       title = paste0("Fragment status in ",  
                      month(ymd(post_bleaching_2024), label = TRUE), " ", 
                      year(ymd(post_bleaching_2024)), " ",
                      "relative to ", 
                      month(ymd(pre_bleaching_baseline_2024), label = TRUE), " ", 
                      year(ymd(pre_bleaching_baseline_2024)))) +
  theme_bw() +
  theme(plot.title = element_text(hjust=0.5))

ggsave(here("restoration_monitoring", "figs", "bleaching_outcomes_2024_nursery.png"), width = 10, height = 5)
```

### Comparing 2024 outcomes to 2023

```{r}
# establish date references

pre_bleaching_baseline_2023 <- ymd("2023-09-01")
post_bleaching_2023 <- ymd("2023-12-01") 

# determine nursery-level pre-bleaching baseline numbers

status_total_pre_bleaching_2023 <- nursery_mon %>%
  filter(date_floor == pre_bleaching_baseline_2023) %>%
  summarize(n_tot_baseline = sum(n_healthy + n_other + n_pale))


# determining condition of fragments at post-bleaching monitoring dates

status_total_post_bleaching_2024 <- nursery_mon %>%
  filter(date_floor == post_bleaching_2023) %>%
  mutate(n_tot_alive = n_healthy + n_bleached + n_pale + n_other + n_diseased) %>%
  summarise(n_healthy = sum(n_healthy + n_other),
            n_pale = sum(n_pale),
            n_bleached = sum(n_bleached),
            n_diseased = sum(n_diseased),
            n_tot_alive = sum(n_tot_alive)) %>%
  bind_cols(status_total_pre_bleaching_2023) %>%
  mutate(n_dead_rel_baseline = n_tot_baseline - n_tot_alive) %>%
  select(n_tot_baseline, n_tot_alive, n_healthy, n_diseased, n_pale, n_bleached, n_dead_rel_baseline) %>%
  mutate(p_healthy = n_healthy/n_tot_alive*100, # relative to current total
         p_diseased = n_diseased/n_tot_alive*100, 
         p_pale = n_pale/n_tot_alive*100,
         p_bleached = n_bleached/n_tot_alive*100,
         p_healthy_rel_baseline = n_healthy/n_tot_baseline*100, # relative to baseline total
         p_diseased_rel_baseline = n_diseased/n_tot_baseline*100,
         p_pale_rel_baseline = n_pale/n_tot_baseline*100,
         p_bleached_rel_baseline = n_bleached/n_tot_baseline*100,
         p_dead_rel_baseline = (n_tot_baseline - n_tot_alive)/n_tot_baseline*100,
         p_survival_rel_baseline = n_tot_alive/n_tot_baseline*100
         ) %>%
  select(n_tot_baseline, n_tot_alive, n_healthy, n_diseased, n_pale, n_bleached, n_dead_rel_baseline, p_healthy, p_diseased, p_pale, p_bleached, p_healthy_rel_baseline, p_diseased_rel_baseline, p_pale_rel_baseline, p_bleached_rel_baseline, p_dead_rel_baseline, p_survival_rel_baseline)

bleaching_comp_yr <- status_total_post_bleaching_2024 %>%
  mutate(year = "2023") %>%
  bind_rows(status_nursery_post_bleaching_2024 %>%
              filter(nursery == "Total") %>%
              mutate(year = "2024")) %>%
  select(year, healthy = p_healthy_rel_baseline, bleached = p_bleached_rel_baseline, pale = p_pale_rel_baseline, diseased = p_diseased_rel_baseline, dead = p_dead_rel_baseline) %>%
  pivot_longer(cols = healthy:dead, 
                      names_to = "condition",
                      values_to = "percent")

ggplot(bleaching_comp_yr %>%
         mutate(condition = fct_relevel(str_to_sentence(condition), c("Dead", "Diseased", "Bleached", "Pale", "Healthy"))),
       aes(x = year, y = percent, fill = condition)) +
  geom_col(color = "black", position = "stack") +
  scale_fill_manual(values = c("slategray", "coral", "azure1", "tan", "goldenrod")) +
  labs(y = "Percent of fragments", 
       x = "", fill = "Condition", 
       title = "Fragment status post-bleaching") +
  theme_bw() +
  theme(plot.title = element_text(hjust=0.5))

  
```

### Comparing only genotypes present in all 4 nurseries over the summer of 2024

- non-normal distribution -> non-parametric tests
- significantly higher survival at Cade's than York, no other significant differences among groups

```{r}
# identify genotypes present at all 4 comparison nurseries
genotype_comparisons <- nursery_mon_bleaching_2024 %>%
  filter(date_floor == pre_bleaching_baseline_2024 &
           nursery %in% c("Cades", "DN", "York", "Lab")
         ) %>%
  select(genotype, nursery) %>%
  distinct() %>%
  group_by(genotype) %>%
  summarize(count = n()) %>%
  filter(count == 4) %>%
  select(genotype)
genotype_comparisons <- as.list(genotype_comparisons$genotype)

# determine genotype-level pre-bleaching baseline numbers at each nursery for genotypes selected for comparison
status_genotype_nursery_baseline <- nursery_mon_bleaching_2024 %>%
  filter(date_floor == pre_bleaching_baseline_2024 & 
           genotype %in% c(genotype_comparisons) &
           nursery %in% c("Cades", "DN", "York", "Lab")
         ) %>%
  group_by(genotype, nursery) %>%
  summarize(n_healthy = sum(n_healthy),
            n_diseased = sum(n_diseased),
            n_pale = sum(n_pale),
            n_bleached = sum(n_bleached),
            n_other = sum(n_other),
            n_tot_baseline_og = sum(n_healthy + n_other + n_pale)
            ) %>%
  left_join(nursery_adjustments %>% # incorporating frags that were removed/re-added for lab, DN, etc.
              group_by(nursery, genotype) %>%
              summarize(n_adj = sum(n_frags))) %>%
  mutate(n_adj = replace_na(n_adj, 0),
         n_tot_baseline = n_tot_baseline_og + n_adj) %>%
  # manually adjusting here two genotypes that do not make sense in terms of total before/after numbers
  mutate(n_tot_baseline = if_else(genotype == "APAL12" & nursery == "York", 17,
                                  if_else(genotype == "APAL52" & nursery == "York", 25,
                                          n_tot_baseline)))

# determine genotype-level post-bleaching numbers at each nursery for genotypes selected for comparison
status_genotype_nursery <- nursery_mon_bleaching_2024 %>%
  filter(date_floor == post_bleaching_2024 
           & genotype %in% c(genotype_comparisons)
           & nursery %in% c("Cades", "DN", "York", "Lab")
         ) %>%
  group_by(genotype, nursery) %>%
  summarize(n_healthy = sum(n_healthy),
            n_diseased = sum(n_diseased),
            n_pale = sum(n_pale),
            n_bleached = sum(n_bleached),
            n_other = sum(n_other),
            n_dead = sum(n_dead)
            ) %>% 
  right_join(status_genotype_nursery_baseline %>%
              select(genotype, nursery, n_tot_baseline)) %>%
  mutate_at(c('n_healthy', 'n_dead', 'n_bleached', 'n_pale', 'n_diseased', 'n_other'),
            ~replace_na(.,0)) %>%
  mutate(n_tot_alive = n_healthy + n_diseased + n_pale + n_bleached + n_other,
         n_dead_rel_baseline = n_tot_baseline - n_tot_alive) %>%
  select(genotype, nursery, n_tot_baseline, n_tot_alive, n_healthy, n_diseased, n_pale, n_bleached, n_dead_rel_baseline) %>%
  distinct() %>%
  mutate(p_healthy = n_healthy/n_tot_alive*100, # relative to current total
         p_diseased = n_diseased/n_tot_alive*100, 
         p_pale = n_pale/n_tot_alive*100,
         p_bleached = n_bleached/n_tot_alive*100,
         p_healthy_rel_baseline = n_healthy/n_tot_baseline*100, # relative to baseline total
         p_diseased_rel_baseline = n_diseased/n_tot_baseline*100,
         p_pale_rel_baseline = n_pale/n_tot_baseline*100,
         p_bleached_rel_baseline = n_bleached/n_tot_baseline*100,
         p_dead_rel_baseline = (n_tot_baseline-n_tot_alive)/n_tot_baseline*100,
         p_survival_rel_baseline = n_tot_alive/n_tot_baseline*100
         ) %>%
  select(genotype, nursery, n_tot_baseline, n_tot_alive, n_healthy, n_diseased, n_pale, n_bleached, n_dead_rel_baseline, p_healthy, p_diseased, p_pale, p_bleached, p_healthy_rel_baseline, p_diseased_rel_baseline, p_pale_rel_baseline, p_bleached_rel_baseline, p_dead_rel_baseline, p_survival_rel_baseline)


# exploratory plots

ggplot(status_genotype_nursery %>%
         group_by(nursery) %>%
         summarize(n_tot_alive = sum(n_tot_alive),
            n_tot_baseline = sum(n_tot_baseline)
            ) %>%
         mutate(p_survival_rel_baseline = n_tot_alive/n_tot_baseline*100
         ) %>%
         mutate(nursery = case_when(nursery == "DN" ~ "Deep",
                                    nursery == "TPB" ~ "Ten Lb.",
                                    .default = nursery)) %>%
         arrange(nursery),
       aes(x = nursery, y = p_survival_rel_baseline)) +
  geom_col(color = "black", fill = "goldenrod") +
  labs(y = "Percent survival", x = "Nursery", fill = "Condition", , title = "Performance of comparison genotypes July-Dec. 2024") +
  theme_bw() +
  theme(plot.title = element_text(hjust=0.5))
ggsave(here("restoration_monitoring", "figs", "bleaching_outcomes_2024_compgenotypes_nursery_survival.png"), width = 6, height = 5)

ggplot(status_genotype_nursery %>%
         group_by(nursery) %>%
         summarize(n_healthy = sum(n_healthy),
            n_diseased = sum(n_diseased),
            n_pale = sum(n_pale),
            n_bleached = sum(n_bleached),
            n_dead_rel_baseline = sum(n_dead_rel_baseline),
            n_tot_alive = sum(n_tot_alive),
            n_tot_baseline = sum(n_tot_baseline)
            ) %>%
         mutate(healthy = n_healthy/n_tot_baseline*100,
                diseased = n_diseased/n_tot_baseline*100,
                pale = n_pale/n_tot_baseline*100,
                bleached = n_bleached/n_tot_baseline*100,
                dead = (n_tot_baseline-n_tot_alive)/n_tot_baseline*100
         ) %>%
         mutate(nursery = case_when(nursery == "DN" ~ "Deep",
                                    nursery == "TPB" ~ "Ten Lb.",
                                    .default = nursery)) %>%
         pivot_longer(cols = healthy:dead, 
                      names_to = "condition", 
                      values_to = "p_frags") %>%
         filter(nursery != "Total") %>%
         mutate(condition = fct_relevel(str_to_sentence(condition), c("Dead", "Diseased", "Bleached", "Pale", "Healthy"))) %>%
         arrange(nursery),
       aes(x = nursery, y = p_frags, fill = condition)) +
  geom_col(color = "black") +
  scale_fill_manual(values = c("slategray", "coral", "azure1", "tan", "goldenrod")) +
  labs(y = "Percent of fragments", x = "Nursery", fill = "Condition", , title = "Performance of comparison genotypes July-Dec. 2024") +
  theme_bw() +
  theme(plot.title = element_text(hjust=0.5))
ggsave(here("restoration_monitoring", "figs", "bleaching_outcomes_2024_compgenotypes_nursery.png"), width = 6, height = 5)

my_comparisons <- list( c("Cades", "Lab"), c("York", "Lab"))
ggplot(status_genotype_nursery, 
       aes(x = nursery, y = p_survival_rel_baseline)) +
  geom_violin(fill = "cadetblue", color = "cadetblue", alpha = 0.2) +
  stat_summary(fun = "mean",
               geom = "crossbar", 
               width = 0.2,
               colour = "coral") +
  geom_beeswarm(color = "cadetblue", alpha = 0.8) +
  stat_compare_means(paired = TRUE, method = "wilcox.test", 
                     comparisons = my_comparisons, label.y = c(105, 110),
                     label = "p.signif",
                     ref.group = "0.5") +
  stat_compare_means(label.y = 120) +
  scale_y_continuous(breaks = c(0, 25, 50, 75, 100)) +
  labs(x = "Nursery", y = "Percent survival", title = "Survival of replicated genotypes across nurseries (July - Dec. 2024)") +
  theme_bw()
ggsave(here("restoration_monitoring", "figs", "bleaching_outcomes_2024_compgenotypes_nursery.png"), width = 6, height = 5)

ggplot(status_genotype_nursery,
       aes(x = nursery, y = p_survival_rel_baseline)) +
  geom_col(aes(fill = nursery), alpha = 0.8, color = "black") +
  scale_fill_manual(values = c("deepskyblue3", "cadetblue4", "goldenrod", "seagreen")) +
  coord_flip() +
  labs(x = "", y = "Percent survival", fill = "") +
  scale_y_continuous(breaks = c(0, 100)) +
  theme_minimal() +
  theme(legend.position = "bottom",
        axis.text.x = element_text(angle = 45, hjust = 1)) +
  facet_wrap(~genotype, ncol = 7)
ggsave(here("restoration_monitoring", "figs", "bleaching_outcomes_2024_compgenotypes.png"), width = 10, height = 5)

# summary statistics
genotype_nursery_stats <- status_genotype_nursery %>%
  group_by(nursery) %>%
  get_summary_stats(p_survival_rel_baseline, type = "mean_sd")
kable(genotype_nursery_stats)

status_genotype_nursery %>%
  group_by(nursery) %>%
  shapiro_test(p_survival_rel_baseline)
ggqqplot(status_genotype_nursery, "p_survival_rel_baseline", facet.by = "nursery")
# non normal distribution -> non-parametric tests -> 

friedman_genotype_nursery <- friedman.test(p_survival_rel_baseline ~ nursery | genotype, data = status_genotype_nursery)
friedman_genotype_nursery
# significant differences among groups -> now need to determine among which groups

# post-hoc analyses
pwc_genotype_nursery <- status_genotype_nursery %>%
  arrange(genotype) %>%
  wilcox_test(p_survival_rel_baseline ~ nursery, paired = TRUE, p.adjust.method = "bonferroni")
print(pwc_genotype_nursery)

dunn_genotype_nursery <- dunn.test(status_genotype_nursery$p_survival_rel_baseline, status_genotype_nursery$nursery, 
                          method = "bonferroni")  # Use Bonferroni correction
print(dunn_genotype_nursery)
```

## Graphing nursery heat stress over time

```{r}
therm_stress_time <- nursery_mon %>%
  select(date_floor, date_label, n_healthy, n_bleached, n_pale, n_diseased, n_dead) %>%
  mutate(n_tot_frags = n_healthy + n_bleached + n_pale + n_diseased + n_dead) %>%
  group_by(date_floor, date_label) %>%
  summarize(n_tot = sum(n_tot_frags),
            Healthy = sum(n_healthy)/n_tot*100,
            Bleached = sum(n_bleached)/n_tot*100,
            Pale = sum(n_pale)/n_tot*100,
            Diseased = sum(n_diseased)/n_tot*100,
            Dead = sum(n_dead)/n_tot*100
            ) %>%
  pivot_longer(cols = Healthy:Dead, names_to = "condition", values_to = "percent") %>%
  mutate(condition = fct_relevel(condition, c("Dead", "Diseased", "Bleached", "Pale", "Healthy")))

ggplot(therm_stress_time,
       aes(x = date_floor, y = percent, group = condition)) +
  geom_bar(aes(fill = condition),
               position = "stack", stat = 'identity', color = "black") +
  scale_fill_manual(values = c("slategray", "coral", "azure1", "tan", "goldenrod")) +
  scale_x_date(date_labels = "%Y-%b", date_breaks  = "1 month") +
  labs(y = "Percent of fragments", x = "Date", fill = "Condition") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

ggsave(here("restoration_monitoring", "figs", "frag_status_over_time.png"), width = 12, height = 5)
```













## Incomplete code/analyses

### Summarizing 2023 and 2024 bleaching by genotype

```{r, eval = FALSE}
# determine genotype-level pre-bleaching baseline numbers
status_genotype_baseline <- nursery_mon %>%
  filter(date_floor == pre_bleaching_baseline_2024) %>%
  group_by(genotype) %>%
  summarize(n_healthy = sum(n_healthy),
            n_diseased = sum(n_diseased),
            n_pale = sum(n_pale),
            n_bleached = sum(n_bleached),
            n_other = sum(n_other)
            ) %>% 
  left_join(frags_lab) %>%
  mutate(n_lab = replace_na(n_lab, 0),
         n_tot_baseline = n_healthy + n_diseased + n_pale + n_bleached + n_other)

# genotype-level snapshot at post-bleaching date, using baseline pre-bleaching data to calculate percent loss
status_genotype <- nursery_mon %>%
  filter(date_floor == post_bleaching_2024) %>% # most recent dates
  select(genotype, n_healthy, n_diseased, n_pale, n_bleached, n_other) %>%
  group_by(genotype) %>%
  summarize(n_healthy = sum(n_healthy),
            n_diseased = sum(n_diseased),
            n_pale = sum(n_pale),
            n_bleached = sum(n_bleached),
            n_other = sum(n_other)
            ) %>% 
  mutate(n_tot_alive = n_healthy + n_diseased + n_pale + n_bleached + n_other) %>%
  left_join(status_genotype_baseline %>%
              select(genotype, n_tot_baseline, n_lab)) %>%
  mutate(n_dead_rel_baseline = n_tot_baseline - n_tot_alive - n_lab) %>%
  select(genotype, n_tot_baseline, n_lab, n_tot_alive, n_healthy, n_diseased, n_pale, n_bleached, n_dead_rel_baseline) %>%
  distinct() %>%
  mutate(p_healthy = n_healthy/n_tot_alive*100, # relative to current total
         p_diseased = n_diseased/n_tot_alive*100, 
         p_pale = n_pale/n_tot_alive*100,
         p_bleached = n_bleached/n_tot_alive*100,
         p_healthy_rel_baseline = n_healthy/n_tot_baseline*100, # relative to baseline total
         p_diseased_rel_baseline = n_diseased/n_tot_baseline*100,
         p_pale_rel_baseline = n_pale/n_tot_baseline*100,
         p_bleached_rel_baseline = n_bleached/n_tot_baseline*100,
         p_dead_rel_baseline = (n_tot_baseline-n_tot_alive)/n_tot_baseline*100,
         p_survival_rel_baseline = n_tot_alive/n_tot_baseline*100
         ) %>%
  select(genotype, n_tot_baseline, n_lab, n_tot_alive, n_healthy, n_diseased, n_pale, n_bleached, n_dead_rel_baseline, p_healthy, p_diseased, p_pale, p_bleached, p_healthy_rel_baseline, p_diseased_rel_baseline, p_pale_rel_baseline, p_bleached_rel_baseline, p_dead_rel_baseline, p_survival_rel_baseline)
### need to figure out negative numbers/NAs here ^

ggplot(status_genotype %>%
         select(genotype, healthy = p_healthy_rel_baseline, bleached = p_bleached_rel_baseline, pale = p_pale_rel_baseline, diseased = p_diseased_rel_baseline, dead = p_dead_rel_baseline) %>%
         pivot_longer(cols = -genotype,
                 names_to = "condition",
                 values_to = "percent"),
       aes(x = genotype, y = percent, group = condition)) +
  geom_col(aes(fill = condition)) +
  scale_fill_manual(breaks = c('dead', 'diseased', 'bleached', 'pale', 'healthy'), values = c("slategray", "coral", "azure1", "tan", "goldenrod")) +
  labs(x = "Genotype", 
       y = "Percent of fragments", 
       fill = "Condition",
       title = paste0("Fragment status in ",  
                      month(ymd(post_bleaching_2024), label = TRUE), " ", 
                      year(ymd(post_bleaching_2024)), " ",
                      "relative to ", 
                      month(ymd(pre_bleaching_baseline_2024), label = TRUE), " ", 
                      year(ymd(pre_bleaching_baseline_2024)))) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  coord_flip()

ggsave(here("restoration_monitoring", "figs", "bleaching_outcomes_2024_genotype.png"), width = 6, height = 14)
```


```{r}
int_bleaching_2023 <- interval(ymd("2023-09-01"), ymd("2023-12-10")) # date range for investigation of 2023 bleaching

# calculate total loss across nurseries

thermtol_geno_date <- nursery_mon %>%
  filter(date == "2023-09-25") %>%
  group_by(genotype) %>%
  summarize(n_og = sum(n_healthy + n_dead + n_bleached + n_pale + n_diseased + n_other)) %>%
  right_join(nursery_mon %>%
               filter(date %within% int_bleaching_2023) %>%
               group_by(genotype, date) %>%
               summarize(n_healthy = sum(n_healthy), n_bleached = sum(n_bleached), n_pale = sum(n_pale),  n_dead = sum(n_dead), n_diseased = sum(n_diseased), n_other = sum(n_other))) %>%
  mutate(p_thermtolerant = n_healthy / n_og) 

thermtol_geno_2023 <- thermtol_geno_date %>%
  group_by(genotype) %>%
  summarize(min_thermtolerance = min(p_thermtolerant)) %>%
  left_join(thermtol_geno_date %>%
              filter(date == "2023-12-04") %>%
              mutate(survivorship = n_healthy / n_og) %>%
              select(genotype, survivorship)
            ) %>%
  mutate(survivorship = replace_na(survivorship, 0))

write.csv(thermtol_geno_2023, here("restoration_monitoring", "data_outputs", "thermtol_2023.csv"), row.names=FALSE)
```

### Disease susceptibility

- rate of infection: number of infected frags / total frags across all structures

```{r}
### need to account for multiple dates - or combine dates
disease <- nursery_mon %>%
  group_by(genotype, date) %>%
  summarize(n_healthy = sum(n_healthy), n_bleached = sum(n_bleached), n_pale = sum(n_pale),  n_dead = sum(n_dead), n_diseased = sum(n_diseased), n_other = sum(n_other)) %>%
  mutate(p_diseased = n_diseased / (n_diseased + n_healthy + n_bleached + n_pale + n_other)) %>% # not relative to dead fragments
  mutate(across(p_diseased, ~ replace(., is.nan(.), 0))) %>%
  group_by(genotype) %>%
  summarize(mean_infection = mean(p_diseased)) # need to check APAL14

write.csv(disease, here("restoration_monitoring", "data_outputs", "disease.csv"), row.names=FALSE)

#SCTLD graph for MRC talk
# remove ACER, APAL, APRO,

ggplot(disease %>% filter(!str_detect(genotype, "ACER|APRO|APAL")),
       aes(x = genotype, y = mean_infection*100)) +
  geom_col(fill = "darkslategray4") +
  ylim(0, 100) +
  labs(x = "Genotype", y = "Mean rate of infection (%)") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

ggsave(here("restoration_monitoring", "figs", "disease_rate_sctld.png"), width = 10, height = 6)
```

### Broodstock survival
```{r}
#Cades: A4
#TPB: A4

broodstock <- nursery_mon %>%
  filter((nursery == "Cades" & location == "A4") | (nursery == "TPB" & location == "A4")) %>%
  mutate(n_tot = n_healthy + n_diseased + n_bleached + n_pale) 

broodstock_survival <- broodstock %>%
  filter(date == "2024-05-20" | date == "2024-06-06") %>%
  select(nursery, genotype, n_tot_baseline = n_tot) %>%
  left_join(broodstock %>%
              filter(date == "2024-11-05" | date == "2024-11-11") %>%
              select(nursery, genotype, n_tot_24.11 = n_tot)
  ) %>%
  mutate(survival = n_tot_24.11/n_tot_baseline*100)

broodstock_survival_nursery <- broodstock_survival %>%
  group_by(nursery) %>%
  summarize(survival = mean(survival))
```












```{r, EVAL = F}
# archive...
# 
# 
# 
# # species codes
#          species = case_when(substr(genotype, 1, 4) == "ACER" ~ "A. cervicornis",
#                              substr(genotype, 1, 4) == "APRO" ~ "A. prolifera",
#                              substr(genotype, 1, 4) == "APAL" ~ "A. palmata",
#                              substr(genotype, 1, 4) == "PPOR" ~ "P. porites",
#                              substr(genotype, 1, 4) == "PDIV" ~ "P. divaricata",
#                              substr(genotype, 1, 4) == "OANN" ~ "O. annularis",
#                              substr(genotype, 1, 4) == "OFAV" ~ "O. faveolata",
#                              substr(genotype, 1, 4) == "OFRA" ~ "O. franksi",
#                              substr(genotype, 1, 4) == "PSTR" ~ "P. strigosa",
#                              substr(genotype, 1, 4) == "PCLI" ~ "P. clivosa",
#                              substr(genotype, 1, 4) == "CNAT" ~ "C. natans",
#                              substr(genotype, 1, 4) == "DLAB" ~ "D. labyrinthiformis",
#                              substr(genotype, 1, 4) == "DCYL" ~ "D. cylindrus",
#                              substr(genotype, 1, 4) == "MCAV" ~ "M. cavernosa"),
```







