---
title: "Nursery monitoring"
author: "Molly Wilson"
date: "2023-09-02"
output: html_document
---

```{r, message = F, warning = F, echo = F}
library(tidyverse)
library(here) 
library(readxl) 
library(janitor)
library(snakecase) # for adjusting capitalization of text within data (e.g., species names)
library(knitr) # for including tables
library(stringr)
library(lubridate)

knitr::opts_chunk$set(message = FALSE, warning = FALSE)

nursery_mon <- read_excel(here("restoration_monitoring", "data_raw", "Genotype tracking.xlsx"), sheet = "nursery monitoring") %>% 
  clean_names() %>%
  mutate(date = ymd(date),
         species = case_when(substr(genotype, 1, 4) == "ACER" ~ "A. cervicornis",
                             substr(genotype, 1, 4) == "APRO" ~ "A. prolifera",
                             substr(genotype, 1, 4) == "APAL" ~ "A. palmata",
                             substr(genotype, 1, 4) == "PPOR" ~ "P. porites",
                             substr(genotype, 1, 4) == "PDIV" ~ "P. divaricata",
                             substr(genotype, 1, 4) == "OANN" ~ "O. annularis",
                             substr(genotype, 1, 4) == "OFAV" ~ "O. faveolata",
                             substr(genotype, 1, 4) == "OFRA" ~ "O. franksi",
                             substr(genotype, 1, 4) == "PSTR" ~ "P. strigosa",
                             substr(genotype, 1, 4) == "PCLI" ~ "P. clivosa",
                             substr(genotype, 1, 4) == "CNAT" ~ "C. natans",
                             substr(genotype, 1, 4) == "DLAB" ~ "D. labyrinthiformis",
                             substr(genotype, 1, 4) == "DCYL" ~ "D. cylindrus",
                             substr(genotype, 1, 4) == "MCAV" ~ "M. cavernosa"),
         ) %>%
  filter(!is.na(date)) %>%
  mutate_at(c('n_healthy', 'n_dead', 'n_bleached', 'n_pale', 'n_diseased', 'n_other'), ~replace_na(.,0))
```

Determining baseline and current monitoring dates
```{r}
# determine baseline dates before 2024 bleaching started
start_mon_cades <- nursery_mon %>%
  filter(nursery == "Cades" & date < ymd("2024-09-01")) %>%
  filter(date == last(date))
start_mon_date_cades <- start_mon_cades$date

start_mon_nonsuch <- nursery_mon %>%
  filter(nursery == "York" & date < ymd("2024-09-01")) %>%
  filter(date == last(date))
start_mon_date_nonsuch <- start_mon_nonsuch$date

start_mon_deep <- nursery_mon %>%
  filter(nursery == "DN") %>%
  filter(date == first(date))
start_mon_date_deep <- start_mon_deep$date

# determine most recent monitoring dates for Cades and Nonsuch to get latest data
last_mon_cades <- nursery_mon %>%
  filter(nursery == "Cades") %>%
  filter(date == last(date))
last_mon_date_cades <- last_mon_cades$date

last_mon_nonsuch <- nursery_mon %>%
  filter(nursery == "York") %>%
  filter(date == last(date))
last_mon_date_nonsuch <- last_mon_nonsuch$date
```

Summarizing current fragment status by genotype
```{r}
# current nursery fragment numbers and status by genotype
therm_stress_baseline_genotype <- nursery_mon %>%
  filter((nursery == "Cades" & date == start_mon_cades$date) |
           (nursery %in% c("TPB", "York") & date == start_mon_nonsuch$date) |
           (nursery == "DN" & date == start_mon_deep$date)
         ) %>%
  group_by(genotype) %>%
  summarize(n_healthy = sum(n_healthy),
            n_diseased = sum(n_diseased),
            n_pale = sum(n_pale),
            n_bleached = sum(n_bleached),
            n_other = sum(n_other)
            ) %>% 
  mutate(n_tot_og = n_healthy + n_diseased + n_pale + n_bleached + n_other)

frag_status_genotype <- nursery_mon %>%
  filter(date %in% c(last_mon_cades$date, last_mon_nonsuch$date)) %>% # most recent dates
  select(genotype, n_healthy, n_diseased, n_pale, n_bleached, n_other) %>%
  group_by(genotype) %>%
  summarize(n_healthy = sum(n_healthy),
            n_diseased = sum(n_diseased),
            n_pale = sum(n_pale),
            n_bleached = sum(n_bleached),
            n_other = sum(n_other)
            ) %>% 
  mutate(n_tot_alive = n_healthy + n_diseased + n_pale + n_bleached + n_other) %>%
  left_join(therm_stress_baseline_genotype %>%
              select(genotype, n_tot_og)) %>%
  mutate(n_dead = n_tot_og-n_tot_alive) %>%
  select(genotype, n_tot_og, n_tot_alive, n_healthy, n_diseased, n_pale, n_bleached, n_dead) %>%
  distinct()

write.csv(frag_status_genotype, here("restoration_monitoring", "data_outputs", "frag_status.csv"), row.names=FALSE)
```

Summarizing current fragment status by nursery
```{r}
# determine nursery-level pre-bleaching baseline numbers
therm_stress_baseline_nursery <- nursery_mon %>%
  filter((nursery == "Cades" & date == start_mon_cades$date) |
           (nursery %in% c("TPB", "York") & date == start_mon_nonsuch$date) |
           (nursery == "DN" & date == start_mon_deep$date)
         ) %>%
  group_by(nursery) %>%
  summarize(n_healthy = sum(n_healthy),
            n_diseased = sum(n_diseased),
            n_pale = sum(n_pale),
            n_bleached = sum(n_bleached),
            n_other = sum(n_other),
            n_dead = sum(n_dead)
            ) %>% 
  mutate(n_tot_og = n_healthy + n_diseased + n_pale + n_bleached + n_dead + n_other)

# determining condition of fragments at most recent monitoring dates
therm_stress_summary_nursery <- nursery_mon %>%
  filter(date %in% c(last_mon_cades$date, last_mon_nonsuch$date)) %>% # most recent dates
  select(nursery, n_healthy, n_bleached, n_pale, n_diseased) %>% # "other" frags not included here for simplicity
  group_by(nursery) %>%
  summarize(n_healthy = sum(n_healthy),
            n_diseased = sum(n_diseased),
            n_pale = sum(n_pale),
            n_bleached = sum(n_bleached)
            ) %>% 
  mutate(n_tot_alive = n_healthy + n_diseased + n_pale + n_bleached) %>%
  left_join(therm_stress_baseline_nursery %>%
              select(nursery, n_tot_og)) %>%
  mutate(n_dead = n_tot_og-n_tot_alive) %>%
  select(nursery, n_tot_og, n_tot_alive, n_healthy, n_diseased, n_pale, n_bleached, n_dead)

therm_stress_summary <- therm_stress_summary_nursery %>%
  summarize(n_healthy = sum(n_healthy),
            n_diseased = sum(n_diseased),
            n_pale = sum(n_pale),
            n_bleached = sum(n_bleached),
            n_dead = sum(n_dead),
            n_tot_alive = sum(n_tot_alive),
            n_tot_og = sum(n_tot_og)
            ) %>%
  mutate(nursery = "Total") %>%
  select(nursery, n_tot_og, n_tot_alive, n_healthy, n_diseased, n_pale, n_bleached, n_dead) %>%
  rbind(therm_stress_summary_nursery) %>%
  mutate(p_healthy = n_healthy/n_tot_og*100,
         p_diseased = n_diseased/n_tot_og*100,
         p_pale = n_pale/n_tot_og*100,
         p_bleached = n_bleached/n_tot_og*100,
         p_dead = (n_tot_og-n_tot_alive)/n_tot_og*100,
         p_survival = n_tot_alive/n_tot_og*100
         ) %>%
  select(nursery, n_tot_og, n_tot_alive, n_healthy, n_diseased, n_pale, n_bleached, n_dead, p_healthy, p_diseased, p_pale, p_bleached, p_dead, p_survival)

# plots

ggplot(therm_stress_summary %>%
         select(nursery, n_healthy, n_diseased, n_pale, n_bleached, n_dead) %>%
         pivot_longer(cols = n_healthy:n_dead, 
                      names_to = "condition", 
                      names_prefix = 'n_',
                      values_to = "n_frags") %>%
         mutate(nursery = fct_relevel(nursery, c('Total','York','TPB', 'DN', 'Cades'))) %>%
         arrange(nursery),
       aes(x = nursery, y = n_frags, fill = condition)) +
  geom_col(color = "black") +
  #scale_x_discrete(labels = c('Total','York','TPB', 'DN', 'Cades')) +
  scale_fill_manual(breaks = c('dead', 'diseased', 'bleached', 'pale', 'healthy'), values = c("slategray", "coral", "azure1", "tan", "goldenrod")) +
  labs(y = "Number of fragments", x = "Nursery", fill = "Condition") +
  theme_bw()

ggplot(therm_stress_summary %>%
         select(nursery, p_healthy, p_diseased, p_pale, p_bleached, p_dead) %>%
         pivot_longer(cols = p_healthy:p_dead, 
                      names_to = "condition", 
                      names_prefix = 'p_',
                      values_to = "percent")  %>%
         mutate(nursery = fct_relevel(nursery, c('Total','York','TPB', 'DN', 'Cades'))) %>%
         arrange(nursery),
       aes(x = nursery, y = percent, fill = condition)) +
  geom_col(color = "black") +
  scale_fill_manual(breaks = c('dead', 'diseased', 'bleached', 'pale', 'healthy'), values = c("slategray", "coral", "azure1", "tan", "goldenrod")) +
  labs(y = "Percent of fragments", x = "Nursery", fill = "Condition") +
  theme_bw()
```

2023 bleaching event
- survivorship
- % therm tolerant (no bleaching or paling) out of living fragments

Notes for future improvement:
- currently doesn't distinguish between heat-related death vs. disease, etc.
- more checks and balances to ensure that no data collected doesn't equate to 0 corals
- how to account for times when structures are fluffled, etc. (i.e., number of fragments increase)
- disease monitoring can be ongoing - and relative to actual number of fragments, while bleaching monitoring would only occur in the fall (when structures are not refragged etc?)

Graphing nursery heat stress over time
```{r}
# think I need to redo this to have n_tot be from the beginning of a time interval?

n_tot <- nursery_mon %>%
  filter(date %in% c(last_mon_cades$date, last_mon_nonsuch$date)) %>% # most recent dates
  # filter(date == last_mon_nonsuch$date) %>% # Nonsuch only
  pivot_longer(cols = n_healthy:n_diseased, names_to = "condition", values_to = "n_frags") %>%
  summarize(n_tot = sum(n_frags))

therm_stress_time <- nursery_mon %>%
  select(date, n_healthy, n_bleached, n_pale, n_diseased) %>%
  group_by(date) %>%
  summarize(Healthy = sum(n_healthy)/n_tot$n_tot*100,
            Bleached = sum(n_bleached)/n_tot$n_tot*100,
            Pale = sum(n_pale)/n_tot$n_tot*100) %>%
  mutate(Dead = 100 - (Healthy + Bleached + Pale)) %>%
  pivot_longer(cols = Healthy:Dead, names_to = "condition", values_to = "percent")
factor(therm_stress_time$condition, levels = c("Dead", "Bleached", "Pale", "Healthy"))

ggplot(therm_stress_time,
       aes(x = as.factor(date), y = percent, group = condition)) +
  geom_bar(aes(fill = condition), position = "stack", stat = 'identity', color = "black") +
  scale_fill_manual(breaks = c('Dead', 'Bleached', 'Pale', 'Healthy'), values = c("slategray", "azure1", "tan", "goldenrod")) +
  labs(y = "Percent of fragments", x = "Date", fill = "Condition") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

Genotype-specific results of 2023 bleaching
```{r}
int_bleaching_2023 <- interval(ymd("2023-09-01"), ymd("2023-12-10")) # date range for investigation of 2023 bleaching

thermtol_geno_date <- nursery_mon %>%
  filter(date == "2023-09-25") %>%
  group_by(genotype) %>%
  summarize(n_og = sum(n_healthy + n_dead + n_bleached + n_pale + n_diseased + n_other)) %>%
  right_join(nursery_mon %>%
               filter(date %within% int_bleaching_2023) %>%
               group_by(genotype, date) %>%
               summarize(n_healthy = sum(n_healthy), n_bleached = sum(n_bleached), n_pale = sum(n_pale),  n_dead = sum(n_dead), n_diseased = sum(n_diseased), n_other = sum(n_other))) %>%
  mutate(p_thermtolerant = n_healthy / n_og) 

thermtol_geno <- thermtol_geno_date %>%
  group_by(genotype) %>%
  summarize(min_thermtolerance = min(p_thermtolerant)) %>%
  left_join(thermtol_geno_date %>%
              filter(date == "2023-12-04") %>%
              mutate(survivorship = n_healthy / n_og) %>%
              select(genotype, survivorship)
            ) %>%
  mutate(survivorship = replace_na(survivorship, 0))

write.csv(thermtol_geno, here("restoration_monitoring", "data_outputs", "thermtol_2023.csv"), row.names=FALSE)
```

Disease susceptibility
- rate of infection: number of infected frags / total frags across all structures

```{r}
disease <- nursery_mon %>%
  group_by(genotype, date) %>%
  summarize(n_healthy = sum(n_healthy), n_bleached = sum(n_bleached), n_pale = sum(n_pale),  n_dead = sum(n_dead), n_diseased = sum(n_diseased), n_other = sum(n_other)) %>%
  mutate(p_diseased = n_diseased / (n_diseased + n_healthy + n_bleached + n_pale + n_other)) %>% # not relative to dead fragments
  mutate(across(p_diseased, ~ replace(., is.nan(.), 0))) %>%
  group_by(genotype) %>%
  summarize(mean_infection = mean(p_diseased)) # need to check APAL14

write.csv(disease, here("restoration_monitoring", "data_outputs", "disease.csv"), row.names=FALSE)
  
```

Number of current fragments
```{r}
# added in Q3 2024

collection <- read_excel(here("restoration_monitoring", "data_raw", "Genotype tracking.xlsx"), sheet = "collection") %>%
  clean_names() %>%
  mutate(collection_date = ymd(date)) %>%
  rename(collection_site = site,
         collection_lat = lat,
         collection_lon = lon)

q3 <- interval(ymd("2024-01-01"), ymd("2024-03-31"))
q3_additions <- collection %>%
  filter(date %within% q3) %>%
  select(genotype) %>%
  left_join(nursery_mon %>% 
              filter(date == "2024-03-25") %>%
              mutate(n_tot = n_healthy + n_diseased + n_bleached + n_pale) %>%
              select(genotype, n_tot)) # last monitoring date in Q3

q3_genotypes <- q3_additions %>% 
  select(genotype) %>%
  distinct()

q3_frags <- q3_additions %>%
  filter(!is.na(n_tot)) %>%
  summarize(n_frags = sum(n_tot))
```
Broodstock survival
```{r}
#Cades: A4
#TPB: A4

broodstock <- nursery_mon %>%
  filter((nursery == "Cades" & location == "A4") | (nursery == "TPB" & location == "A4")) %>%
  mutate(n_tot = n_healthy + n_diseased + n_bleached + n_pale) %>%
  group_by(genotype) %>%
  summarize()
```














Old code..... need to organize

```{r, EVAL = F}
# archive...
therm_stress_geno <- nursery_mon %>%
  filter(date == "2023-09-25" & !location %in% c("C1", "D1", "E1", "F1")) %>% # excluding genebanks
  mutate(n_og = n_healthy + n_dead + n_bleached + n_pale + n_diseased) %>%
  select(nursery, location, genotype, n_og) %>%
  left_join(nursery_mon %>%
              filter(date == "2023-12-04") %>% # comparing to December 2023 data collection
              select(nursery, location, genotype, n_healthy, n_bleached, n_pale, n_diseased), 
            by = c("nursery", "location", "genotype")) %>%
  mutate_at(c('n_healthy', 'n_bleached', 'n_pale', 'n_diseased'), ~replace_na(.,0)) %>%
  mutate(p_healthy = n_healthy/n_og*100,
         p_bleached = n_bleached/n_og*100,
         p_pale = n_pale/n_og*100,
         p_diseased = n_diseased/n_og*100,
         p_dead = (n_og - (n_healthy + n_bleached + n_pale + n_diseased))/n_og*100) %>%
  group_by(genotype) %>%
  summarise(dead_percent = mean(p_dead), # took averages across structures for each genotype
            dead_se = sd(p_dead)/sqrt(n()),
            healthy_percent = mean(p_healthy),
            healthy_se = sd(p_healthy)/sqrt(n()),
            bleached_percent = mean(p_bleached),
            bleached_se = sd(p_bleached)/sqrt(n()),
            pale_percent = mean(p_pale),
            pale_se = sd(p_pale)/sqrt(n()),
            diseased_percent = mean(p_diseased),
            diseased_se = sd(p_diseased)/sqrt(n()),
            thermstressed_percent = mean(p_bleached + p_pale),
            thermstressed_se = sd(p_bleached + p_pale)/sqrt(n()))
write.csv(therm_stress_geno, here("restoration_monitoring", "data_outputs", "bleaching_outcomes.csv"))

therm_stress_long_temp <- therm_stress_geno %>%
  pivot_longer(cols = -genotype,
                 names_to = c('variable', '.value'),
                 names_sep = '_') %>%
  rename(condition = variable) 

therm_stress_long <- therm_stress_long_temp %>%
  left_join(therm_stress_long_temp %>%
              filter(condition == "dead") %>%
              arrange(percent) %>%
              mutate(order = row_number()) %>%
              select(genotype, order))

ggplot(therm_stress_long %>%
         filter(condition %in% c("healthy", "dead", "thermstressed", "diseased")) %>%
         mutate(genotype = fct_reorder(genotype, order)),
       aes(x = genotype, y = percent, group = condition)) +
  geom_col(aes(fill = condition)) +
  scale_fill_manual(values = c("darkseagreen", "coral", "goldenrod", "azure1")) +
  labs(x = "Genotype", y = "Percent of fragments", fill = "Condition") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
ggsave(here("restoration_monitoring", "figs", "bleaching_outcomes_23.12.04.png"), width = 12, height = 5)
```







