---
title: "Outplant monitoring"
author: "Molly Wilson"
date: "2023-02-19"
output: 
  html_document:
    code_folding: hide
---

# {.tabset}


```{r, message = F, warning = F, echo = F}
library(tidyverse)
library(here) 
library(readxl) 
library(janitor)
library(snakecase) # for adjusting capitalization of text within data (e.g., species names)
library(knitr) # for including tables
library(stringr)
library(lubridate)

knitr::opts_chunk$set(message = FALSE, warning = FALSE)
```

### Outplant statistics
```{r}
outplants <- read_excel(here("restoration_monitoring", "data_raw", "Genotype tracking.xlsx"), sheet = "outplants") %>% 
  clean_names() %>%
  mutate(genotype = substr(id, 1, 6),
         species = case_when(substr(id, 1, 4) == "ACER" ~ "A. cervicornis",
                             substr(id, 1, 4) == "APRO" ~ "A. prolifera",
                             substr(id, 1, 4) == "APAL" ~ "A. palmata"),
         species_com = case_when(substr(id, 1, 4) == "ACER" ~ "Staghorn",
                             substr(id, 1, 4) == "APRO" ~ "Fused staghorn",
                             substr(id, 1, 4) == "APAL" ~ "Elkhorn"),
         date = ymd(date)) %>%
  filter(!is.na(date)) %>%
  select(date, site, species, genotype, id, n_frags, n_colonies)

frags_tot <- sum(outplants$n_frags)
colonies_tot <- sum(outplants$n_colonies)
genotypes_tot <- count(outplants %>%
  distinct(genotype))
genotypes_spp_site <- outplants %>%
  group_by(species, site) %>%
  distinct(genotype) %>%
  count()
counts_spp <- outplants %>%
  group_by(species) %>%
  summarise(frags = sum(n_frags),
            p_frags = sum(n_frags)/frags_tot*100,
            colonies = sum(n_colonies),
            p_colonies = sum(n_colonies)/colonies_tot*100)

outplant_activity <- outplants %>%
  mutate(month = month.abb[month(date, label = TRUE)],
         year = year(date)) %>%
  group_by(date) %>%
  summarize(colonies_planted = sum(n_colonies))

# add graph for this
```
**Total number of fragments planted:** `r frags_tot`

**Total number of colonies planted:** `r colonies_tot`

**Total number of genotypes planted:** `r genotypes_tot`

**Number of genotypes planted by species and site:**
`r kable(genotypes_spp_site)`

**Number of fragments and colonies planted by species:**
`r kable(counts_spp)`

### Survivorship 
```{r}
outplant_monitoring <- read_excel(here("restoration_monitoring", "data_raw", "genotype tracking.xlsx"), sheet = "outplant monitoring") %>% 
  clean_names() %>%
  filter(!is.na(colonies_live)) %>% # filtering out any genotypes with missing data
  mutate(colonies_planted = as.numeric(colonies_planted),
         date_monitored = ymd(date_monitored),
         # if year/month is needed: format_ISO8601(ymd(date_monitored), precision = "ym"),
         date_planted = ymd(date_planted))
test <- outplant_monitoring %>% filter(colonies_live > colonies_planted) # check data for extra colonies

survivorship_id <- outplant_monitoring %>%
  filter(date_monitored == max(outplant_monitoring$date_monitored)) %>%
  pivot_longer(
     cols = starts_with("live_c"),
     names_to = "colony",
     names_prefix = "live_c",
     values_to = "survivorship",
     values_drop_na = TRUE #  drops all colonies planted as indv (APALs and some APROs)
  ) %>%
  group_by(date_monitored, id) %>%
  summarise(survivorship_col = mean(survivorship)) %>%
  # pivot_wider(names_from = date_monitored, 
  #             values_from = survivorship_col) %>% # was using this when trying to look across different dates
  right_join(outplant_monitoring %>%
               filter(date_monitored == max(outplant_monitoring$date_monitored)) %>%
               select(date_monitored, site, id, date_planted, colonies_planted, colonies_live), by = c("date_monitored", "id")) %>%
  mutate(age = (date_planted %--% date_monitored) / months(1),
         survivorship_col = if_else(is.na(survivorship_col), 100, survivorship_col),
         survivorship_pop = survivorship_col * colonies_live / colonies_planted) %>%
  relocate(survivorship_col, .before = survivorship_pop) %>%
  mutate(genotype = substr(id, 1, 6),
         species = case_when(substr(id, 1, 4) == "ACER" ~ "A. cervicornis",
                             substr(id, 1, 4) == "APRO" ~ "A. prolifera",
                             substr(id, 1, 4) == "APAL" ~ "A. palmata"),
         species_com = case_when(substr(id, 1, 4) == "ACER" ~ "Staghorn",
                             substr(id, 1, 4) == "APRO" ~ "Fused staghorn",
                             substr(id, 1, 4) == "APAL" ~ "Elkhorn"),
         site = case_when(site == "York Island" ~ "York",
                             site == "Ten Pound Bay" ~ "Ten Pound",
                             site == "Ricketts" ~ "Ricketts"))

# overall survivorship across all outplants
survivorship <- mean(survivorship_id$survivorship_pop)

ggplot(survivorship_id, aes(x = survivorship_pop)) +
  geom_density(fill = "steelblue", alpha = .4) +
  xlim(0, 100) +
  labs(x = "Survivorship of outplanted population", y = "Density") +
  geom_vline(xintercept = survivorship, linetype="dotted") +
  annotate("text", x = 70, y = .026, label = paste("Mean = ", round(survivorship, digits = 1), "%")) +
  theme_bw()
ggsave(here("restoration_monitoring", "figs", "survivorship_dens.png"))
```


### Investigating survivorship by...

## ~ species
```{r}
survivorship_spp <- survivorship_id %>%
  group_by(species, species_com) %>%
  summarise(survivorship = mean(survivorship_pop),
            se = sd(survivorship_pop)/sqrt(n()),
            replicates = n())

kable(survivorship_spp)

ggplot(survivorship_spp, aes(x = species_com, y = survivorship)) +
  geom_col(color = "black", fill = "paleturquoise3", alpha = 0.7) +
  geom_errorbar(aes(ymin = survivorship - se, ymax = survivorship + se), width = .2,
                 position = position_dodge(.9)) +
  lims(y = c(0,100)) +
  labs(y = "Outplant survivorship (%)", x = "") +
  theme_bw() +
  theme(axis.text = element_text(size = 14), axis.title = element_text(size = 14))
ggsave(here("restoration_monitoring", "figs", "survivorship_spp_col.png"), width = 6.5, height = 5)

ggplot(survivorship_id, aes(x = survivorship_pop))+
  geom_density(alpha = 0.7, fill = "paleturquoise3") +
  facet_grid(rows = vars(species)) +
  labs(x = "Survivorship of outplanted population", y = "Density") +
  xlim(0, 100) +
  scale_y_continuous(breaks = c(0.03, 0.06)) +
  geom_vline(data = survivorship_spp, 
             aes(xintercept = survivorship), linetype = "dotted") +
  theme_bw() +
  theme(strip.text = element_text(face = "italic"))
ggsave(here("restoration_monitoring", "figs", "survivorship_spp_dens.png"))
```

## ~ outplant age
```{r}
ggplot(survivorship_id, aes(x = age, y = survivorship_pop)) +
  geom_point() +
  facet_wrap(. ~ species) +
  labs(x = "Age (months)", y = "Survivorship of population") +
  theme_bw()
```

Not really worth looking at now, as we have a small range in terms of outplant age (and a big gap between July and October)

## ~ genotype
```{r}
survivorship_genotype <- survivorship_id %>%
  group_by(species, genotype) %>%
  summarise(survivorship_mean = mean(survivorship_pop),
            se = sd(survivorship_pop)/sqrt(n()),
            replicates = n()) %>% 
  arrange(-survivorship_mean)

kable(survivorship_genotype)
```

## ~ site

```{r}
survivorship_spp_site <- survivorship_id %>%
  group_by(species, site) %>%
  summarise(survivorship = mean(survivorship_pop),
            se = sd(survivorship_pop)/sqrt(n()),
            replicates = n())

ggplot(survivorship_spp_site, aes(x = site, y = survivorship)) +
  geom_col(color = "black", fill = "steelblue", alpha = 0.4) +
  facet_grid(. ~ species) +
  geom_errorbar(aes(ymin = survivorship - se, ymax = survivorship + se), width = .2,
                 position = position_dodge(.9)) +
  lims(y = c(0,100)) +
  labs(y = "Survivorship of all outplants", x = "") +
  theme_bw() +
  theme(strip.text = element_text(face = "italic"))

# now with only genotypes that have been replicated at all sites
survivorship_genotype_site <- survivorship_id %>%
  group_by(species, site, genotype) %>%
  summarise(survivorship = mean(survivorship_pop),
            se = sd(survivorship_pop)/sqrt(n()),
            replicates = n()) %>% 
  arrange(-survivorship)

genotypes_all_sites <- survivorship_genotype_site %>%
  group_by(genotype) %>%
  summarise(n_sites = n()) %>%
  filter(n_sites == 3) # genotypes with outplants at all 3 sites

survivorship_genotype_site <- survivorship_genotype_site %>%
  filter(genotype %in% genotypes_all_sites$genotype)

survivorship_spp_site_reps <- survivorship_id %>%
  filter(genotype %in% genotypes_all_sites$genotype) %>%
  group_by(species, site) %>%
  summarise(survivorship = mean(survivorship_pop),
            se = sd(survivorship_pop)/sqrt(n()),
            replicates = n())

ggplot(survivorship_spp_site_reps, aes(x = site, y = survivorship)) +
  geom_col(color = "black", fill = "steelblue", alpha = 0.4) +
  facet_grid(. ~ species) +
  geom_errorbar(aes(ymin = survivorship - se, ymax = survivorship + se), width = .2,
                 position = position_dodge(.9)) +
  lims(y = c(0,100)) +
  labs(y = "Survivorship of replicated genotypes", x = "") +
  theme_bw() +
  theme(strip.text = element_text(face = "italic"))

kable(survivorship_spp_site_reps)
```


#### Comparing genotypes with replicates at all 3 sites
```{r}
survivorship_genotype_site <- survivorship_id %>%
  group_by(species, site, genotype) %>%
  summarise(survivorship = mean(survivorship_pop),
            se = sd(survivorship_pop)/sqrt(n()),
            replicates = n()) %>% 
  arrange(-survivorship)

genotypes_all_sites <- survivorship_genotype_site %>%
  group_by(genotype) %>%
  summarise(n_sites = n()) %>%
  filter(n_sites == 3) # genotypes with outplants at all 3 sites

survivorship_genotype_site <- survivorship_genotype_site %>%
  filter(genotype %in% genotypes_all_sites$genotype)

ggplot(survivorship_genotype_site, aes(x = genotype, y = survivorship, color = site)) +
  geom_point(alpha = 0.9) +
  scale_color_manual(values = c("coral", "aquamarine3", "midnightblue")) +
  lims(y = c(0,100)) +
  labs(y = "Survivorship", x = "", color = "Restoration site") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```


```{r, eval = F}
ggplot(survivorship_genotype_site %>% filter(species == "A. palmata"), aes(x = genotype, y = survivorship)) +
  geom_col(aes(fill = site), position = "dodge", alpha = 0.9, color = "black") +
  scale_fill_manual(values = c("steelblue1", "aquamarine3", "gray")) +
  labs(x = "Genotype", y = "Percent survival", fill = "Outplant site") +
  lims(y = c(0,100)) +
  theme_bw()

ggplot(survivorship_genotype_site %>% filter(species == "A. prolifera"), aes(x = genotype, y = survivorship)) +
  geom_col(aes(fill = site), position = "dodge", alpha = 0.9, color = "black") +
  scale_fill_manual(values = c("steelblue1", "aquamarine3", "gray")) +
  labs(x = "Genotype", y = "Percent survival", fill = "Outplant site") +
  lims(y = c(0,100)) +
  theme_bw()

ggplot(survivorship_genotype_site %>% filter(species == "A. cervicornis"), aes(x = genotype, y = survivorship)) +
  geom_col(aes(fill = site), position = "dodge", alpha = 0.9, color = "black") +
  scale_fill_manual(values = c("steelblue1", "aquamarine3", "gray")) +
  labs(x = "Genotype", y = "Percent survival", fill = "Outplant site") +
  lims(y = c(0,100)) +
  theme_bw()

```





