---
title: "Species interaction study"
author: "Molly Wilson"
date: "2025-06-10"
output: 
  html_document:
    code_folding: hide
---

```{r, message = F, warning = F, echo = F}
library(tidyverse)
library(here) 
library(readxl) 
library(janitor)
library(stringr)
library(lubridate)

knitr::opts_chunk$set(message = FALSE, warning = FALSE)
```

Steps:

1a. Calculate changes in buoyant weights
1b. Test for differences in growth rates across groups
2a. Calculate survival rates
2b. Test for differences in survival rates

# Import data

```{r}
weights <- read_excel(here("spp_interaction_exp", "data_raw", "Species interaction data.xlsx"), 
                  sheet = "Buoyant weights") %>%
  clean_names() %>%
  mutate(type = case_when(table %in% c("C1", "C2") ~ "Control",
                          table %in% c("T1", "T2") ~ "Treatment"))
```


# Growth rates
1. Remove NAs
2. Calculate adjusted masses
3. Get mean adjusted mass

```{r}

growth <- weights %>%
  filter(date != as.Date("2025-06-06")) %>% # remove intermediate sampling date to look at long term growth
  filter(is.na(partial_death) | partial_death != "with") %>% # remove measurements that include dead tissue weight
  select("id", "date", starts_with("buoy_wght_"), "density_seawater_kg_m3") %>%
  na.omit() %>%
  pivot_longer(
    cols = starts_with("buoy_wght_"),
    names_to = "replicate",
    values_to = "buoy_wght_g"
  ) %>%
  mutate(adj_wght_g = buoy_wght_g/(1-density_seawater_kg_m3/1000/2.94)) %>%
  group_by(id, date) %>%
  summarize(adj_wght_g = mean(adj_wght_g)) %>%
  group_by(id) %>%
  mutate(growth_rate = (adj_wght_g - lag(adj_wght_g)) / lag(adj_wght_g)) %>%
  ungroup() %>%
  na.omit() %>%
  mutate(treatment = case_when(substr(id, 1, 1) == "C" ~ "Control",
                          substr(id, 1, 1) == "T" ~ "Treatment"),
         table = substr(id, 1, 2),
         genotype = substr(id, 4, 5)
         ) 

ggplot(growth, aes(x = treatment, y = growth_rate, fill = genotype)) +
  geom_violin(position = position_dodge(width = 0.8), trim = FALSE) +
  geom_boxplot(width = 0.1, position = position_dodge(width = 0.8), outlier.shape = NA, color = "black", alpha = 0.3) +
  labs(
    x = "Treatment",
    y = "Growth rate",
    fill = "Genotype"
  ) +
  theme_classic(base_size = 14)

ggplot(growth, aes(x = treatment, y = growth_rate, fill = table)) +
  geom_violin(position = position_dodge(width = 0.8), trim = FALSE) +
  geom_boxplot(width = 0.1, position = position_dodge(width = 0.8), outlier.shape = NA, color = "black", alpha = 0.3) +
  labs(
    x = "Treatment",
    y = "Growth rate",
    fill = "Table"
  ) +
  theme_classic(base_size = 14)

sample_sizes <- growth %>%
  group_by(treatment, table, genotype) %>%
  summarize(n = n())

model_growth1 <- lm(growth_rate ~ treatment + table + genotype, data = growth)
summary(model_growth1)
model_growth2 <- lm(growth_rate ~ treatment + genotype, data = growth)
summary(model_growth2)

# check normality?
```

# Survival rates
1. Create df with condition (alive/dead) per individual
2. Determine predictors of condition (alive or dead)

```{r}
# do I have intermediate death data somewhere? Check distinction between dead vs missing
survival <- weights %>%
  group_by(id) %>%
  filter(date == max(date)) %>%
  ungroup() %>%
  select(id, condition) %>%
  filter(condition != "missing original") %>%
  mutate(treatment = case_when(substr(id, 1, 1) == "C" ~ "Control",
                          substr(id, 1, 1) == "T" ~ "Treatment"),
         table = substr(id, 1, 2),
         genotype = substr(id, 4, 5)
         ) %>%
  mutate(condition = factor(condition, levels = c("alive", "dead"))) # for model

model_survival <- glm(condition ~ treatment + table + genotype, 
             data = survival, 
             family = binomial)

summary(model_survival) # no statistically significant differences
```

