---
title: "Shark tagging"
author: "Molly Wilson"
date: "2025-11-06"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(readxl)
library(here)
library(purrr)
library(lubridate)
library(janitor)
library(sf)
library(ggplot2)
library(ggspatial)
library(gganimate)
library(patchwork)
# library(png)
# library(jpeg)
# library(grid)
# library(cowplot)
# library(gridExtra)
```

*CONVERT ALL DATES TO PLAIN TEXT WITHIN EXCEL FIRST TO AVOID FORMATTING ISSUES WITH INCONSISTENT DATA SOURCES*

# Import + format data

## Tagging data
```{r}
tagging <- read_excel(here("tagging", "data_raw", "Shark tagging data.xlsx"), sheet = "Tagging") %>% 
  clean_names() %>%
  filter(!str_detect(as.character(date), "[A-Za-z]") & sat_ptt != "NA") %>% # remove any placeholder dates
  mutate(date = as.Date(as.numeric(date), origin = "1899-12-30"),
         length_sex = paste0(tl, "cm-", sex)) %>%
  select(pttid = sat_ptt, date, lat = latitude, lon = longitude, sex, species, length_tot = tl, length_sex)

tagging <- read_excel(here("tagging", "data_raw", "Shark tagging data.xlsx"), sheet = "Tagging") %>% 
  clean_names() %>%
  filter(!str_detect(as.character(date), "[A-Za-z]") & sat_ptt != "NA") %>% 
  mutate(
    date = as.Date(as.numeric(date), origin = "1899-12-30"),
    tl_m  = tl / 100,
    tl_ft = tl / 30.48,
    sex_full = recode(sex,
                      "M" = "Male",
                      "F" = "Female"),
    length_sex = paste0(
      round(tl_m, 2), "m / ",
      round(tl_ft, 1), "ft ",
      sex_full)
  ) %>%
  select(
    pttid = sat_ptt, date,lat = latitude, lon = longitude, sex, species, length_tot = tl, length_sex)

```

## Raw ARGOS data

- each shark has individual sheet
- each sheet has multiple data types with different header rows bound together
- function below splits data types within each sheet and then binds sheets together into one df

*make sure all dates are in "date" ("YYYY-MM-DD") format within excel*
*make sure all sheets have DTPNG/DTTEMP/DTDPT between data chunks within excel*

```{r}
file_path <- here("tagging", "data_raw", "Argos_compiled.xlsx")
sheets <- excel_sheets(file_path)

process_sheet_split <- function(sheet_name) {
  raw <- read_excel(file_path, sheet = sheet_name, col_names = FALSE)
  
  type_rows <- which(raw[[1]] %in% c("DTPNG", "DTTEMP", "DTDPT"))
  header_rows <- which(raw[[1]] == "PTTID")
  
  if(length(type_rows) == 0 | length(header_rows) == 0) return(NULL)
  
  extract_block <- function(i) {
    type_row <- type_rows[i]
    header_row <- header_rows[i]
    end_row <- if(i < length(type_rows)) type_rows[i+1]-1 else nrow(raw)
    
    header <- as.character(unlist(raw[header_row, ]))
    data <- raw[(header_row+1):end_row, ]
    
    # Clean header
    header[is.na(header) | header == ""] <- paste0("col_", seq_len(sum(is.na(header) | header == "")))
    names(data) <- header
    data <- janitor::clean_names(data)
    
    # Remove empty rows
    data <- data %>% filter(!if_all(everything(), is.na))
    
    # Convert PTTID to numeric (extract numbers only)
    if("pttid" %in% names(data)) {
      data <- data %>% 
        mutate(pttid = str_extract(pttid, "\\d+")) %>%  # extract numeric part
        mutate(pttid = as.numeric(pttid))               # convert to numeric
    }
    
    list(type = raw[[1]][type_row], data = data)
  }
  
  blocks <- map(seq_along(type_rows), extract_block)
  
  # Split into named list by type
  split_blocks <- blocks %>% 
    set_names(map_chr(blocks, "type")) %>% 
    map("data")
  
  split_blocks
}

all_blocks <- map(sheets, process_sheet_split)

# create blocks by data type

pttid_list <- map_dfr(all_blocks, ~ .x[["DTPNG"]]) %>%
  pull(pttid) %>%
  unique()

pings_argos <- map_dfr(all_blocks, ~ .x[["DTPNG"]]) %>%
  mutate(
    date = case_when(
      # Excel numeric dates (including "45838" stored as character)
      !is.na(suppressWarnings(as.numeric(date_detected))) ~
        as.Date(as.numeric(date_detected), origin = "1899-12-30"),

      # Everything else: let lubridate figure it out
      TRUE ~
        as.Date(parse_date_time(
          date_detected,
          orders = c("ymd", "mdy", "dmy")
        ))
    )
  ) %>%
  select(pttid, date, lat = latitude_n, lon = longitude_w) %>%
  mutate(category = "Ping") %>%
  rbind(tagging %>%
          filter(pttid %in% pttid_list) %>%
          select(pttid, date, lat, lon) %>%
          mutate(category = "Deployment")) %>%
  left_join(tagging %>%
              select(pttid, sex, species, length_tot, length_sex),
            by = c("pttid"))

temps <- map_dfr(all_blocks, ~ .x[["DTTEMP"]]) %>%
  select(pttid, date, sst = sea_surface_temp_c, adt = at_depth_temp_c) %>%
  mutate(date = case_when( # dealing with multiple excel date types
    !is.na(suppressWarnings(as.numeric(date))) ~
      as.Date(as.numeric(date), origin = "1899-12-30"),
    TRUE ~
      as.Date(parse_date_time(
        date,
        orders = c("ymd", "mdy", "dmy")))),
    across(c(sst, adt), as.numeric))

depth <- map_dfr(all_blocks, ~ .x[["DTDPT"]]) %>%
  select(pttid, date, contains("depth")) %>%
  mutate(date = case_when( # dealing with multiple excel date types
    !is.na(suppressWarnings(as.numeric(date))) ~
      as.Date(as.numeric(date), origin = "1899-12-30"),
    TRUE ~
      as.Date(parse_date_time(
        date,
        orders = c("ymd", "mdy", "dmy")))),
    across(contains("depth"), as.numeric))

shark_data <- left_join(temps, depth, by = c("pttid", "date")) %>%
  mutate(pttid = as.character(pttid)) %>%
  left_join(pings_argos %>% select(pttid, date, lat, lon), by = c("pttid", "date")) %>%
  left_join(tagging %>%
              select(pttid, sex, species, length_tot, length_sex), 
            by = c("pttid"))
```
## Raw Swimway data

```{r}
pings_swimway <- read_excel(here("tagging", "data_raw", "Sattag data OpSwimaway 2023.xlsx"), sheet = "Tag data formatted") %>% 
  clean_names() %>%
  fill(pttid, sex, species) %>%
  mutate(date = dmy(date)) %>%
  left_join(tagging %>% select(pttid, length_tot, length_sex), by = "pttid") %>%
  select(pttid, date, lat = latitude, lon = longitude, category, sex, species, length_tot, length_sex)
```


## Integrate data sources

```{r}
pings <- rbind(pings_argos, pings_swimway) %>%
  mutate(species = as.character(species)) %>% # for filtering in maps
  arrange(date)
```


# Mapping - ESRI

*Projection = 3857 (must reproject dataset)*

## Prep

```{r}
# link ESRI layers
esri_ocean <- paste0('https://services.arcgisonline.com/arcgis/rest/services/','Ocean/World_Ocean_Base/MapServer/tile/${z}/${y}/${x}.jpeg')
esri_world_imagery <- "https://services.arcgisonline.com/arcgis/rest/services/World_Imagery/MapServer/tile/${z}/${y}/${x}.jpeg"

# function to prep data and bounding boxes based on dif data subsets
prepare_shark_df <- function(pings, pttids = NULL, buffer_prop = 0.5, crs_out = 3857) {
  
  # filter PTTIDs if provided
  data_sub <- if (!is.null(pttids)) {
    pings %>% filter(pttid %in% pttids)
  } else {
    pings
  }
  
  # convert to numeric
  data_sub$lon <- as.numeric(data_sub$lon)
  data_sub$lat <- as.numeric(data_sub$lat)
  
  # convert to sf and project
  sf_proj <- st_as_sf(data_sub, coords = c("lon", "lat"), crs = 4326) %>%
    st_transform(crs_out)
  
  # extract x, y
  df_proj <- sf_proj %>%
    mutate(
      x = st_coordinates(.)[,1],
      y = st_coordinates(.)[,2]
    ) %>%
    st_set_geometry(NULL)
  
  # calculate buffered bbox from subset
  expand_range <- function(x, buffer = 0.05) {
    r <- range(x, na.rm = TRUE)
    r + diff(r) * buffer * c(-1, 1)
  }
  
  lon_lim <- expand_range(data_sub$lon, buffer_prop)
  lat_lim <- expand_range(data_sub$lat, buffer_prop)
  
  bb_buffered <- st_as_sfc(st_bbox(c(
    xmin = lon_lim[1], xmax = lon_lim[2],
    ymin = lat_lim[1], ymax = lat_lim[2]
  ), crs = 4326))
  
  bb_3857 <- st_bbox(st_transform(bb_buffered, crs_out))
  xlim_3857 <- bb_3857[c("xmin", "xmax")]
  ylim_3857 <- bb_3857[c("ymin", "ymax")]
  
  list(
    df_proj = df_proj,
    xlim = xlim_3857,
    ylim = ylim_3857
  )
}

# color palette for tracks (add more if needed, or combine with linetypes?)
shark_palette <- c(
  "coral1",
  "goldenrod1",
  "seagreen1",
  "turquoise2",
  "magenta",
  "azure1",
  "orange",
  "aquamarine",
  "darkolivegreen1",
  "hotpink",
  "slategray1",
  "orangered"
)

# function for plotting shark tracks
p_shark <- function(pings,
                    pttids_subset = NULL,
                    species_filter = NULL,
                    buffer_prop = 0.5,
                    target_ratio = NULL,
                    start_date = NULL,
                    show_legend = TRUE,
                    animate = FALSE) {
  
  # ---- filter data ----
  data_sub <- pings
  
  if(!is.null(species_filter)) {
    data_sub <- data_sub %>% filter(species == species_filter)
  }
  
  if(!is.null(pttids_subset)) {
    data_sub <- data_sub %>% filter(pttid %in% pttids_subset)
  }
  
  if(!is.null(start_date)) {
    data_sub <- data_sub %>% filter(date >= as.Date(start_date))
  }
  
  # ---- projection + bbox ----
  prep <- prepare_shark_df(data_sub, buffer_prop = buffer_prop)
  df_proj <- prep$df_proj
  
  # ---- adjust bbox for target ratio if requested ----
  xlim_adj <- prep$xlim
  ylim_adj <- prep$ylim
  
  if(!is.null(target_ratio)){
    x_rng <- diff(prep$xlim)
    y_rng <- diff(prep$ylim)
    
    if((x_rng / y_rng) < target_ratio){
      # make x wider
      x_mid <- mean(prep$xlim)
      x_half <- (target_ratio * y_rng) / 2
      xlim_adj <- x_mid + c(-x_half, x_half)
    } else {
      # make y taller
      y_mid <- mean(prep$ylim)
      y_half <- x_rng / target_ratio / 2
      ylim_adj <- y_mid + c(-y_half, y_half)
    }
  }
  
  # ---- base map ----
  p <- ggplot() +
    annotation_map_tile(type = esri_world_imagery,
                        progress = "none",
                        zoomin = 0) +
    coord_sf(crs = st_crs(3857),
             xlim = xlim_adj,   # USE adjusted limits
             ylim = ylim_adj,   # USE adjusted limits
             expand = FALSE) +
    theme_bw(base_size = 18) +
    theme(
      panel.background = element_rect(fill = "steelblue"),
      panel.grid = element_blank(),
      legend.position = ifelse(animate, "none", "right"),
      legend.key = element_blank()
    ) +
    geom_path(
      data = df_proj,
      aes(x = x, y = y, group = pttid, color = pttid),
      linewidth = 1,
      alpha = 0.7,
      linetype = "11"
    ) +
    geom_point(
      data = df_proj,
      aes(x = x, y = y, group = pttid, color = pttid, shape = category),
      size = ifelse(animate, 2, 1),
      alpha = 0.7
    ) +
    scale_shape_manual(
      name = "Type",
      values = c(
        "Deployment" = 8,
        "Ping" = 16)
    ) +
    labs(x = NULL, y = NULL)
  
  # ---- automatic palette ----
  pttid_levels <- sort(unique(df_proj$pttid))
  colors <- shark_palette[seq_along(pttid_levels)] # Trim palette to needed length
  colors <- setNames(colors, pttid_levels) # Name colors so they match PTTIDs
  p <- p + scale_color_manual(values = colors,
                              name = "PTTID")
  
  # safety check
  if(length(pttid_levels) > length(shark_palette)) {
    stop("Not enough colors in shark_palette! Add more to shark_palette vector.")
  }
  
  # ---- turn animation ON if requested ----
  if(animate) {
    p <- p +
      labs(title = 'Tiger shark movement: {format(frame_along, "%B %d, %Y")}') +
      transition_reveal(along = date) +
      theme(
        plot.title = element_text(
          hjust = 0,
          size = 26,
          face = "bold"
        )
      )
  }
  
  # ---- turn legend ON/OFF if requested ----
  if (!show_legend) {
    p <- p + theme(legend.position = "none")
  }
  
  return(p)
}
```

## Plot

### GEF

```{r}
gef_pttids <- c("283833", "260593", "26354", "36353")

# static plot
p_static_gef <- p_shark(pings,
                        pttids_subset = gef_pttids)
ggsave(here("tagging", "figs", "shark_track_gef.png"), plot = p_static_gef, width = 10, height = 10)

# animated plots
p_anim_gef <- p_shark(
  pings,
  pttids_subset = gef_pttids,
  animate = TRUE
)

anim_save(
  here("tagging", "figs", "shark_track_gef.gif"),
  animate(
    p_anim_gef,
    nframes = 150,
    fps = 10,
    width = 600,
    height = 800,
    renderer = gifski_renderer()
  )
)

animate( # add library(av) if encountering issues
  p_anim_gef,
  nframes = 150,
  fps = 10,
  width = 600,
  height = 800,
  renderer = av_renderer(
    here("tagging", "figs", "shark_track_gef.mp4")
  )
)
```

### ISRA - all sharks by species
```{r}
# G. cuvier
p_static_tigers <- p_shark(pings, 
                           species_filter  = "Galeocerdo cuvier",
                           target_ratio = 1.5)
ggsave(here("tagging", "figs", "shark_track_gcuvier.png"), plot = p_static_tigers, width = 10, height = 8)

# C. falciformis
p_static_silkies <- p_shark(pings, 
                            species_filter = "Carcharhinus falciformis",
                            target_ratio = 1.5)
ggsave(here("tagging", "figs", "shark_track_cfalciformis.png"), plot = p_static_silkies, width = 10, height = 8)
```

### ISRA - individual sharks

```{r}
# for loop for each pttid
all_pttids <- unique(pings$pttid)
for(ptt in all_pttids) {
  
  # Generate static map
  p <- p_shark(pings, 
               pttids_subset = ptt,
               target_ratio = 1,
               buffer_prop = .75)
  
  # Save as PNG
  ggsave(
    filename = here("tagging", "figs", "indv_pttids", paste0("shark_", ptt, ".png")),
    plot = p,
    width = 8,
    height = 10,
    dpi = 300
  )
}
```


# Mapping - bathy

## Prep
```{r}
library(terra)

# bathymetry raster (user-defined GeoTIFF): Sentinel2 2022 ATG-specific from TNC: https://caribbeanscienceatlas.tnc.org/maps/28371a902f414fef830cc2e6ff9b3d8f

# set lims for crop to save processing time
lon_lim <- c(-62.22, -61.6)
lat_lim <- c(16.9, 17.3)

# create bbox in same CRS as raster (EPSG:4326)
bbox_4326 <- st_as_sfc(st_bbox(c(
  xmin = lon_lim[1],
  xmax = lon_lim[2],
  ymin = lat_lim[1],
  ymax = lat_lim[2]
), crs = 4326))

# load, crop, and convert to df
sentinel_df <- rast(here("mapping", "rasters", "sentinel_atg", "atg_bath_Sentinel2_2022_clip.tif")) %>%
  crop(bbox_4326) %>%
  as.data.frame(xy = TRUE) %>%
  setNames(c("x", "y", "depth"))

# downsample for faster rendering if needed
sentinel_df_small <- sentinel_df %>% sample_frac(0.5)
```


## Plot
```{r}
# static background (rendered once)
p_base <- ggplot() +
  geom_raster(data = sentinel_df, aes(x = x, y = y, fill = depth)) +
  scale_fill_gradient2(low = "tan", mid = "lightblue", high = "steelblue4", midpoint = 0, name = "Depth (m)") +
  coord_fixed(xlim = c(-62.15, -61.6), ylim = c(16.9, 17.3)) +
  theme_bw() +
  theme(panel.background = element_rect(fill = "steelblue4"),
        panel.grid = element_blank())

# animate overlay
p_anim <- p_base +
  geom_path(data = shark_df_3857, aes(x = x, y = y, group = pttid, color = pttid),
            size = 1, alpha = 0.7, linetype = 2) +
  geom_point(data = shark_df_3857, aes(x = x, y = y, group = pttid, color = pttid),
             size = 2) +
  scale_color_manual(values = c("coral2", "gold")) +
  labs(x = NULL, y = NULL, color = "Shark ID",
       title = "Shark movement over time: {frame_along}") +
  transition_reveal(along = date)

gganimate::animate(p_anim, nframes = 150, fps = 10)
anim_save(here("mapping", "figs", "shark_track_bathy.gif"), animation = anim)
```


# Depth / Temp

## Temperature

```{r}
p_temp_mult <- ggplot(shark_data, aes(x = date, group = pttid)) +
  geom_line(data = shark_data %>%
              pivot_longer(cols = c(sst, adt),
                           names_to = "type", values_to = "temp") %>%
              mutate(type = recode(type,
                                   sst = "Sea surface temp.",
                                   adt = "At-depth temp.")),
            aes(y = temp, color = type, group = interaction(pttid, type)),
            linewidth = 0.4) +
  scale_color_manual(values = c("Sea surface temp." = "coral",
                                "At-depth temp." = "goldenrod")) +
  theme_bw() +
  theme(legend.position = "top") +
  labs(color = NULL, y = "Temperature (C)", x = "")
ggsave(here("tagging", "figs", "shark_temp_mult.png"), plot = p_temp_mult, width = 6, height = 6)

p_temp_panel <- p_temp +
  facet_wrap(~ pttid, ncol = 2)
ggsave(here("tagging", "figs", "shark_temp_panel.png"), plot = p_temp_panel, width = 6, height = 6)
```

## Depth

```{r}
# still plot - mult
p_depth_mult <- ggplot(shark_data, aes(x = date)) +
  geom_ribbon(aes(ymin = min_depth_m, ymax = max_depth_m, group = pttid),
              fill = "gray", alpha = 0.4) +
  geom_line(data = shark_data %>%
              pivot_longer(cols = c(min_depth_m, mean_depth_m, max_depth_m),
                           names_to = "type", values_to = "depth") %>%
              mutate(type = recode(type,
                                   min_depth_m = "Min. depth",
                                   mean_depth_m = "Mean depth",
                                   max_depth_m = "Max. depth")),
            aes(y = depth, color = type, group = interaction(pttid, type)),
            linewidth = 0.8) +
  scale_y_reverse() +
  scale_color_manual(values = c("Min. depth" = "lightblue",
                                "Mean depth" = "gray50",
                                "Max. depth" = "steelblue")) +
  theme_bw() +
  theme(legend.position = "top") +
  labs(color = "", y = "Depth (m)", x = "")
ggsave(here("tagging", "figs", "shark_depth_mult.png"), plot = p_depth_mult, width = 6, height = 6)

# still plot - indv
p_depth_panel <- p_depth_mult +
  facet_wrap(~ pttid, ncol = 2)
ggsave(here("tagging", "figs", "shark_depth_panel.png"), plot = p_depth_panel, width = 8, height = 6)

# animated plots - turn off ribbon?
p_anim <- p_depth_panel +
  facet_wrap(~ pttid) +
  theme_bw(base_size = 18) +
  theme(legend.position = "top") +
  transition_reveal(date)
anim <- gganimate::animate(p_anim, nframes = 100, fps = 10, width = 800, height = 600)
anim_save(here("tagging", "figs", "shark_depth_indv.gif"), animation = anim)
```

## Panels

```{r}
# Plot panels
p_overlay <- p_temp / p_depth
ggsave(here("tagging", "figs", "depth_temp_OG.png"), plot = p_overlay, width = 6, height = 6)
```























# Archive

## Basic plots

*Projection = 4326 (same as dataset)*



```{r}
### Prep
# load Antigua shapefile (switch to Caribbean shapefile eventually)
atg <- st_read(here("mapping", "shapefiles", "atg_adm_2019_shp", "atg_admbnda_adm1_2019.shp")) %>%
  st_union() %>%
  st_sf()

# convert data to sf with og 4326 projection
shark_sf_4326 <- st_as_sf(pings, coords = c("lon", "lat"), crs = 4326)

# extract point data
shark_df_4326 <- shark_sf_4326 %>%
  mutate(x = st_coordinates(.)[,1],
         y = st_coordinates(.)[,2]) %>%
  st_set_geometry(NULL)

# standardize coords (ANU)
# lat_lim <- c(16.8, 17.3)
# lon_lim <- c(-62.15, -61.6)

# generating coordinate buffer
expand_range <- function(x, buffer = 0.3) { # buffer as proportion of data range
  r <- range(x, na.rm = TRUE)
  r + diff(r) * buffer * c(-1, 1)
}
```



```{r}
### Plot

#### Static
# plot pings - static
buffer_prop = .5
lon_lim <- expand_range(shark_df_4326$x, buffer_prop)
lat_lim <- expand_range(shark_df_4326$y, buffer_prop)

ggplot() +
  geom_sf(data = atg, fill = "lightblue", color = "lightblue") +  # islands
  geom_path(data = shark_df_4326, aes(x = x, y = y, group = pttid, color = pttid),
            size = .5, alpha = 0.5, linetype = 2) +
  geom_point(data = shark_df_4326, aes(x = x, y = y, group = pttid, color = pttid),
             size = 1) +
  # scale_color_manual(values = c("coral2", "gold")) +
  # xlim(c(lon_min, lon_max)) + ylim(c(lat_min, lat_max)) +
  coord_sf(xlim = lon_lim, ylim = lat_lim, expand = FALSE) +
  labs(x = NULL, y = NULL, color = "ID", title = "Shark movement over time") +
  theme_bw() +
  theme(panel.background = element_rect(fill = "steelblue"),
        panel.grid = element_blank())
#ggsave(here("tagging", "figs", "pings_static.png"))
```



```{r}
#### Animation

##### Compiled sharks
# animated plot - mult sharks
p_anim <- ggplot() +
  geom_sf(data = atg, fill = "lightblue", color = "lightblue") +  # islands
  geom_path(data = shark_df_4326, aes(x = x, y = y, group = pttid, color = pttid),
            size = 1, alpha = 0.7, linetype = 2) +
  geom_point(data = shark_df_4326, aes(x = x, y = y, group = pttid, color = pttid),
             size = 2) +
  # xlim(c(lon_min, lon_max)) + ylim(c(lat_min, lat_max)) +
  coord_sf(xlim = lon_lim, ylim = lat_lim, expand = FALSE) +
  labs(x = NULL, y = NULL, color = "ID", title = "Shark movement over time: {frame_along}") +
  theme_bw() +
  theme(panel.background = element_rect(fill = "steelblue"),
        panel.grid = element_blank()) +
  transition_reveal(date)

anim <- gganimate::animate(p_anim, nframes = 150, fps = 10, width = 600, height = 400, renderer = gifski_renderer())
anim_save(here("tagging", "figs", "shark_track_mult.gif"), animation = anim)
```



```{r}
##### Indvidual sharks
animate_indv <- function(
  pttid,
  shark_df,
  atg,
  buffer_prop = 0.5,
  out_dir = "tagging/figs"
) {

  # subset data ONCE
  df_i <- shark_df %>% 
    filter(pttid == !!pttid)

  # compute per-shark limits
  lon_lim <- expand_range(df_i$x, buffer_prop)
  lat_lim <- expand_range(df_i$y, buffer_prop)

  p_anim <- ggplot() +
    geom_sf(data = atg, fill = "lightblue", color = "lightblue") +
    geom_path(
      data = df_i,
      aes(x = x, y = y),
      color = "coral2", size = 1, alpha = 0.7, linetype = 2
    ) +
    geom_point(
      data = df_i,
      aes(x = x, y = y),
      color = "coral2", size = 2
    ) +
    coord_sf(
      xlim = lon_lim,
      ylim = lat_lim,
      expand = FALSE
    ) +
    labs(
      x = NULL, y = NULL,
      title = glue::glue("Shark {pttid} movement over time: {{frame_along}}")
    ) +
    theme_bw(base_size = 18) +
    theme(
      panel.background = element_rect(fill = "steelblue"),
      panel.grid = element_blank(),
      aspect.ratio = 1 # force square regardless of lims
    ) +
    transition_reveal(date)

  anim <- gganimate::animate(
    p_anim,
    nframes = 150, fps = 10,
    width = 1000, height = 750,
    renderer = gifski_renderer()
  )

  gif_path <- here::here(
    out_dir,
    paste0("shark_track_", pttid, ".gif")
  )

  anim_save(gif_path, animation = anim)

  message("✅ Saved: ", gif_path)
}

unique_ids <- unique(shark_df_4326$pttid)
purrr::walk(unique_ids, animate_indv, shark_df = shark_df_4326, atg = atg)
```
```{r}
animate_indv <- function(
  pttid,
  shark_df,
  atg,
  buffer_prop = 0.1,
  out_dir = "tagging/figs"
) {

  # subset shark data
  df_i <- shark_df %>% filter(pttid == !!pttid)

  if(nrow(df_i) == 0) {
    warning("No data for pttid: ", pttid)
    return(NULL)
  }

  # helper to expand range
  expand_range <- function(x, buffer = 0.1) {
    r <- range(x, na.rm = TRUE)
    r + diff(r) * buffer * c(-1, 1)
  }

  # project islands and shark points to UTM (zone 20N)
  atg_proj <- st_transform(atg, 32620)
  df_i_sf <- st_as_sf(df_i, coords = c("x", "y"), crs = 4326) %>%
    st_transform(32620)

  # compute per-shark limits in meters
  coords <- st_coordinates(df_i_sf)
  x_lim <- expand_range(coords[,1], buffer_prop)
  y_lim <- expand_range(coords[,2], buffer_prop)

  # build animation
  p_anim <- ggplot() +
    geom_sf(data = atg_proj, fill = "lightblue", color = "lightblue") +
    geom_path(data = as.data.frame(coords), aes(X, Y),
              color = "coral2", size = 1, alpha = 0.7, linetype = 2) +
    geom_point(data = as.data.frame(coords), aes(X, Y),
               color = "coral2", size = 2) +
    coord_sf(xlim = x_lim, ylim = y_lim, expand = FALSE) +
    labs(
      x = NULL, y = NULL,
      title = glue::glue("Shark {pttid} movement over time: {{frame_along}}")
    ) +
    theme_bw(base_size = 18) +
    theme(
      panel.background = element_rect(fill = "steelblue"),
      panel.grid = element_blank()
    ) +
    transition_reveal(df_i$date)

  # animate and save square GIF
  anim <- gganimate::animate(
    p_anim,
    nframes = 150, fps = 10,
    width = 800, height = 800,  # square
    renderer = gifski_renderer()
  )

  gif_path <- here::here(
    out_dir,
    paste0("shark_track_", pttid, ".gif")
  )

  anim_save(gif_path, animation = anim)
  message("✅ Saved: ", gif_path)
}

unique_ids <- unique(shark_df_4326$pttid)
purrr::walk(unique_ids, animate_indv, shark_df = shark_df_4326, atg = atg)

```

## Old ESRIs

```{r}
# plot formula
p_static <- function(pings, pttids_subset = NULL, species_subset = NULL, buffer_prop = 0.5) {
  
  # Filter by species if provided
  data_sub <- pings
  if(!is.null(species_subset)) {
    data_sub <- data_sub %>% filter(species == species_subset)
  }
  
  # Filter by PTTIDs if provided
  if(!is.null(pttids_subset)) {
    data_sub <- data_sub %>% filter(pttid %in% pttids_subset)
  }
  
  # Prepare subset + bbox
  prep <- prepare_shark_df(data_sub, buffer_prop = buffer_prop)
  df_proj <- prep$df_proj
  
  # Build base plot with bbox from subset
  p_base <- ggplot() +
    annotation_map_tile(type = esri_world_imagery, progress = "none", zoomin = 0) +
    coord_sf(crs = st_crs(3857), xlim = prep$xlim, ylim = prep$ylim, expand = FALSE) +
    theme_bw(base_size = 18) +
    theme(
      panel.background = element_rect(fill = "steelblue"),
      panel.grid = element_blank(),
      legend.position = "right",
      legend.key = element_blank()
    )
  
  # Add paths + points
  p <- p_base +
    geom_path(
      data = df_proj,
      aes(x = x, y = y, group = pttid, color = pttid),
      linewidth = 1, alpha = 0.7, linetype = "11"  # denser dashes
    ) +
    geom_point(
      data = df_proj,
      aes(x = x, y = y, group = pttid, color = pttid, shape = category),
      size = 1
    ) +
    scale_shape_manual(
      name = "Type",
      values = c(1, 16)  # 1 = hollow, 16 = solid
    )
  
  # Optional: if exactly 4 PTTIDs, assign custom colors
  pttid_levels <- unique(df_proj$pttid)
  if(length(pttid_levels) == 4) {
    colors <- c("coral1", "goldenrod1", "seagreen1", "turquoise2")
    colors <- setNames(colors, pttid_levels)
    p <- p + scale_color_manual(name = "Shark PTTID", values = colors)
  }
  
  # No title
  p <- p + labs(x = NULL, y = NULL)
  
  return(p)
}

# GEF PTTIDs only
p_static_gef <- p_static(pings, pttids_subset = c("283833", "260593", "26354", "36353")) +
  scale_color_manual(
    name = "Shark PTTID",
    values = c("coral1", "goldenrod1", "seagreen1", "turquoise2")
  )
ggsave(here("tagging", "figs", "shark_track_gef.png"), plot = p_static_gef, width = 10, height = 10)

# All PTTIDs
p_static_all <- p_static(pings)
ggsave(here("tagging", "figs", "shark_track_all.png"), plot = p_static_all, width = 10, height = 8)

# G. cuvier
p_static_tigers <- p_static(pings, species_subset = "Galeocerdo cuvier")
ggsave(here("tagging", "figs", "shark_track_gcuvier.png"), plot = p_static_tigers, width = 10, height = 8)

# C. falciformis
p_static_silkies <- p_static(pings, species_subset = "Carcharhinus falciformis")
ggsave(here("tagging", "figs", "shark_track_cfalciformis.png"), plot = p_static_silkies, width = 10, height = 8)
```

```{r, eval = FALSE}
# base / background (rendered once)
p_base <- ggplot() +
  annotation_map_tile(type = esri_world_imagery, progress = "none", zoomin = 0) +
  coord_sf(crs = st_crs(3857), xlim = xlim_3857, ylim = ylim_3857, expand = FALSE) +
  theme_bw(base_size = 18) +
  theme(
    panel.background = element_rect(fill = "steelblue"),
    panel.grid = element_blank(),
    legend.position = "right",
    legend.key = element_blank(),
    plot.title = element_text(hjust = 0.5, size = 16))
```

```{r, eval= FALSE}
# animate overlay (rendered repeatedly)
shark_key <- tagging %>%
  select(pttid, length_sex) %>%
  distinct()

shark_df_plot <- shark_df_3857 %>%
  left_join(shark_key, by = "pttid") %>%
  filter(date > as.Date("2025-01-01"))

p_anim <- p_base +
  geom_path(data = shark_df_plot, 
            aes(x = x, y = y, group = pttid, color = pttid),
            linewidth = 1, alpha = 0.7, linetype = 2) +
  geom_point(data = shark_df_plot,
             aes(x = x, y = y, group = pttid, color = pttid),
             size = 2) +
  scale_color_manual(
    name = "Total length & Sex",
    values = c("coral1", "goldenrod1", "seagreen1", "turquoise2"),
    labels = setNames(shark_key$length_sex, shark_key$pttid)
  ) +
  labs(x = NULL, y = NULL, color = "Total length & Sex",
       title = 'Tiger shark movement: {format(frame_along, "%B %d, %Y")}') +
  transition_reveal(along = date) +
  theme(legend.position = "none",
        plot.title = element_text(hjust = 0, size = 27, face = "bold"))

# render + save
# anim <- animate(
#   p_anim,
#   nframes = 150, fps = 10,
#   width = 600, height = 800,
#   renderer = av_renderer("animation.mp4")
# )
# anim_save(here("tagging", "figs", "shark_track_mult_esri.gif"), animation = anim)

animate(
  p_anim,
  nframes = 150,
  fps = 10,
  width = 600,
  height = 800,
  renderer = av_renderer(here::here("tagging", "figs", "shark_track_mult_esri.mp4"))
)
```


