---
title: "Water quality testing"
author: "Molly Wilson"
date: "2026-01-01"
output: 
  html_document:
    toc: true
    code_folding: hide
    number_sections: false
    theme: cerulean
    highlight: haddock
---
# To-dos:

- automatically incorporate 0-detection months for enterococcus data

# Set up
```{r, message = F, warning = F, echo = F}
library(tidyverse)
library(here) 
library(readxl) 
library(janitor)
library(scales)
library(lubridate)
library(ggpubr)
library(sf)
library(smplot2)
library(fuzzyjoin)

knitr::opts_chunk$set(message = FALSE, warning = FALSE)
```

## Importing, cleaning, and merging datasets

Data cleaning notes for future:
- April 2024 had two testing dates instead of one in May; July 2025 had two testing dates instead of one in August - currently accounted for except July 2025 could be refined as Cades and Jumby sites were added at the end of the month and are currently considered July while 7/29/25 Nonsuch sites are considered August as they were already tested 7/01
- Cades Sculpture Nursery = Cades Nursery (for now, should eventually remove overlap on 4/10)
- Missing 1m depth for CRS in Jan 2025
- YSI omissions and corrections are now made directly within kor_cleaning.Rmd (output = ysi_clean.csv)

```{r}
# setting folder locations

fig_folder <- here("water_quality", "figs", "2026 Jun")

# import various datasets

sites <- read_excel(here("water_quality", "data_raw", "Water quality data.xlsx"), sheet = "sites") %>% 
  clean_names()

site_levels <- sites %>%
  filter(active == "Y") %>%
  pull(site_name) %>%
  unique()

ysi <- list.files(
  here("water_quality", "data_outputs", "kor_exports"),
  pattern = "\\.csv$",
  full.names = TRUE
) %>%
  map_dfr(read_csv, show_col_types = FALSE) %>%
  arrange(desc(date))
write_csv(ysi, here("water_quality", "data_outputs", "ysi_clean.csv"))

entero <- read_excel(here("water_quality", "data_raw", "Water quality data.xlsx"), sheet = "enterococcus") %>% 
  clean_names() %>%
  mutate(n_colonies = as.numeric(str_replace_all(n_colonies, ">", "")))

boat_activity <- read_excel(here("water_quality", "data_raw", "Water quality data.xlsx"), sheet = "boat activity") %>% 
  clean_names() %>%
  mutate(site_code = if_else(site_code == "GIA/LBI", "GIA", site_code)) %>% # coding merged GIA/LBI sites as GIA for now
  select(date, site_code, n_boats)

precip <- read_excel(here("water_quality", "data_raw", "Water quality data.xlsx"), sheet = "precipitation") %>% 
  clean_names() %>%
  select(date, precipitation_in) %>%
  right_join(data.frame(date = seq(ymd('2023-10-10'), as.Date(today()), by='day'))) %>%
  mutate(precipitation_in = replace_na(precipitation_in, 0))

sargassum_local <- read_excel(here("water_quality", "data_raw", "Water quality data.xlsx"), sheet = "sargassum") %>%
  clean_names() %>%
  mutate(date = as.Date(date)) %>%
  select(date, site_code, sargassum, leachate)

sargassum_regional <- read_excel(here("water_quality", "data_raw", "Water quality data.xlsx"), sheet = "sargassum_regional") %>%
  clean_names() %>%
  mutate(date = my(paste(month, year)))

# merge into one dataset

data_mon_dpt <- ysi %>%
  # drop_na() %>%
  left_join(sites, by = c("site_code")) %>%
  left_join(entero, by = c("date", "site_code")) %>%
  left_join(boat_activity, by = c("date", "site_code")) %>%
  left_join(sargassum_local, by = c("date", "site_code")) %>%
  # filter(!is.na(site_name)) %>%
  filter(site_name %in% sites$site_name[sites$active == "Y"], # filtering for only sites with names in active list, but this includes sites with inactive codes (e.g., GEB) but active names
  !(site_code %in% c("GIE", "YIE") & as.Date(date) >= as.Date("2024-08-19"))) %>% # removing YIE/YIC and GIE/GIC overlap during transition of control sites
  mutate(site_code = recode(site_code, "GIE" = "GIC", "YIE" = "YIC"), # making controls continuous
         longitude = as.numeric(longitude),
         latitude = as.numeric(latitude),
         n_colonies = replace_na(as.numeric(str_replace(n_colonies, ">" ,"")), 0),
         entero_yn = if_else(n_colonies == 0, "Absent",
                             if_else(n_colonies > 0, "Present", "NA")),
         n_boats = replace_na(n_boats, 0),
         leachate = replace_na(leachate, 0),
         doy = yday(date),
         month = month(ymd(date), label = TRUE),
         year = year(ymd(date)),
         site_cat = if_else(site_code %in% c("YIE", "GIE", "YIC", "GIC", "JC", "CRS"), "Control", "Treatment"),
         turbidity_fnu = if_else(date %in% c(as.Date("2026-07-08"), as.Date("2023-10-10")), NA_real_, turbidity_fnu)) %>% # sensor giving false readings on these dates
  relocate(date, doy, year, 
           # date_label, date_cat, 
           site_code, site_name, latitude, longitude, depth, monitoring_type) %>%
  select(-notes)
# need to select only active sites, but combine a few incl. control sites

data_mon <- data_mon_dpt %>%
  mutate(depth = if_else(site_code_depth == "CRS3" & date == as.Date("2025-01-20"), 1, depth)) %>% # missing 1m depth for CRS in Jan 2025, so switching 3m to use as 1m instead before dropping depth
  filter(depth == 1) %>% # using only long-term monitoring sites and surface measurements
  select(-depth) %>%
  mutate(site_name = factor(site_name, levels = site_levels)) %>%
  mutate(site_name = fct_rev(site_name)) %>%
  # adding date label to account for months with two testings/no testings
  arrange(site_code, date) %>%
  group_by(site_code) %>%
  mutate(
    month = floor_date(date, "month"),
    month_n = ave(month, month, FUN = length),
    month_row = ave(month, month, FUN = seq_along),
    next_month = month %m+% months(1)
  ) %>%
  mutate(
    next_month_exists = next_month %in% month
  ) %>%
  mutate(
    date_label = case_when(
      month_n > 1 &
      month_row == month_n &
      !next_month_exists ~ next_month,
      TRUE ~ month
    )
  ) %>%
  ungroup() %>%
  mutate(date_label = format(date_label, "%b %Y")) %>%
  select(-c(month_n, month_row, next_month, next_month_exists))
  
#### check samples per date_label - still some issues here?? ####
chk <- data_mon %>%
  count(site_name, date_label) %>%
  filter(n > 1)
```

## Setting target references, plotting ranges, and parameter labels for graphs

```{r}
# sampling date ranges
date_min <- min(data_mon$date)
date_max <- max(data_mon$date)

# pH
lab_ph <- "pH"
ref_ph = data.frame(xmin = -Inf, xmax = Inf, ymin = 7.7, ymax = 8.5)
lab_ref_ph <- "Target range (Rogers et al. 2001)"
ylims_ph <- c((min(data_mon$ph, na.rm = T) - .7), 9)

# turbidity (need to update reference here)
lab_turb <- "Turbidity (FNU)"
ref_turb = data.frame(xmin = -Inf, xmax = Inf, ymin = 0, ymax = 2)
lab_ref_turb <- "Target range (EPA)"
ylims_turb <- c((-1), (max(data_mon$turbidity_fnu, na.rm = T) + 1))

# do
lab_do <- expression("Dissolved oxygen (mg "*L^-1*")")
ref_do = data.frame(xmin = -Inf, xmax = Inf, ymin = 5, ymax = Inf)
lab_ref_do <- "Target range (Long et al. 2013)"
ref_do_hypoxia = 2
lab_ref_do_hypoxia <- "Hypoxia threshold"
ylims_do <- c(0, (max(data_mon$do_mg_l, na.rm = T) + 1))

# temp
lab_temp <- expression("Temperature ("*~degree*C*")")
ref_temp = data.frame(xmin = -Inf, xmax = Inf, ymin = 23, ymax = 29.6)
lab_ref_temp <- "Target range (Coral Reef Alliance)"
ref_temp_bleaching = 30.63
lab_ref_temp_bleaching <- "Coral bleaching threshold (NOAA)"
ylims_temp <- c(22, (max(data_mon$temp_c, na.rm = T) + 1))

# enterococcus
lab_entero <- expression(paste(italic("Enterococcus" ), " spp. (CFU "*mL^-1*")"))
ref_entero = data.frame(xmin = -Inf, xmax = Inf, ymin = 0, ymax = 7/100) # below 7cfu/100mL (EPA)
lab_ref_entero <- "Target range (EPA)"
ylims_entero <- c(0, (max(data_mon$n_colonies, na.rm = T) + 10))

# phycoerythrin
lab_phyco <- "Phycoerythrin (RFU)"
ylims_phyco <- c((min(data_mon$tal_pe_rfu, na.rm = T) - 1), (max(data_mon$tal_pe_rfu, na.rm = T) + 1))

# chlorophyll a
lab_chlor <- expression(paste("Chlorophyll ", italic("a"), " (RFU)"))
ylims_chlor <- c((min(data_mon$chlorophyll_rfu, na.rm = T) - 1), (max(data_mon$chlorophyll_rfu, na.rm = T) + 1))

# precipitation
lab_precip <- expression("Precipitation (in. "*day^-1*")")

# aesthetics for reference rectangles
c_range <- "gray20" # setting color for target range in graphs
a_range <- 0.15 # setting alpha (transparency) for target range in graphs
```

## Writing functions for graph types

```{r}
# flipped point plots by site, with and without reference boxes

flpoint_site <- function(data_wq, y, ylab, ylims) {
  ggplot() +
    geom_point(data = data_wq, 
             aes(site_name, {{y}}),
             alpha = 0.6) +
    labs(x = "", y = ylab) +
    ylim(ylims) +
    theme_bw() +
    coord_flip()
}

flpoint_site_ref <- function(ref_data, data_wq, y, ylab, ylims) {
  ggplot() +
    geom_rect(data = ref_data, 
              aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax, fill = "What"),
              alpha = a_range) +
    scale_fill_manual(values = c_range, guide = "none") +
    geom_point(data = data_wq, 
             aes(site_name, {{y}}),
             alpha = 0.6) +
    labs(x = "", y = ylab) +
    ylim(ylims) +
    theme_bw() +
    coord_flip()
}

# line plots over time, with and without reference boxes

line_time <- function(data_wq, y, ylab, ylims) {
  ggplot() +
    geom_line(data = data_wq %>%
                filter(site_cat == "Treatment"),
            aes(x = date, y = {{y}}, group = site_code),
            color = "lightblue",
            alpha = 0.9) +
    geom_point(data = data_wq %>%
                filter(site_cat == "Treatment"),
            aes(x = date, y = {{y}}, group = site_code),
            color = "lightblue",
            alpha = 0.9) +
    geom_line(data = data_wq %>%
                filter(site_cat == "Control"),
            aes(x = date, y = {{y}}, group = site_code), 
            color = "dodgerblue4",
            alpha = 0.9) +
    geom_point(data = data_wq %>%
                filter(site_cat == "Control"),
            aes(x = date, y = {{y}}, group = site_code), 
            color = "dodgerblue4",
            alpha = 0.9) +
    geom_hline(data = data_wq, # adding geom_hline to work create color legend now that layers are split up 
              aes(yintercept = -100, color = site_cat)) +
    scale_color_manual(values = c("dodgerblue4", "lightblue"), labels = c("Control sites", "Treatment sites")) +
    guides(fill = guide_legend(order = 1), color = guide_legend(order = 2)) +
    scale_x_datetime(date_breaks = "2 months",
                   date_labels = "%b %Y",
                   expand = c(0.02, 0.02)) +
    labs(x = "", y = ylab, color = "") +
    ylim(ylims) +
    theme_bw() +
    theme(legend.position = "top",
          axis.text.x = element_text(angle = 45, hjust = 1))
}


line_time_ref <- function(data_ref, lab_ref, data_wq, y, ylab, ylims) {
  ggplot() +
    geom_rect(data = data_ref, 
              aes(xmin = as.POSIXct(-Inf), xmax = as.POSIXct(Inf), ymin = ymin, ymax = ymax, fill = lab_ref),
              alpha = a_range) +
    scale_fill_manual(values = c_range) +
    geom_line(data = data_wq %>%
                filter(site_cat == "Treatment"),
            aes(x = date, y = {{y}}, group = site_code),
            color = "lightblue",
            label = "Treatment",
            alpha = 0.9) +
    geom_point(data = data_wq %>%
                filter(site_cat == "Treatment"),
            aes(x = date, y = {{y}}, group = site_code),
            color = "lightblue",
            alpha = 0.9) +
    geom_line(data = data_wq %>%
                filter(site_cat == "Control"),
            aes(x = date, y = {{y}}, group = site_code), 
            color = "dodgerblue4",
            alpha = 0.9) +
    geom_point(data = data_wq %>%
                filter(site_cat == "Control"),
            aes(x = date, y = {{y}}, group = site_code), 
            color = "dodgerblue4",
            alpha = 0.9) +
    geom_hline(data = data_wq,
              aes(yintercept = -100, color = site_cat)) +
    scale_color_manual(values = c("dodgerblue4", "lightblue"), labels = c("Control sites", "Treatment sites")) +
    guides(fill = guide_legend(order = 1), color = guide_legend(order = 2)) +
    scale_x_datetime(date_breaks = "2 months",
                   date_labels = "%b %Y",
                   expand = c(0.02, 0.02)) +
    labs(x = "", y = ylab, fill = "", color = "") +
    ylim(ylims) +
    theme_bw() +
    theme(legend.position = "top",
          axis.text.x = element_text(angle = 45, hjust = 1))
}
```

# Methods - site map

## Shapefile
```{r}
# Site only data

# Antigua shapefile
atg <- st_read(here("mapping", "shapefiles", "atg_adm_2019_shp", "atg_admbnda_adm1_2019.shp")) %>%
  st_union() %>%
  st_sf()

# Antigua boundaries
lons = c(-61.94, -61.62)
lats = c(16.98, 17.19)

ggplot() +
  geom_sf(data = atg, fill = "slategray", color = "slategray") + # ATG basemap
  coord_sf(xlim = lons, ylim = lats, expand = FALSE) + # setting map boundaries
  geom_point(data_mon %>% 
               filter(active == "Y") %>% # don't include site_codes with overlapping site names
               select(site_code, site_cat, latitude, longitude) %>%
               distinct(), 
             mapping = aes(x = longitude, y = latitude, shape = site_cat), # 
             alpha = 0.9, color = "black", size = 2) + # add sites
  scale_shape_manual(values = c(1, 16)) +
  labs(shape = "") +
  theme_void() +
  theme(axis.title = element_blank(), 
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        legend.position = "top",
        panel.background = element_rect(fill = "lightblue", color = NA)  # map panel background
        #plot.background = element_rect(fill = "lightblue", color = NA)    # entire plot background
 )

ggsave(here(fig_folder, "sites_map_pres.png"), width = 5, height = 4)
```

## ESRI leaflet
```{r}
library(leaflet)

sites_ongoing <- sites %>%
  filter(monitoring_type == "ongoing") %>%
  mutate(longitude = as.numeric(longitude))

leaflet(sites_ongoing) %>%
  addProviderTiles(providers$Esri.WorldImagery) %>%
  addCircleMarkers(
    lng = ~longitude,
    lat = ~latitude,
    radius = 3,
    color = "orange",       # outline
    fillColor = "orange",   # fill
    fillOpacity = 1,
    weight = 1,
    popup = ~site_name,
    label = ~site_code
  )
```


# Results

## Plotting standard graphs by parameter

```{r}
# plot dimensions
plot_dim_w <- 8.2
plot_dim_h <- 4

# ph
ph_point <- flpoint_site_ref(ref_ph, data_mon, ph, lab_ph, ylims_ph)
ph_line <- line_time_ref(ref_ph, lab_ref_ph, data_mon, ph, lab_ph, ylims_ph)

ggarrange(ph_point, ph_line, 
          ncol = 2, nrow = 1, common.legend = TRUE, legend = "top")
ggsave(here(fig_folder, "ph.png"),  width = plot_dim_w, height = plot_dim_h)

# turbidity
turb_point <- flpoint_site_ref(ref_turb, data_mon, turbidity_fnu, lab_turb, ylims_turb)
turb_line <- line_time_ref(ref_turb, lab_ref_turb, data_mon, turbidity_fnu, lab_turb, ylims_turb)

turb_point <- flpoint_site_ref(ref_turb, filter(data_mon, as.character(date) != "2023-10-10" & turbidity_fnu > -1), turbidity_fnu, lab_turb, ylims_turb) # removing first date and negative values, calibration was off
turb_line <- line_time_ref(ref_turb, lab_ref_turb, filter(data_mon, as.character(date) != "2023-10-10" & turbidity_fnu > -1), turbidity_fnu, lab_turb, ylims_turb)

ggarrange(turb_point, turb_line, 
          ncol = 2, nrow = 1, common.legend = TRUE, legend = "top")
ggsave(here(fig_folder, "turbidity.png"),  width = plot_dim_w, height = plot_dim_h)

# do
do_point <- flpoint_site_ref(ref_do, data_mon, do_mg_l, lab_do, ylims_do)
do_line <- line_time_ref(ref_do, lab_ref_do, data_mon, do_mg_l, lab_do, ylims_do) +
  geom_hline(aes(yintercept = ref_do_hypoxia, linetype = lab_ref_do_hypoxia),
             color = "salmon") +
  scale_linetype_manual(values = "dashed", name = "")

ggarrange(do_point, do_line, 
          ncol = 2, nrow = 1, common.legend = TRUE, legend = "top")
ggsave(here(fig_folder, "do.png"),  width = plot_dim_w, height = plot_dim_h)

# temp
data_mon_temp <- data_mon %>% filter(site_code != "GEB" & date != as.Date("2025-12-02")) # erroneous temp measurement
temp_point <- flpoint_site_ref(ref_temp, data_mon_temp, temp_c, lab_temp, ylims_temp) +
  geom_hline(yintercept = ref_temp_bleaching, color = "salmon", linetype = "dashed")
temp_line <- line_time_ref(ref_temp, lab_ref_temp, data_mon_temp, temp_c, lab_temp, ylims_temp) +
  geom_hline(aes(yintercept = ref_temp_bleaching, linetype = lab_ref_temp_bleaching),
             color = "salmon") +
  scale_linetype_manual(values = "dashed", name = "")

ggarrange(temp_point, temp_line, 
          ncol = 2, nrow = 1, common.legend = TRUE, legend = "top")
ggsave(here(fig_folder, "temp.png"),  width = plot_dim_w, height = plot_dim_h)

# chlorophyll a
chlor_point <- flpoint_site(data_mon, chlorophyll_rfu, lab_chlor, ylims_chlor)
chlor_line <- line_time(data_mon, chlorophyll_rfu, lab_chlor, ylims_chlor)

ggarrange(chlor_point, chlor_line, 
          ncol = 2, nrow = 1, common.legend = TRUE, legend = "top")
ggsave(here(fig_folder, "chlor.png"),  width = plot_dim_w, height = plot_dim_h)

# phycoerythrin
phyco_point <- flpoint_site(data_mon, tal_pe_rfu, lab_phyco, ylims_phyco)
phyco_line <- line_time(data_mon, tal_pe_rfu, lab_phyco, ylims_phyco)

ggarrange(phyco_point, phyco_line, 
          ncol = 2, nrow = 1, common.legend = TRUE, legend = "top")
ggsave(here(fig_folder, "phyco.png"),  width = plot_dim_w, height = plot_dim_h)

# enterococcus
entero_point <- flpoint_site_ref(ref_entero, data_mon, n_colonies, lab_entero, ylims_entero)
entero_line <- line_time_ref(ref_entero, lab_ref_entero, data_mon, n_colonies, lab_entero, ylims_entero)

ggarrange(entero_point, entero_line, 
          ncol = 2, nrow = 1, common.legend = TRUE, legend = "top")
ggsave(here(fig_folder, "entero.png"),  width = plot_dim_w, height = plot_dim_h)
# line only for exec sum
ggsave(here(fig_folder, "entero_line.png"),  plot = entero_line, width = plot_dim_w, height = plot_dim_h)
```

## Additional enterococcus graphs

##### Enterocuccus interannual comparisons

```{r}
# separate data by year
# group by site_year, color by year

temp <- data_mon %>% 
         select(date, month, year, site_code, n_colonies) %>%
         distinct() %>%
  group_by(month, year, site_code) %>%
  summarize(count = n()) %>%
  filter(count > 1)

month_breaks <- yday(ymd(paste0("2023-", 1:12, "-15")))  # mid-month DOYs
month_labels <- month(ymd(paste0("2023-", 1:12, "-01")), label = TRUE, abbr = TRUE)

ggplot(data_mon %>% 
         filter(doy <180) %>%
         select(doy, year, site_code, n_colonies) %>%
         distinct(), 
       aes(x = doy, y = n_colonies, group = interaction(site_code, year), color = factor(year))) +
  geom_line(alpha = .6, size = 1) +
  scale_x_continuous(
    breaks = month_breaks,
    labels = month_labels,
    name = "Month"
  ) +
  labs(color = "Year", y = lab_entero) +
  theme_minimal()
ggsave(here(fig_folder, "entero_interannual.png"), width = 6, height = 4)
```

##### Enterococcus maps

```{r, error = FALSE}
# clean dataset (adding Aug 2025 = 0 here due to missing data)
data_mon_entero <- data_mon %>%
  select(year, date, date_label, site_name, latitude, longitude, n_colonies, entero_yn) %>%
  mutate(assigned_month = my(date_label))

# Antigua shapefile
atg <- st_read(here("mapping", "shapefiles", "atg_adm_2019_shp", "atg_admbnda_adm1_2019.shp")) %>%
  st_union() %>%
  st_sf()

# check lat/long
summary(data_mon$longitude)
summary(data_mon$latitude)

# set lat/lon for graph (map) boundaries
xlim <- range(data_mon$longitude, na.rm = TRUE) + c(-.01, .01)
ylim <- range(data_mon$latitude, na.rm = TRUE) + c(-.01, .01)

# Nonsuch only boundaries
# lons = c(-61.72, -61.65)
# lats = c(17.03, 17.10)

ggplot() +
  geom_sf(data = atg, fill = "slategray", color = "slategray") + # ATG basemap
  coord_sf(xlim = xlim, ylim = ylim, expand = FALSE) + # setting map boundaries
  geom_point(data_mon_entero %>% filter(year >= 2024), 
             mapping = aes(x = longitude, y = latitude, color = entero_yn, size = n_colonies), 
             alpha = 0.7) + # add sites
  scale_color_manual(values = c("turquoise", "tomato")) +
  facet_wrap(~ factor(assigned_month,
             levels = sort(unique(assigned_month)),
             labels = format(sort(unique(assigned_month)), "%b %Y")),
             ncol = 12) +
  labs(size = expression("Concentration (CFU "*mL^-1*")"),
       color = expression("Enterococci detection")) +
  theme_bw() +
  theme(axis.title = element_blank(), 
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        legend.position = "top")

ggsave(here(fig_folder, "entero_map.png"), width = 12, height = 6)
```

##### Boat play

```{r}
ggplot(data_mon %>% filter(site_code %in% c("GIA", "RBA", "NBA", "GOE")), 
       aes(x = n_boats, y = n_colonies, shape = site_name, color = site_name)) +
  geom_point(alpha = 0.6, size = 2) +
  labs(x = "Number of boats in anchorages", y = lab_entero, color = "", shape = "") +
  theme_bw()
ggsave(here(fig_folder, "entero_boats.png"), width = 6, height = 4)
```

##### Enterococcus ~ chlorophyll play

```{r}
entero_phyco <- ggplot(data_mon,
       aes(x = tal_pe_rfu, y = n_colonies)) +
  geom_point(alpha = 0.6, size = 2) +
  sm_statCorr() +
  labs(x = lab_phyco, y = lab_entero) +
  theme_bw()

entero_chlor <- ggplot(data_mon, 
       aes(x = chlorophyll_rfu, y = n_colonies)) +
  geom_point(alpha = 0.6, size = 2) +
  sm_statCorr() +
  labs(x = lab_chlor, y = lab_entero) +
  theme_bw()

ggarrange(entero_phyco, entero_chlor, 
          ncol = 2, nrow = 1)
ggsave(here(fig_folder, "entero_chlor.png"), width = 6, height = 3)
```

## Precipitation

Didn't find much relationship between precipitation and any parameters at this point, it will be helpful to have more data over time. Including exploratory graphs at this point for reference.

```{r}
precip_daily <- precip %>%
  mutate(date = as.Date(date)) %>%
  complete(
    date = seq(min(date), max(date), by = "day"),
    fill = list(precipitation_in = 0)
  ) %>%
  mutate(
    year = factor(year(date)),
    doy = yday(date)
  )

ggplot(precip_daily %>% filter(year(date) >= 2024)) + # first complete year
  geom_line(aes(doy, precipitation_in, colour = year, group = year), alpha = 0.7) +
  scale_x_continuous(
    breaks = yday(as.Date(paste0("2001-", sprintf("%02d", 1:12), "-01"))),
    labels = month.abb
  ) +
  scale_color_manual(values = c(
  "2024" = "navy",  # blue
  "2025" = "#E69F00",  # orange
  "2026" = "#009E73"   # green
)) +
  labs(x = "", y = lab_precip, colour = "Year") +
  theme_bw()
ggsave(here(fig_folder, "precip.png"), width = 6, height = 4)



# trying to calculate the amount of rainfall preceding a testing day
ysi_dates <- ysi %>%
  select(date) %>%
  distinct() # need to clarify which dates were actual monitoring days

# this should be period between testing intervals
precip_mo <- data_mon %>%
  filter(!as.character(date) %in% c("2024-03-11", "2024-08-20", "2023-10-10")) %>% # test dates within 1 day of another test date or dates with no prior data
  select(end_date = date) %>%
  distinct() %>%
  mutate(start_date = end_date - days(31)) %>%
  # mutate(interval = interval(ymd(start_date), ymd(end_date))) %>% # find a way to label this
  fuzzy_left_join(precip, by = c("start_date" = "date", "end_date" = "date"),
                          match_fun = list(`<=`, `>`)) %>%
  group_by(end_date) %>%
  summarize(precip_prior = sum(precipitation_in)) %>%
  mutate(interval = "Month")

precip_wk <- data_mon %>%
  filter(!as.character(date) %in% c("2024-03-11", "2024-08-20", "2023-10-10")) %>% # dates within 1 day of another test or dates with no prior data
  select(end_date = date) %>%
  distinct() %>%
  mutate(start_date = end_date - days(7)) %>%
  # mutate(interval = interval(ymd(start_date), ymd(end_date))) %>% # find a way to label this
  fuzzy_left_join(precip, by = c("start_date" = "date", "end_date" = "date"),
                          match_fun = list(`<=`, `>`)) %>%
  group_by(end_date) %>%
  summarize(precip_prior = sum(precipitation_in)) %>%
  mutate(interval = "Week")

precip_48hr <- data_mon %>%
  filter(!as.character(date) %in% c("2024-03-11", "2024-08-20", "2023-10-10")) %>% # dates within 1 day of another test or dates with no prior data
  select(end_date = date) %>%
  distinct() %>%
  mutate(start_date = end_date - days(2)) %>%
  # mutate(interval = interval(ymd(start_date), ymd(end_date))) %>% # find a way to label this
  fuzzy_left_join(precip, by = c("start_date" = "date", "end_date" = "date"),
                          match_fun = list(`<=`, `>`)) %>%
  group_by(end_date) %>%
  summarize(precip_prior = sum(precipitation_in)) %>%
mutate(interval = "48 hrs")

precip_prior <- rbind(precip_mo, precip_wk, precip_48hr) %>%
  mutate(month = month(ymd(end_date), label = TRUE),
         year = year(ymd(end_date)),
         date_label = paste(as.character(month), as.character(year)),
         date_label = if_else(as.character(end_date) == "2024-04-29", "May 2024", date_label),
         date_label = fct_reorder(date_label, end_date))

# this isn't working fully... multiple values per date_label
# ggplot(precip_prior, aes(x = date_label, y = precip_prior, group = factor(interval, levels = c("Month", "Week", "48 hrs")))) +
#   geom_col(aes(fill = interval), color = "black", position = "dodge") +
#   scale_fill_manual(values = c("cadetblue1", "seagreen", "goldenrod")) +
#   labs(x = "Test Date", y = "Prior precipitation (in)", fill = "Interval prior") +
#   theme_bw() +
#   theme(axis.text.x = element_text(angle = 45, hjust = 1))
#   
# ggsave(here(fig_folder, "precip_prior.png"), width = 6, height = 4)
  
```
```{r}
# dual y attempts

scale = .01 # relates y axis 1 with y axis 2

point_precip <- function(data, y, ylab, scale) {
  ggplot() +
    geom_point(data = data, 
               aes(x = date, y = {{y}}),
               color = "salmon") +
    geom_line(data = precip,
               aes(x = date, y = precipitation_in/scale),
              color = "black") +
    scale_y_continuous(
      name = ylab,
      sec.axis = sec_axis(~.*scale, name = "Precipitation (in/day)")
  ) +
    labs(x = "") +
    theme_bw()
}

point_precip(data_mon, n_colonies, lab_entero, 0.01)
```

## Sargassum

### Regional influx

```{r}
# sargassum accumulation 2023-2025 (from Ruleo)

ggplot() +
  geom_col(data = sargassum_regional, 
             aes(x = as.POSIXct(date), y = quantity_m_tons),
           fill = "orange3") +
  labs(x = "", y = "Caribbean & GOM sargassum influx (mil tonnes)") +
  scale_x_datetime(date_breaks = "3 months",
                   date_labels = "%b %Y",
                   expand = c(0.01, 0.01)) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
ggsave(here(fig_folder, "sargassum_regional.png"), width = 6, height = 4)
```

### Panels

```{r}
sites_sarg <- c("EBB", "MRCH", "BDB", "LDB", "FHB", "NBRM", "NBRB")
date_min_sarg <- min(sargassum_local$date) # date at which we started recording leachate levels

# reference values from Aug. 2021 decay at FHB

ref_sarg_ph <- 7.25
ref_sarg_do <- 0.26 # mg/L # changing to hypoxia threshold here
ref_sarg_turb <- 26.2 # NTU (YSI says no conversion to FNU needed if YSI was used to collect both)
ref_sarg_entero <- 10/100 # CFU/mL

label <- "Decay reference"
ref_sarg_df <- data.frame(label, ref_sarg_ph, ref_sarg_do, ref_sarg_turb, ref_sarg_entero) # had to create df to be able to label reference line in a faceted plot

sarg <- function(data_ref, ref_sarg, y, ylab, ylims) {
  ggplot() +
    geom_rect(data = data_ref, 
              aes(xmin = as.POSIXct(-Inf), xmax = as.POSIXct(Inf), ymin = ymin, ymax = ymax, fill = "Target range"),
              alpha = a_range) +
    scale_fill_manual(values = c_range) +
    geom_hline(data = ref_sarg_df, aes(yintercept = ref_sarg, linetype = label),
             color = "salmon") +
    scale_linetype_manual(values = "dashed") +
    geom_point(data = data_mon_dpt %>% 
                 filter(site_code %in% sites_sarg & date >= date_min_sarg) ,
       aes(x = date, y = {{y}}, color = as.character(depth)),
       alpha = 0.8, size = 2) +
    scale_color_manual(values = c("cadetblue", "darkblue"), labels = c("1m", "3m")) +
    # scale_shape_manual(values = c(16, 18), labels = c("1m", "3m")) +
    facet_wrap(. ~ site_name, nrow = 1, labeller = labeller(site_name = label_wrap_gen(17))) +
    scale_x_datetime(date_breaks = "4 months",
                   date_labels = "%b %Y") +
    ylim(ylims) +
    labs(x = "", y = ylab, shape = "Depth", color = "Depth", fill = "", linetype = "") +
    guides(color = guide_legend(order = 1), linetype = guide_legend(order = 2), linetype = guide_legend(order = 3)) +
    theme_bw()
}

# sargassum leachate levels
leachate_sarg <- ggplot() +
  geom_col(data = data_mon_dpt %>% 
             filter(site_code %in% sites_sarg & date >= date_min_sarg) %>%
             select(date, site_name, leachate) %>%
             distinct(),
       aes(x = date, y = leachate),
       fill = "orange3") +
  scale_x_datetime(date_breaks = "4 months",
                   date_labels = "%b %Y") +
  ylim(0, 2) +
  scale_y_continuous(breaks = seq(0, 1, by = 1)) +
  labs(y = "Leachate score", x = "") +
  facet_wrap(. ~ site_name, nrow = 1, labeller = labeller(site_name = label_wrap_gen(17))) +
  theme_bw() +
  theme(axis.text.x = (element_text(angle = 45, hjust = 1)))

# do
do_sarg <- sarg(ref_do, ref_do_hypoxia, do_mg_l, lab_do, ylims = c(0,9.5)) +
  theme(legend.position = "top",
        axis.text.x = element_blank())

# ph
ph_sarg <- sarg(ref_ph, ref_sarg_ph, ph, lab_ph, ylims_ph) +
  theme(legend.position = "none",
        axis.text.x = element_blank())

# turbidity - excluded for now
turb_sarg <- sarg(ref_turb, ref_sarg_turb, turbidity_fnu, lab_turb, c(0,30)) +
  theme(legend.position = "none",
        axis.text.x = (element_text(angle = 45, hjust = 1)))

ggarrange(
  # Top row: two plots with shared legend
  ggarrange(do_sarg, ph_sarg, 
            ncol = 1, 
            common.legend = TRUE, 
            legend = "top",
            heights = c(4, 4)),
  # Bottom row: third plot without legend
  leachate_sarg + theme(legend.position = "none"),
  nrow = 2,
  heights = c(4, 2)  # Give 2/3 of the height to the top row, 1/3 to the bottom
)
ggsave(here(fig_folder, "sargassum_local.png"),  width = 8, height = 8)

```

### Leachate heatmap

```{r}
# leachate visualizations

plot_dat <- data_mon %>%
  distinct(month, date_label) %>%
  arrange(month) %>%
  pull(date_label)

ggplot(
  data_mon %>%
    mutate(date_label = factor(date_label, levels = plot_dat)) %>%
    complete(site_name, date_label, fill = list(leachate = 0)) %>%
    mutate(leachate = factor(leachate)),
  aes(date_label, site_name, fill = leachate)) +
  scale_fill_manual(
    values = c(
      "0" = "#F7FBFF",
      "1" = "#6BAED6",
      "2" = "#08519C"
    )) +
  geom_tile(color = "grey90", linewidth = 0.3) +
  labs(x = NULL, y = NULL, fill = "Leachate score") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),
        legend.position = "top")
ggsave(here(fig_folder, "sargassum_leachate.png"),  width = 6, height = 4)
```


## Fish kill investigation
```{r}
# subset of sites: control, Little Deep, Big Deep, Boom, Nonsuch Mangroves
sites_select <- c("EBB", "BDB", "LDB", "NBRM", "YIC", "YIE")
data_sub <- data_mon %>% 
  filter(site_code %in% c(sites_select)) %>%
  mutate(site_name = recode(site_name, "Nonsuch Bay Resort Mangrove" = "Nonsuch Mangroves"))

do_point_sub <- flpoint_site_ref(ref_do, data_sub, do_mg_l, lab_do, ylims_do) +
  geom_point(
    data = data_sub,
    aes(site_name, do_mg_l, colour = site_cat),
    alpha = 0.8
  ) +
  scale_colour_manual(values = c("dodgerblue4", "lightblue")) +
  geom_hline(aes(yintercept = ref_do_hypoxia, linetype = lab_ref_do_hypoxia),
             color = "salmon") +
  labs(colour = "") +
  scale_linetype_manual(values = "dashed", name = "")
ggsave(here(fig_folder, "do_subset_line.png"),  width = 5, height = 2)

do_line_sub <- line_time_ref(ref_do, lab_ref_do, data_sub, do_mg_l, lab_do, ylims_do) +
  geom_hline(aes(yintercept = ref_do_hypoxia, linetype = lab_ref_do_hypoxia),
             color = "salmon") +
  scale_linetype_manual(values = "dashed", name = "")

ggarrange(do_point_sub, do_line_sub, 
          ncol = 2, nrow = 1, common.legend = TRUE, legend = "top")
ggsave(here(fig_folder, "do_subset.png"),  width = plot_dim_w, height = plot_dim_h)
```


