---
title: "Water quality testing"
author: "Molly Wilson"
date: "2026-01-01"
output: 
  html_document:
    toc: true
    code_folding: hide
    number_sections: false
    theme: cerulean
    highlight: haddock
---
# Set up
```{r, message = F, warning = F, echo = F}
library(tidyverse)
library(here) 
library(readxl) 
library(janitor)
library(scales)
library(lubridate)
library(sf)

knitr::opts_chunk$set(message = FALSE, warning = FALSE)
```

## Importing, cleaning, and merging datasets

Data cleaning notes for future:
- April 2024 had two testing dates instead of one in May
- Missing 1m depth for CRS in Jan 2025
- YSI omissions and corrections are now made directly within kor_cleaning.Rmd (output = ysi_clean.csv)

```{r}
# setting folder locations

fig_folder <- here("water_quality", "figs", "2026 Jan")

# import various datasets

sites <- read_excel(here("water_quality", "data_raw", "Water quality data.xlsx"), sheet = "sites") %>% 
  clean_names()

ysi <- read_csv(here("water_quality", "data_outputs", "ysi_clean.csv"))

entero <- read_excel(here("water_quality", "data_raw", "Water quality data.xlsx"), sheet = "enterococcus") %>% 
  clean_names() %>%
  mutate(n_colonies = as.numeric(str_replace_all(n_colonies, ">", "")))

boat_activity <- read_excel(here("water_quality", "data_raw", "Water quality data.xlsx"), sheet = "boat activity") %>% 
  clean_names() %>%
  mutate(site_code = if_else(site_code == "GIA/LBI", "GIA", site_code)) %>% # coding merged GIA/LBI sites as GIA for now
  select(date, site_code, n_boats)

precip <- read_excel(here("water_quality", "data_raw", "Water quality data.xlsx"), sheet = "precipitation") %>% 
  clean_names() %>%
  right_join(data.frame(date = seq(ymd('2023-10-10'), as.Date(today()), by='day'))) %>%
  mutate(precipitation_in = replace_na(precipitation_in, 0))

sargassum_local <- read_excel(here("water_quality", "data_raw", "Water quality data.xlsx"), sheet = "sargassum") %>%
  clean_names() %>%
  mutate(date = as.Date(date)) %>%
  select(date, site_code, sargassum, leachate)

# merge into one dataset

data_mon <- ysi %>%
  # drop_na() %>%
  left_join(sites, by = c("site_code")) %>%
  left_join(entero, by = c("date", "site_code")) %>%
  left_join(boat_activity, by = c("date", "site_code")) %>%
  left_join(sargassum_local, by = c("date", "site_code")) %>%
  mutate(depth = if_else(site_code_depth == "CRS3" & date == as.Date("2025-01-20"), 1, depth)) %>% # missing 1m depth for CRS in Jan 2025, so switching 3m to use as 1m instead before dropping depth
  filter(!is.na(site_name) & depth == 1) %>% # using only long-term monitoring sites and surface measurements
  mutate(longitude = as.numeric(longitude),
         latitude = as.numeric(latitude),
         n_colonies = replace_na(as.numeric(str_replace(n_colonies, ">" ,"")), 0),
         entero_yn = if_else(n_colonies == 0, "Absent",
                             if_else(n_colonies > 0, "Present", "NA")),
         n_boats = replace_na(n_boats, 0),
         leachate = replace_na(leachate, 0),
         doy = yday(date),
         month = month(ymd(date), label = TRUE),
         year = year(ymd(date)),
         date_label = paste(as.character(month), as.character(year)),
         date_label = if_else(as.character(date) == "2024-04-29", "May 2024", date_label), # label as next month due to spread
         date_cat = if_else(substr(date_label, 1, 3) %in% c("Dec", "Jan", "Feb", "Mar", "Apr"), "High season", "Low season"),
         site_cat = if_else(site_code %in% c("YIE", "GIE", "YIC", "GIC", "JC", "CRS"), "Control", "Treatment")) %>%
  relocate(date, doy, year, date_label, date_cat, site_code, site_name, latitude, longitude)
  # distinct()

chk <- data_mon %>%
  complete(date_label, site_code, fill = list(depth = NA)) %>% # fill can be NA or other defaults
  select(date_label, site_code, date) %>%
  mutate(
    date_label_date = parse_date_time(date_label, orders = "b Y"), # "Apr 2024" → 2024-04-01
    date_label_date = as.Date(date_label_date)
  ) %>%
  arrange(site_code, date_label_date)
```

## Setting target references, plotting ranges, and parameter labels for graphs

```{r}
# sampling date ranges
date_min <- min(data_mon$date)
date_max <- max(data_mon$date)

# pH
lab_ph <- "pH"
ref_ph = data.frame(xmin = -Inf, xmax = Inf, ymin = 7.7, ymax = 8.5)
lab_ref_ph <- "Target range (Rogers et al. 2001)"
ylims_ph <- c((min(data_mon$ph, na.rm = T) - .7), 9)

# turbidity (need to update reference here)
lab_turb <- "Turbidity (FNU)"
ref_turb = data.frame(xmin = -Inf, xmax = Inf, ymin = 0, ymax = 2)
lab_ref_turb <- "Target range (EPA)"
ylims_turb <- c((min(data_mon$turbidity_fnu, na.rm = T) - 1), (max(data_mon$turbidity_fnu, na.rm = T) + 1))

# do
lab_do <- expression("Dissolved oxygen (mg "*L^-1*")")
ref_do = data.frame(xmin = -Inf, xmax = Inf, ymin = 5, ymax = Inf)
lab_ref_do <- "Target range (Long et al. 2013)"
ref_do_hypoxia = 2
lab_ref_do_hypoxia <- "Hypoxia threshold"
ylims_do <- c(0, (max(data_mon$do_mg_l, na.rm = T) + 1))

# temp
lab_temp <- expression("Temperature ("*~degree*C*")")
ref_temp = data.frame(xmin = -Inf, xmax = Inf, ymin = 23, ymax = 29.6)
lab_ref_temp <- "Target range (Coral Reef Alliance)"
ref_temp_bleaching = 30.63
lab_ref_temp_bleaching <- "Coral bleaching threshold (NOAA)"
ylims_temp <- c(22, (max(data_mon$temp_c, na.rm = T) + 1))

# enterococcus
lab_entero <- expression(paste(italic("Enterococcus" ), " spp. (CFU "*mL^-1*")"))
ref_entero = data.frame(xmin = -Inf, xmax = Inf, ymin = 0, ymax = 7/100) # below 7cfu/100mL (EPA)
lab_ref_entero <- "Target range (EPA)"
ylims_entero <- c(0, (max(data_mon$n_colonies, na.rm = T) + 10))

# phycoerythrin
lab_phyco <- "Phycoerythrin (RFU)"
ylims_phyco <- c((min(data_mon$tal_pe_rfu, na.rm = T) - 1), (max(data_mon$tal_pe_rfu, na.rm = T) + 1))

# chlorophyll a
lab_chlor <- expression(paste("Chlorophyll ", italic("a"), " (RFU)"))
ylims_chlor <- c((min(data_mon$chlorophyll_rfu, na.rm = T) - 1), (max(data_mon$chlorophyll_rfu, na.rm = T) + 1))

# precipitation
lab_precip <- expression("Precipitation (in. "*day^-1*")")

# aesthetics for reference rectangles
c_range <- "gray20" # setting color for target range in graphs
a_range <- 0.15 # setting alpha (transparency) for target range in graphs
```

## Writing functions for graph types

```{r}
# flipped point plots by site, with and without reference boxes

flpoint_site <- function(data_wq, y, ylab, ylims) {
  ggplot() +
    geom_point(data = data_wq, 
             aes(site_name, {{y}}),
             alpha = 0.6) +
    labs(x = "", y = ylab) +
    ylim(ylims) +
    theme_bw() +
    coord_flip()
}

flpoint_site_ref <- function(ref_data, data_wq, y, ylab, ylims) {
  ggplot() +
    geom_rect(data = ref_data, 
              aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax, fill = "What"),
              alpha = a_range) +
    scale_fill_manual(values = c_range, guide = "none") +
    geom_point(data = data_wq, 
             aes(site_name, {{y}}),
             alpha = 0.6) +
    labs(x = "", y = ylab) +
    ylim(ylims) +
    theme_bw() +
    coord_flip()
}

# line plots over time, with and without reference boxes

line_time <- function(data_wq, y, ylab, ylims) {
  ggplot() +
    geom_line(data = data_wq %>%
                filter(site_cat == "Treatment"),
            aes(x = date, y = {{y}}, group = site_code),
            color = "lightblue",
            alpha = 0.9) +
    geom_point(data = data_wq %>%
                filter(site_cat == "Treatment"),
            aes(x = date, y = {{y}}, group = site_code),
            color = "lightblue",
            alpha = 0.9) +
    geom_line(data = data_wq %>%
                filter(site_cat == "Control"),
            aes(x = date, y = {{y}}, group = site_code), 
            color = "dodgerblue4",
            alpha = 0.9) +
    geom_point(data = data_wq %>%
                filter(site_cat == "Control"),
            aes(x = date, y = {{y}}, group = site_code), 
            color = "dodgerblue4",
            alpha = 0.9) +
    geom_hline(data = data_wq, # adding geom_hline to work create color legend now that layers are split up 
              aes(yintercept = -100, color = site_cat)) +
    scale_color_manual(values = c("dodgerblue4", "lightblue"), labels = c("Control sites", "Treatment sites")) +
    guides(fill = guide_legend(order = 1), color = guide_legend(order = 2)) +
    scale_x_datetime(date_breaks = "2 months",
                   date_labels = "%b %Y",
                   expand = c(0.02, 0.02)) +
    labs(x = "", y = ylab, color = "") +
    ylim(ylims) +
    theme_bw() +
    theme(legend.position = "top",
          axis.text.x = element_text(angle = 45, hjust = 1))
}


line_time_ref <- function(data_ref, lab_ref, data_wq, y, ylab, ylims) {
  ggplot() +
    geom_rect(data = data_ref, 
              aes(xmin = as.POSIXct(-Inf), xmax = as.POSIXct(Inf), ymin = ymin, ymax = ymax, fill = lab_ref),
              alpha = a_range) +
    scale_fill_manual(values = c_range) +
    geom_line(data = data_wq %>%
                filter(site_cat == "Treatment"),
            aes(x = date, y = {{y}}, group = site_code),
            color = "lightblue",
            label = "Treatment",
            alpha = 0.9) +
    geom_point(data = data_wq %>%
                filter(site_cat == "Treatment"),
            aes(x = date, y = {{y}}, group = site_code),
            color = "lightblue",
            alpha = 0.9) +
    geom_line(data = data_wq %>%
                filter(site_cat == "Control"),
            aes(x = date, y = {{y}}, group = site_code), 
            color = "dodgerblue4",
            alpha = 0.9) +
    geom_point(data = data_wq %>%
                filter(site_cat == "Control"),
            aes(x = date, y = {{y}}, group = site_code), 
            color = "dodgerblue4",
            alpha = 0.9) +
    geom_hline(data = data_wq,
              aes(yintercept = -100, color = site_cat)) +
    scale_color_manual(values = c("dodgerblue4", "lightblue"), labels = c("Control sites", "Treatment sites")) +
    guides(fill = guide_legend(order = 1), color = guide_legend(order = 2)) +
    scale_x_datetime(date_breaks = "2 months",
                   date_labels = "%b %Y",
                   expand = c(0.02, 0.02)) +
    labs(x = "", y = ylab, fill = "", color = "") +
    ylim(ylims) +
    theme_bw() +
    theme(legend.position = "top",
          axis.text.x = element_text(angle = 45, hjust = 1))
}
```

# Methods - site map

```{r}
# Site only data

# Antigua shapefile
atg <- st_read(here("mapping", "shapefiles", "atg_adm_2019_shp", "atg_admbnda_adm1_2019.shp")) %>%
  st_union() %>%
  st_sf()

# Antigua boundaries
lons = c(-61.92, -61.65)
lats = c(16.98, 17.19)

ggplot() +
  geom_sf(data = atg, fill = "slategray", color = "slategray") + # ATG basemap
  coord_sf(xlim = lons, ylim = lats, expand = FALSE) + # setting map boundaries
  geom_point(data_mon %>% 
               select(site_code, site_cat, latitude, longitude) %>%
               distinct(), 
             mapping = aes(x = longitude, y = latitude, shape = site_cat), 
             alpha = 0.9, color = "black", size = 1) + # add sites
  scale_shape_manual(values = c(1, 16)) +
  labs(shape = "") +
  theme_minimal() +
  theme(axis.title = element_blank(), 
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        legend.position = "top",
        panel.background = element_rect(fill = "lightblue", color = NA)  # map panel background
        #plot.background = element_rect(fill = "lightblue", color = NA)    # entire plot background
 )

### missing Jumby Control lat/lon
ggsave(here(fig_folder, "sites_map.png"), width = 5, height = 4)
```

# Results

## Plotting standard graphs by parameter

```{r}
# plot dimensions
plot_dim_w <- 8.2
plot_dim_h <- 4

# ph
ph_point <- flpoint_site_ref(ref_ph, data_mon, ph, lab_ph, ylims_ph)
ph_line <- line_time_ref(ref_ph, lab_ref_ph, data_mon, ph, lab_ph, ylims_ph)

ggarrange(ph_point, ph_line, 
          ncol = 2, nrow = 1, common.legend = TRUE, legend = "top")
ggsave(here(fig_folder, "ph.png"),  width = plot_dim_w, height = plot_dim_h)

# turbidity
turb_point <- flpoint_site_ref(ref_turb, filter(data_mon, as.character(date) != "2023-10-10"), turbidity_fnu, lab_turb, ylims_turb) # removing first date, calibration was off
turb_line <- line_time_ref(ref_turb, lab_ref_turb, filter(data_mon, as.character(date) != "2023-10-10"), turbidity_fnu, lab_turb, ylims_turb)

ggarrange(turb_point, turb_line, 
          ncol = 2, nrow = 1, common.legend = TRUE, legend = "top")
ggsave(here(fig_folder, "turbidity.png"),  width = plot_dim_w, height = plot_dim_h)

# do
do_point <- flpoint_site_ref(ref_do, data_mon, do_mg_l, lab_do, ylims_do)
do_line <- line_time_ref(ref_do, lab_ref_do, data_mon, do_mg_l, lab_do, ylims_do) +
  geom_hline(aes(yintercept = ref_do_hypoxia, linetype = lab_ref_do_hypoxia),
             color = "salmon") +
  scale_linetype_manual(values = "dashed", name = "")

ggarrange(do_point, do_line, 
          ncol = 2, nrow = 1, common.legend = TRUE, legend = "top")
ggsave(here(fig_folder, "do.png"),  width = plot_dim_w, height = plot_dim_h)

# temp
data_mon_temp <- data_mon %>% filter(site_code != "GEB")
temp_point <- flpoint_site_ref(ref_temp, data_mon_temp, temp_c, lab_temp, ylims_temp) +
  geom_hline(yintercept = ref_temp_bleaching, color = "salmon", linetype = "dashed")
temp_line <- line_time_ref(ref_temp, lab_ref_temp, data_mon_temp, temp_c, lab_temp, ylims_temp) +
  geom_hline(aes(yintercept = ref_temp_bleaching, linetype = lab_ref_temp_bleaching),
             color = "salmon") +
  scale_linetype_manual(values = "dashed", name = "")

ggarrange(temp_point, temp_line, 
          ncol = 2, nrow = 1, common.legend = TRUE, legend = "top")
ggsave(here(fig_folder, "temp.png"),  width = plot_dim_w, height = plot_dim_h)

# chlorophyll a
chlor_point <- flpoint_site(data_mon, chlorophyll_rfu, lab_chlor, ylims_chlor)
chlor_line <- line_time(data_mon, chlorophyll_rfu, lab_chlor, ylims_chlor)

ggarrange(chlor_point, chlor_line, 
          ncol = 2, nrow = 1, common.legend = TRUE, legend = "top")
ggsave(here(fig_folder, "chlor.png"),  width = plot_dim_w, height = plot_dim_h)

# phycoerythrin
phyco_point <- flpoint_site(data_mon, tal_pe_rfu, lab_phyco, ylims_phyco)
phyco_line <- line_time(data_mon, tal_pe_rfu, lab_phyco, ylims_phyco)

ggarrange(phyco_point, phyco_line, 
          ncol = 2, nrow = 1, common.legend = TRUE, legend = "top")
ggsave(here(fig_folder, "phyco.png"),  width = plot_dim_w, height = plot_dim_h)

# enterococcus
entero_point <- flpoint_site_ref(ref_entero, data_mon, n_colonies, lab_entero, ylims_entero)
entero_line <- line_time_ref(ref_entero, lab_ref_entero, data_mon, n_colonies, lab_entero, ylims_entero)

ggarrange(entero_point, entero_line, 
          ncol = 2, nrow = 1, common.legend = TRUE, legend = "top")
ggsave(here(fig_folder, "entero.png"),  width = plot_dim_w, height = plot_dim_h)
```

## Additional enterococcus graphs

##### Enterocuccus interannual comparisons

```{r}
# separate data by year
# group by site_year, color by year

temp <- data_mon %>% 
         select(date, month, year, site_code, n_colonies) %>%
         distinct() %>%
  group_by(month, year, site_code) %>%
  summarize(count = n()) %>%
  filter(count > 1)

month_breaks <- yday(ymd(paste0("2023-", 1:12, "-15")))  # mid-month DOYs
month_labels <- month(ymd(paste0("2023-", 1:12, "-01")), label = TRUE, abbr = TRUE)

ggplot(data_mon %>% 
         filter(doy <180) %>%
         select(doy, year, site_code, n_colonies) %>%
         distinct(), 
       aes(x = doy, y = n_colonies, group = interaction(site_code, year), color = factor(year))) +
  geom_line(alpha = .6, size = 1) +
  scale_x_continuous(
    breaks = month_breaks,
    labels = month_labels,
    name = "Month"
  ) +
  labs(color = "Year", y = lab_entero) +
  theme_minimal()
ggsave(here(fig_folder, "entero_interannual.png"), width = 6, height = 4)
```

##### Enterococcus maps

```{r, error = FALSE}
# Antigua shapefile
atg <- st_read(here("mapping", "shapefiles", "atg_adm_2019_shp", "atg_admbnda_adm1_2019.shp")) %>%
  st_union() %>%
  st_sf()

# check lat/long
summary(data_mon$longitude)
summary(data_mon$latitude)

# set lat/lon for graph (map) boundaries
xlim <- range(data_mon$longitude, na.rm = TRUE) + c(-.01, .01)
ylim <- range(data_mon$latitude, na.rm = TRUE) + c(-.01, .01)

# Nonsuch only boundaries
lons = c(-61.72, -61.65)
lats = c(17.03, 17.10)

ggplot() +
  geom_sf(data = atg, fill = "slategray", color = "slategray") + # ATG basemap
  coord_sf(xlim = lons, ylim = lats, expand = FALSE) + # setting map boundaries
  geom_point(data_mon, 
             mapping = aes(x = longitude, y = latitude, color = entero_yn, size = n_colonies), 
             alpha = 0.7) + # add sites
  scale_color_manual(values = c("turquoise", "tomato")) +
  facet_wrap(. ~ factor(date_label, levels = c("Oct 2023", "Nov 2023", "Dec 2023", "Jan 2024", "Feb 2024", "Mar 2024", "Apr 2024", "May 2024", "Jun 2024", "Jul 2024", "Aug 2024", "Sep 2024", "Oct 2024", "Nov 2024", "Dec 2024", "Jan 2025", "Feb 2025", "Mar 2025", "Apr 2025", "May 2025", "Jun 2025")),
             ncol = 12) +
  labs(size = expression("Concentration (CFU "*mL^-1*")"),
       color = expression("Enterococci detection")) +
  theme_bw() +
  theme(axis.title = element_blank(), 
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        legend.position = "top")

ggsave(here(fig_folder, "entero_map.png"), width = 10, height = 9)
```

##### Boat play

```{r}
ggplot(data_mon %>% filter(site_code %in% c("GIA", "RBA", "NBA", "GOE")), 
       aes(x = n_boats, y = n_colonies, shape = site_name, color = site_name)) +
  geom_point(alpha = 0.6, size = 2) +
  labs(x = "Number of boats in anchorages", y = lab_entero, color = "", shape = "") +
  theme_bw()
ggsave(here(fig_folder, "entero_boats.png"), width = 6, height = 4)
```

##### Enterococcus ~ chlorophyll play

```{r}
entero_phyco <- ggplot(data_mon,
       aes(x = tal_pe_rfu, y = n_colonies)) +
  geom_point(alpha = 0.6, size = 2) +
  sm_statCorr() +
  labs(x = lab_phyco, y = lab_entero) +
  theme_bw()

entero_chlor <- ggplot(data_mon, 
       aes(x = chlorophyll_rfu, y = n_colonies)) +
  geom_point(alpha = 0.6, size = 2) +
  sm_statCorr() +
  labs(x = lab_chlor, y = lab_entero) +
  theme_bw()

ggarrange(entero_phyco, entero_chlor, 
          ncol = 2, nrow = 1)
ggsave(here(fig_folder, "entero_chlor.png"), width = 6, height = 3)
```

## Precipitation

Didn't find much relationship between precipitation and any parameters at this point, it will be helpful to have more data over time. Including exploratory graphs at this point for reference.

```{r}
ggplot() +
  geom_line(data = precip %>% 
              filter(as.Date(date) < date_max), 
             aes(x = as.Date(date), y = precipitation_in)) +
  labs(x = "", y = lab_precip) +
  scale_x_date(
    limits = as.Date(c("2023-11-01", "2025-06-30")),
    date_breaks = "2 months",
    date_labels = "%b %Y") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
ggsave(here(fig_folder, "precip.png"), width = 6, height = 4)

# trying to calculate the amount of rainfall preceding a testing day
ysi_dates <- ysi %>%
  select(date) %>%
  distinct() # need to clarify which dates were actual monitoring days

# this should be period between testing intervals
precip_mo <- data_mon %>%
  filter(!as.character(date) %in% c("2024-03-11", "2024-08-20", "2023-10-10")) %>% # test dates within 1 day of another test date or dates with no prior data
  select(end_date = date) %>%
  distinct() %>%
  mutate(start_date = end_date - days(31)) %>%
  # mutate(interval = interval(ymd(start_date), ymd(end_date))) %>% # find a way to label this
  fuzzy_left_join(precip, by = c("start_date" = "date", "end_date" = "date"),
                          match_fun = list(`<=`, `>`)) %>%
  group_by(end_date) %>%
  summarize(precip_prior = sum(precipitation_in)) %>%
  mutate(interval = "Month")

precip_wk <- data_mon %>%
  filter(!as.character(date) %in% c("2024-03-11", "2024-08-20", "2023-10-10")) %>% # dates within 1 day of another test or dates with no prior data
  select(end_date = date) %>%
  distinct() %>%
  mutate(start_date = end_date - days(7)) %>%
  # mutate(interval = interval(ymd(start_date), ymd(end_date))) %>% # find a way to label this
  fuzzy_left_join(precip, by = c("start_date" = "date", "end_date" = "date"),
                          match_fun = list(`<=`, `>`)) %>%
  group_by(end_date) %>%
  summarize(precip_prior = sum(precipitation_in)) %>%
  mutate(interval = "Week")

precip_48hr <- data_mon %>%
  filter(!as.character(date) %in% c("2024-03-11", "2024-08-20", "2023-10-10")) %>% # dates within 1 day of another test or dates with no prior data
  select(end_date = date) %>%
  distinct() %>%
  mutate(start_date = end_date - days(2)) %>%
  # mutate(interval = interval(ymd(start_date), ymd(end_date))) %>% # find a way to label this
  fuzzy_left_join(precip, by = c("start_date" = "date", "end_date" = "date"),
                          match_fun = list(`<=`, `>`)) %>%
  group_by(end_date) %>%
  summarize(precip_prior = sum(precipitation_in)) %>%
mutate(interval = "48 hrs")

precip_prior <- rbind(precip_mo, precip_wk, precip_48hr) %>%
  mutate(month = month(ymd(end_date), label = TRUE),
         year = year(ymd(end_date)),
         date_label = paste(as.character(month), as.character(year)),
         date_label = if_else(as.character(end_date) == "2024-04-29", "May 2024", date_label),
         date_label = fct_reorder(date_label, end_date))

ggplot(precip_prior, aes(x = date_label, y = precip_prior, group = factor(interval, levels = c("Month", "Week", "48 hrs")))) +
  geom_col(aes(fill = interval), color = "black", position = "dodge") +
  scale_fill_manual(values = c("cadetblue1", "seagreen", "goldenrod")) +
  labs(x = "Test Date", y = "Prior precipitation (in)", fill = "Interval prior") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
  
ggsave(here(fig_folder, "precip_prior.png"), width = 6, height = 4)
  
```
```{r}
# dual y attempts

scale = .01 # relates y axis 1 with y axis 2

point_precip <- function(data, y, ylab, scale) {
  ggplot() +
    geom_point(data = data, 
               aes(x = date, y = {{y}}),
               color = "salmon") +
    geom_line(data = precip,
               aes(x = date, y = precipitation_in/scale),
              color = "black") +
    scale_y_continuous(
      name = ylab,
      sec.axis = sec_axis(~.*scale, name = "Precipitation (in/day)")
  ) +
    labs(x = "") +
    theme_bw()
}

point_precip(data_mon, n_colonies, lab_entero, 0.01)
```

## Sargassum

```{r}
# sargassum accumulation 2023-2024 (from Ruleo)

ggplot() +
  geom_col(data = sargassum_regional, 
             aes(x = as.POSIXct(date), y = sargassum_mil_tonnes),
           fill = "orange3") +
  labs(x = "", y = "Caribbean & GOM sargassum influx (mil tonnes)") +
  scale_x_datetime(date_breaks = "2 months",
                   date_labels = "%b %Y",
                   expand = c(0.01, 0.01)) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
ggsave(here(fig_folder, "sargassum.png"), width = 6, height = 4)
```


```{r}
sites_sarg <- c("EBB", "MRCH", "BDB", "LDB", "FHB", "NBRM", "NBRB")
date_min_sarg <- min(sargassum_local$date) # date at which we started recording leachate levels

# reference values from Aug. 2021 decay at FHB

ref_sarg_ph <- 7.25
ref_sarg_do <- 0.26 # mg/L # changing to hypoxia threshold here
ref_sarg_turb <- 26.2 # NTU (YSI says no conversion to FNU needed if YSI was used to collect both)
ref_sarg_entero <- 10/100 # CFU/mL

label <- "Decay reference"
ref_sarg_df <- data.frame(label, ref_sarg_ph, ref_sarg_do, ref_sarg_turb, ref_sarg_entero) # had to create df to be able to label reference line in a faceted plot

sarg <- function(data_ref, ref_sarg, y, ylab, ylims) {
  ggplot() +
    geom_rect(data = data_ref, 
              aes(xmin = as.POSIXct(-Inf), xmax = as.POSIXct(Inf), ymin = ymin, ymax = ymax, fill = "Target range"),
              alpha = a_range) +
    scale_fill_manual(values = c_range) +
    geom_hline(data = ref_sarg_df, aes(yintercept = ref_sarg, linetype = label),
             color = "salmon") +
    scale_linetype_manual(values = "dashed") +
    geom_point(data = wq_data %>% 
                 filter(site_code %in% sites_sarg & date >= date_min_sarg) ,
       aes(x = date, y = {{y}}, color = as.character(depth_m)),
       alpha = 0.8, size = 2) +
    scale_color_manual(values = c("cadetblue", "darkblue"), labels = c("1m", "3m")) +
    # scale_shape_manual(values = c(16, 18), labels = c("1m", "3m")) +
    facet_wrap(. ~ site_name, nrow = 1, labeller = labeller(site_name = label_wrap_gen(17))) +
    scale_x_datetime(date_breaks = "4 months",
                   date_labels = "%b %Y") +
    ylim(ylims) +
    labs(x = "", y = ylab, shape = "Depth", color = "Depth", fill = "", linetype = "") +
    guides(color = guide_legend(order = 1), linetype = guide_legend(order = 2), linetype = guide_legend(order = 3)) +
    theme_bw()
}

# sargassum leachate levels
leachate_sarg <- ggplot() +
  geom_col(data = wq_data %>% 
             filter(site_code %in% sites_sarg & date >= date_min_sarg) %>%
             select(date, site_name, leachate) %>%
             distinct(),
       aes(x = date, y = leachate),
       fill = "orange3") +
  scale_x_datetime(date_breaks = "2 months",
                   date_labels = "%b %Y") +
  ylim(0, 2) +
  scale_y_continuous(breaks = seq(0, 1, by = 1)) +
  labs(y = "Leachate score", x = "") +
  facet_wrap(. ~ site_name, nrow = 1, labeller = labeller(site_name = label_wrap_gen(17))) +
  theme_bw() +
  theme(axis.text.x = (element_text(angle = 45, hjust = 1)))


# do
do_sarg <- sarg(ref_do, ref_do_hypoxia, do_mg_l, lab_do, ylims = c(0,9.5)) +
  theme(legend.position = "top",
        axis.text.x = element_blank())

# ph
ph_sarg <- sarg(ref_ph, ref_sarg_ph, ph, lab_ph, ylims_ph) +
  theme(legend.position = "none",
        axis.text.x = element_blank())

# turbidity - excluded for now
turb_sarg <- sarg(ref_turb, ref_sarg_turb, turbidity_fnu, lab_turb, c(0,30)) +
  theme(legend.position = "none",
        axis.text.x = (element_text(angle = 45, hjust = 1)))

ggarrange(
  # Top row: two plots with shared legend
  ggarrange(do_sarg, ph_sarg, 
            ncol = 1, 
            common.legend = TRUE, 
            legend = "top",
            heights = c(4, 4)),
  # Bottom row: third plot without legend
  leachate_sarg + theme(legend.position = "none"),
  nrow = 2,
  heights = c(4, 2)  # Give 2/3 of the height to the top row, 1/3 to the bottom
)
ggsave(here(fig_folder, "sarg.png"),  width = 8, height = 8)

```

