---
title: "MMA Seagrass"
author: "Molly Wilson"
date: "6/18/2022"
output: 
  html_document:
    code_folding: hide
---
# {.tabset}

```{r, message = F, warning = F, echo = F}
library(tidyverse)
library(here) 
library(readxl) 
library(janitor)
library(snakecase) # for adjusting capitalization of text within data (e.g., species names)
library(knitr) # for including tables

knitr::opts_chunk$set(message = FALSE, warning = FALSE) # sets default for each chunk
```

## Seagrass

```{r}
# Import and clean data

sg_quads <- read_excel(here("mma", "data", "mma_seagrass.xlsx"), sheet = "quadrants") %>%  
  clean_names()
```

### Percent cover by category
* technically I think we should be removing other invertebrates...?
```{r}
# expanding dataset by category
sg_cat <- sg_quads %>%
  expand(nesting(site, transect, quadrant), category) %>%
  left_join(sg_quads %>%
              select(site, transect, quadrant, category, percent, surveyor)) %>%
  mutate(percent = if_else(is.na(percent), 0, percent)) %>%
  group_by(site, transect, quadrant, category) %>%
  summarize(pc = sum(percent))

sg_pc_cat_site <- sg_cat %>%
  group_by(site, transect, category) %>%
  summarize(pc_t = mean(pc),
            n = n()) %>% # testing to make sure there are 10 entries per category within each transect
  group_by(site, category) %>%
  summarize(pc_mean = mean(pc_t),
            pc_se = sd(pc_t)/sqrt(n()))

sg_pc_cat <- sg_pc_cat_site %>%
  group_by(category) %>%
  summarize(pc_mean_sum = mean(pc_mean),
            pc_se = sd(pc_mean)/sqrt(n())) %>%
  rename(pc_mean = pc_mean_sum)

# test percent cover adds up to 100 per quadrant, transect, and site
test1 <- sg_cat %>%
  group_by(site, transect, quadrant) %>%
  summarize(total = sum(pc)) %>%
  filter(total != 100) %>%
  left_join(sg_quads %>%
              select(surveyor, site, transect, quadrant)) %>%
  distinct()
test2 <- sg_pc_cat_site %>%
  group_by(site) %>%
  summarize(total = sum(pc_mean))

# graphs
cat_palette <- c("olivedrab", "wheat", "palevioletred4", "tan3", "darkseagreen4", "slategray3", "coral2")

ggplot(sg_pc_cat, 
       aes(x = reorder(category, pc_mean), y = pc_mean, fill = category)) +
  geom_col(color = "black", alpha = 0.9) +
  geom_errorbar(aes(ymin = pc_mean - pc_se, ymax = pc_mean + pc_se), width = .2,
                 position = position_dodge(.9)) +
  scale_fill_manual(values = cat_palette) +
  coord_flip() +
  labs(y = "Mean percent cover", x = "Category", fill = "") +
  theme_bw() +
  theme(axis.text.y = element_text(face = "italic"))

kable(sg_pc_cat %>% select(category, pc_mean, pc_se))

ggplot(sg_pc_cat_site,
       aes(x = category, y = pc_mean, fill = category)) +
  geom_col(color = "black", alpha = 0.9) +
  geom_errorbar(aes(ymin = pc_mean - pc_se, ymax = pc_mean + pc_se), width = .2,
                 position = position_dodge(.9)) +
    scale_fill_manual(values = cat_palette) +
  facet_wrap(. ~ site) +
  labs(y = "Mean percent cover", x = "Category", fill = "") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, h = 1, face = "italic"))

ggplot(sg_pc_cat_site, 
       aes(x = "category", y = pc_mean, fill = category)) + 
  geom_bar(width = 1, stat = "identity", color = "black") +
  coord_polar("y", start=0) +
  scale_fill_manual(values = cat_palette) +
  facet_wrap(vars(site), nrow = 2) +
  theme_void() +
  theme(legend.position = "bottom",
        legend.title = element_blank(),
        panel.spacing = unit(1, "lines"))

kable(sg_pc_cat_site %>% select(site, category, pc_mean, pc_se))
```

### Percent cover by species
* Only hard corals and seagrasses were identified to the species level
* Algae was identified to genus level (would love to figure out how to make "spp." not italicized, or could convert all to genus!)

```{r}
# expanding dataset by species
sg_spp <- sg_quads %>%
  expand(nesting(site, transect, quadrant), species) %>% ## try adding nesting(species, category)
  left_join(sg_quads %>% select(site, transect, quadrant, species, percent)) %>%
  left_join(sg_quads %>% select (species, category)) %>% # not sure why this wasn't working in the same line
  mutate(percent = if_else(is.na(percent), 0, percent)) %>% # can't get replace_na to work so using this instead
  distinct()

# calculating percent cover by transect -> site
sg_pc_spp_site <- sg_spp %>%
  group_by(site, transect, species, category) %>%
  summarize(pc_t = mean(percent)) %>%
  group_by(site, species, category) %>%
  summarize(pc_mean = mean(pc_t),
            pc_se = sd(pc_t)/sqrt(n()))

# calculating percent cover across all sites
sg_pc_spp <- sg_pc_spp_site %>%
  group_by(species, category) %>%
  summarize(pc_mean_sum = mean(pc_mean),
            pc_se = sd(pc_mean)/sqrt(n())) %>%
  rename(pc_mean = pc_mean_sum)

# testing accuracy of site-level calculations
test1 <- sg_quads %>% filter(site == "Garden of Eden"  & species == "Penicillus spp.")
test2 <- sg_spp %>% filter(site == "Garden of Eden" & species == "Penicillus spp.")
test3 <- sg_spp %>%
  group_by(site, transect, species, category) %>%
  summarize(pc_t = mean(percent)) %>%
  filter(site == "Garden of Eden" & species == "Penicillus spp.")
test4 <- sg_pc_spp_site %>% filter(site == "Garden of Eden" & species == "Penicillus spp.")


```

#### Seagrasses - percent cover by species
```{r}
# graphs
ggplot(sg_pc_spp %>% 
         filter(category == "Seagrass"), 
       aes(x = reorder(species, pc_mean), y = pc_mean)) +
  geom_col(fill = "darkseagreen4", color = "black", alpha = 0.9) +
  geom_errorbar(aes(ymin = pc_mean - pc_se, ymax = pc_mean + pc_se), width = .2,
                 position = position_dodge(.9)) +
  coord_flip() +
  labs(y = "Mean percent cover", x = "Species", fill = "") +
  theme_bw() +
  theme(axis.text.y = element_text(face = "italic"))

kable(sg_pc_spp %>% filter(category == "Seagrass") %>% select(species, pc_mean, pc_se))

ggplot(sg_pc_spp_site %>% 
         filter(category == "Seagrass"), 
       aes(x = species, y = pc_mean)) +
  geom_col(fill = "darkseagreen4", color = "black", alpha = 0.9) +
  geom_errorbar(aes(ymin = pc_mean - pc_se, ymax = pc_mean + pc_se), width = .2,
                 position = position_dodge(.9)) +
  facet_wrap(. ~ site) +
  labs(y = "Mean percent cover", x = "Species", fill = "") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, h = 1, face = "italic"))

kable(sg_pc_spp_site %>% filter(category == "Seagrass") %>% select(site, species, pc_mean, pc_se))
```

#### Algae - percent cover by species/genus
```{r}
# graphs
ggplot(sg_pc_spp %>% 
         filter(category == "Algae"), 
       aes(x = reorder(species, pc_mean), y = pc_mean)) +
  geom_col(fill = "olivedrab", color = "black", alpha = 0.9) +
  geom_errorbar(aes(ymin = pc_mean - pc_se, ymax = pc_mean + pc_se), width = .2,
                 position = position_dodge(.9)) +
  coord_flip() +
  labs(y = "Mean percent cover", x = "Species", fill = "") +
  theme_bw() +
  theme(axis.text.y = element_text(face = "italic"))

kable(sg_pc_spp %>% filter(category == "Algae") %>% select(species, pc_mean, pc_se))

ggplot(sg_pc_spp_site %>% 
         filter(category == "Algae"), 
       aes(x = species, y = pc_mean)) +
  geom_col(fill = "olivedrab", color = "black", alpha = 0.9) +
  geom_errorbar(aes(ymin = pc_mean - pc_se, ymax = pc_mean + pc_se), width = .2,
                 position = position_dodge(.9)) +
  facet_wrap(. ~ site) +
  labs(y = "Mean percent cover", x = "Species", fill = "") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, h = 1, face = "italic"))

kable(sg_pc_spp_site %>% filter(category == "Algae") %>% select(site, species, pc_mean, pc_se))
```

#### Stony coral - percent cover by species
```{r}
# graphs

ggplot(sg_pc_spp %>% 
         filter(category == "Stony coral"), 
       aes(x = reorder(species, pc_mean), y = pc_mean)) +
  geom_col(fill = "coral2", color = "black", alpha = 0.9) +
  geom_errorbar(aes(ymin = pc_mean - pc_se, ymax = pc_mean + pc_se), width = .2,
                 position = position_dodge(.9)) +
  coord_flip() +
  labs(y = "Mean percent cover", x = "Species", fill = "") +
  theme_bw() +
  theme(axis.text.y = element_text(face = "italic"))

kable(sg_pc_spp %>% filter(category == "Stony coral") %>% select(species, pc_mean, pc_se))

ggplot(sg_pc_spp_site %>% 
         filter(category == "Stony coral"), 
       aes(x = species, y = pc_mean)) +
  geom_col(fill = "coral2", color = "black", alpha = 0.9) +
  geom_errorbar(aes(ymin = pc_mean - pc_se, ymax = pc_mean + pc_se), width = .2,
                 position = position_dodge(.9)) +
  facet_wrap(. ~ site) +
  labs(y = "Mean percent cover", x = "Species", fill = "") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, h = 1, face = "italic"))

kable(sg_pc_spp_site %>% filter(category == "Stony coral") %>% select(site, species, pc_mean, pc_se))
```

## Invertebrates
* Only inverts observed here were conch and hermit crab, so maybe just focus on conch for this analysis
```{r}
# import and clean data
sg_inverts <- read_excel(here("mma", "data", "mma_seagrass.xlsx"), sheet = "inverts") %>%  
   clean_names()
```

### Conch density by phase and site
```{r}
# expand data frame to include sites/transects where no conch were present
sg_conch <- sg_quads %>% # need to use the quads data to expand, because not all sites/transects are in invert data
  expand(nesting(site, transect)) %>%
  left_join(sg_inverts %>%
              select(site, transect, phase)) %>%
  expand(nesting(site, transect), phase) %>% # incorporating phase combinations
  filter(!is.na(phase)) %>%
  left_join(sg_inverts %>%
              filter(species == "Aliger gigas") %>%
              select(site, transect, phase, count)) %>%
  mutate(count = if_else(is.na(count), 0, count))

# calculate densities
sg_conch_dens <- sg_conch %>%
  group_by(site, transect, phase) %>%
  summarize(dens_t = sum(count)/20) %>% # indv/m2 (10m * 2m transects) 
  group_by(site, phase) %>%
  summarize(dens_mean = mean(dens_t),
            dens_se = sd(dens_t)/sqrt(n()))

# graph
ggplot(sg_conch_dens, aes(x = phase, y = dens_mean)) +
  geom_col(fill = "pink", color = "black", alpha = 0.9) +
  geom_errorbar(aes(ymin = dens_mean - dens_se, ymax = dens_mean + dens_se), width = .2,
                 position = position_dodge(.9)) +
  facet_wrap(. ~ site) +
  labs(y = "Mean density (indv./m2)", x = "Phase") +
  theme_bw()

kable(sg_conch_dens %>% select(phase, dens_mean, dens_se))
```

