---
title: "MMA Reef"
author: "Molly Wilson"
date: "6/22/2022"
output: 
  html_document:
    code_folding: hide
---

# {.tabset}

```{r, message = F, warning = F, echo = F}
library(tidyverse)
library(here) 
library(readxl) 
library(janitor)
library(snakecase) # for adjusting capitalization of text within data (e.g., species names)
library(knitr) # for including tables

knitr::opts_chunk$set(message = FALSE, warning = FALSE)
```

## Benthic

```{r}
# import and clean data

rf_benthic <- read_excel(here("mma", "data", "mma_coral_benthic.xlsx"), sheet = "benthic") %>%  
  clean_names() %>%
  filter(!is.na(site)) %>% # remove any incomplete rows at the end of the data
  # separate(species_code, c("species_code", "indicator")) %>% # separating variables with an underscore into two columns
  mutate(species = to_any_case(species, case = "sentence"), # this makes sure species names are correctly capitalized
         site = to_any_case(site, case = "title"),
         # consolidating categories here for graphs:
         cat_c = case_when(category_code %in% c("LC", "SLC") ~ "Hard corals",
                                  category_code %in% c("MA") ~ "Macroalgae",
                                  category_code %in% c("TA") ~ "Turf algae",
                                  category_code %in% c("CCA", "CCA_ND") ~ "CCA",
                                  category_code %in% c("OINV", "SPON", "AINV", "CYAN", "PEY") ~ "Other competitors",
                                  category_code %in% c("SAND", "HOLE", "SG", "PAVE") ~ "Other substratum"
                                    ),
         # adding algal type for graphs about palatability:
         algal_type = case_when(type_code %in% c("BFMA", "GFMA", "RFMA") ~ "Fleshy macroalgae",
                                type_code %in% c("GCMA", "RCMA") ~ "Calcareous macroalgae",
                                type_code %in% c("TA", "TAS", "STA") ~ "Turf algae"),
         # certain substrates are not suitable for coral or algal growth, so should not detract from percent cover
         av_sub_yn = if_else(category_code %in% c("SAND", "HOLE", "SG", "PAVE"), "no", "yes")
         ) %>%
    filter(!is.na(cat_c) & !is.na(species)) # need GRB to fix some data errors, but using this for now
```

### Percent cover by site and category

```{r}
# "expand" dataframe so that it contains all benthic categories per site/transect/meter -> join full dataset so that any meters where a species was absent it will show up as 0 (as opposed to having no entry)
rf_pc_cat_m <- rf_benthic %>%
  filter(av_sub_yn == "yes") %>% # percent cover is relative to available substrate
  expand(nesting(site, transect, meter), cat_c) %>%
  # this is where we add in our actual data to this expanded template:
  left_join(rf_benthic %>% 
              filter(av_sub_yn == "yes") %>% # only looking at what is considered available substrate (no sand, etc.)
              group_by(site, transect, meter) %>%
              mutate(n_pts = n()) %>% # showing total number of points per meter that are considered available substrate
              ungroup() %>%
              group_by(site, transect, meter, n_pts, cat_c) %>%
              summarize(pc_m = 100*n()/n_pts) %>% # n() counts the number of entries within a given group
              ungroup() %>%
              distinct() %>%
              select(-n_pts), 
            by = c("site", "transect", "meter", "cat_c")) %>%
  mutate(pc_m = if_else(is.na(pc_m), 0, pc_m))

# average these meter-level results within transect, then within sites
rf_pc_cat_site <- rf_pc_cat_m %>%
  group_by(site, transect, cat_c) %>%
  summarize(pc_t = mean(pc_m)) %>%
  ungroup() %>%
  group_by(site, cat_c) %>%
  summarize(n_test = n(),
            pc_mean = mean(pc_t),
            pc_se = sd(pc_t)/sqrt(n())
            ) %>%
  filter(cat_c != "Other substratum") %>% 
  mutate(cat_c = factor(cat_c, levels = c("Hard corals", "CCA", "Macroalgae", "Turf algae", "Other competitors")))

# quick check to make sure everything adds up to 100% at the meter and site level
test1 <- rf_pc_cat_m %>%
  group_by(site, transect, meter) %>%
  summarize(total = sum(pc_m))
test2 <- rf_pc_cat_site %>%
  group_by(site) %>%
  summarize(total = sum(pc_mean))
```


```{r}
cat_palette <- c("coral2", "pink", "darkolivegreen", "darkkhaki", "slategray3")
ggplot(rf_pc_cat_site, 
       aes(x = "cat_c", y = pc_mean, fill = cat_c)) + 
  geom_bar(width = 1, stat = "identity", color = "black") +
  coord_polar("y", start=0) +
  scale_fill_manual(values = cat_palette) +
  facet_wrap(vars(site), nrow = 2) +
  theme_void() +
  theme(legend.position = "bottom",
        legend.title = element_blank(),
        panel.spacing = unit(1, "lines"))

kable(rf_pc_cat_site %>% select(site, category = cat_c, pc_mean, pc_se))
ggsave(here("mma", "figs", "reef_benthic_pc.png"))
```


### Percent cover by species

```{r}
# calculating percent cover by species at each meter
rf_pc_spp_m <- rf_benthic %>%
  filter(av_sub_yn == "yes") %>%
  expand(nesting(site, transect, meter), nesting(species, cat_c)) %>%
  left_join(rf_benthic %>% 
              filter(av_sub_yn == "yes") %>% 
              group_by(site, transect, meter) %>%
              mutate(n_pts = n()) %>% 
              ungroup() %>%
              group_by(site, transect, meter, n_pts, species, cat_c) %>%
              summarize(pc_m = 100*n()/n_pts) %>%
              ungroup() %>%
              distinct() %>%
              select(-n_pts), 
            by = c("site", "transect", "meter", "species", "cat_c")) %>%
  mutate(pc_m = if_else(is.na(pc_m), 0, pc_m))

# calculating mean percent cover by transect -> site
rf_pc_spp_site <- rf_pc_spp_m %>%
  group_by(site, transect, species, cat_c) %>%
  summarize(pc_t = mean(pc_m)) %>%
  ungroup() %>%
  group_by(site, species, cat_c) %>%
  summarize(pc_mean = mean(pc_t),
            pc_se = sd(pc_t)/sqrt(n())
            ) %>%
  ungroup() %>%
  distinct()

# calculating mean percent cover by species across all sites
rf_pc_spp <- rf_pc_spp_site %>%
  group_by(species, cat_c) %>%
  summarize(pc = mean(pc_mean),
            pc_se = sd(pc_mean)/sqrt(n())) %>%
  rename(pc_mean = pc) 

# testing that total percent cover per transect and site add up to 100
test1 <- rf_pc_spp_m %>%
  group_by(site, transect, meter) %>%
  summarize(total = sum(pc_m))
test2 <- rf_pc_spp_site %>%
  group_by(site) %>%
  summarize(total = sum(pc_mean))
test3 <- sum(rf_pc_spp$pc_mean)
```

#### Stony coral - percent cover by species across all sites
```{r, fig.height = 2}
ggplot(rf_pc_spp %>%
         filter(cat_c == "Hard corals"),
       aes(x = reorder(species, pc_mean), y = pc_mean)) +
  geom_col(fill = "coral2", color = "black", alpha = 0.9) +
  geom_errorbar(aes(ymin = pc_mean - pc_se, ymax = pc_mean + pc_se), width = .2,
                 position = position_dodge(.9)) +
  coord_flip() +
  labs(y = "Mean percent cover", x = "", fill = "") +
  theme_bw() +
  theme(axis.text.y = element_text(face = "italic"))

kable(rf_pc_spp %>% filter(cat_c == "Hard corals") %>% select(species, pc_mean, pc_se))
ggsave(here("mma", "figs", "reef_coral_spp.png"))
```

### Macroalgal explorations

#### Macroalgae - percent cover by species and sites 
```{r, fig.height = 6}
ggplot(rf_pc_spp_site %>%
         filter(cat_c == "Macroalgae"),
       aes(x = reorder(species, pc_mean), y = pc_mean)) +
  geom_col(fill = "darkolivegreen", color = "black", alpha = 0.9) +
  geom_errorbar(aes(ymin = pc_mean - pc_se, ymax = pc_mean + pc_se), width = .2,
                 position = position_dodge(.9)) +
  facet_wrap(. ~ site, ncol = 3) +
  coord_flip() +
  labs(y = "Mean percent cover", x = "Species", fill = "") +
  theme_bw() +
  theme(axis.text.y = element_text(face = "italic"))

ggsave(here("mma", "figs", "ma_site.png"))
```

#### Canopy heights - all species
```{r}
ma_height <- rf_benthic %>%
  filter(category_code == "MA" & height > 0) %>%
  group_by(site, transect) %>%
  summarize(height = mean(height)) %>%
  filter(!is.na(height)) %>%
  group_by(site) %>%
  summarize(height_mean = mean(height),
            height_se = sd(height)/sqrt(n()))

ggplot(ma_height, aes(x = site, y = height_mean)) +
  geom_col(fill = "darkolivegreen", color = "black", alpha = 0.9) +
  geom_errorbar(aes(ymin = height_mean - height_se, ymax = height_mean + height_se), width = .2,
                 position = position_dodge(.9)) +
  labs(x = "", y = "Macroalgal canopy height (cm) (all spp)") +
  theme_bw()
ggsave(here("mma", "figs", "reef_ma_height.png"))

#kable(ma_height)
ggsave(here("mma", "figs", "ma_height_allspp.png"))
```

#### Canopy heights - dictyota only
```{r}
ma_height_palatable <- rf_benthic %>%
  filter(species_code == "DICT" & height > 0) %>%
  group_by(site, transect) %>%
  summarize(height = mean(height)) %>%
  filter(!is.na(height)) %>%
  group_by(site) %>%
  summarize(height_mean = mean(height),
            height_se = sd(height)/sqrt(n()))

ggplot(ma_height_palatable, aes(x = site, y = height_mean)) +
  geom_col(fill = "darkolivegreen", color = "black", alpha = 0.9) +
  geom_errorbar(aes(ymin = height_mean - height_se, ymax = height_mean + height_se), width = .2,
                 position = position_dodge(.9)) +
  labs(x = "", y = "Macroalgal canopy height (cm) (dict only)") +
  theme_bw()
ggsave(here("mma", "figs", "reef_ma_height.png"))

#kable(ma_height)
ggsave(here("mma", "figs", "ma_height_dict.png"))
```

#### Macroalgal index - all species
```{r}
ma_index <- rf_pc_cat_site %>%
  filter(cat_c == "Macroalgae") %>%
  left_join(ma_height) %>%
  mutate(index = height_mean * pc_mean,
         index = if_else(is.na(index), 0, index))
  
ggplot(ma_index, aes(x = site, y = index)) +
  geom_col(fill = "darkolivegreen", color = "black", alpha = 0.9) +
  labs(x = "", y = "Macroalgal index (all spp)") +
  theme_bw()

#kable(ma_index %>% select(site, index))
ggsave(here("mma", "figs", "ma_index_allspp.png"))
```
#### Macroalgal index - dictyota only
```{r}
# calculating average canopy height by species at each transect
ma_height_dict <- rf_benthic %>%
  filter(species_code %in% c("DICT", "LOBO") & height > 0) %>%
  group_by(site, transect) %>%
  summarize(height = mean(height)) %>%
  filter(!is.na(height)) %>%
  group_by(site) %>%
  summarize(height_mean = mean(height),
            height_se = sd(height)/sqrt(n()))

ma_index_dict <- rf_pc_spp_site %>%
  filter(species == "Dictyota spp") %>%
  left_join(ma_height_dict) %>%
  mutate(index = height_mean * pc_mean,
         index = if_else(is.na(index), 0, index))

ggplot(ma_index_dict, aes(x = site, y = index)) +
  geom_col(fill = "darkolivegreen", color = "black", alpha = 0.9) +
  labs(x = "", y = "Macroalgal index (Dictyota spp.)") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
ggsave(here("mma", "figs", "ma_index_dict.png"))
```

```{r}
# ma_index <- rf_pc_spp_site %>%
#   filter(species == "Dictyota spp") %>%
#   left_join(ma_height) %>%
#   mutate(index = height_mean * pc_mean,
#          index = if_else(is.na(index), 0, index))
#   
# ggplot(ma_index, aes(x = site, y = index)) +
#   geom_col(fill = "darkolivegreen", color = "black", alpha = 0.9) +
#   labs(x = "", y = "Macroalgal index") +
#   theme_bw()
# 
# kable(ma_index %>% select(site, index))
```

### Coral recruits
```{r}
rf_recruits <- read_excel(here("mma", "data", "mma_coral_benthic.xlsx"), sheet = "recruits") %>%  
  clean_names() %>%
  pivot_longer(cols = starts_with("count"),
               names_to = "size",
               names_pattern = "count_(.*)",
               values_to = "count") %>%
  mutate(size = if_else(size == "0_2cm", "<2 cm", "2-4 cm")) %>%
  filter(species != "#N/A") %>%
  select(site, transect, meter, size, code, species, count)
```

#### Density of coral recruits (all species) by site
```{r, eval = F}
# by size
rf_recruits_size <- rf_recruits %>%
  group_by(site, transect, size, meter) %>%
  summarize(count = sum(count)) %>%
  group_by(site, transect, size) %>%
  summarize(density = mean(count/(5 * 0.25 * 0.25))) %>% # 5 25cm x 25cm quadrats per transect
  group_by(site, size) %>%
  summarize(density_mean = mean(density),
            density_se = sd(density)/sqrt(n())) # se = mean here, which I assume just means there is a lot of variation/not many samples

ggplot(rf_recruits_size, aes(x = site, y = density_mean)) +
  geom_col(fill = "coral2", color = "black", alpha = 0.9) +
  geom_errorbar(aes(ymin = density_mean - density_se, ymax = density_mean + density_se), width = .2,
                 position = position_dodge(.9)) +
  labs(x = "", y = expression(Larval~recruit~density~(recruits~m^-2))) +
  facet_wrap(. ~ size) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, h = 1))

kable(rf_recruits_size)
```

```{r}
# grouped across sizes
rf_recruits_tot <- rf_recruits %>%
  group_by(site, transect, meter) %>%
  summarize(count = sum(count)) %>%
  group_by(site, transect) %>%
  summarize(density = mean(count/(5 * 0.25 * 0.25))) %>% 
  group_by(site) %>%
  summarize(density_mean = mean(density),
            density_se = sd(density)/sqrt(n())) # se = mean here, which I assume just means there is a lot of variation/not many samples

ggplot(rf_recruits_tot, aes(x = site, y = density_mean)) +
  geom_col(fill = "coral2", color = "black", alpha = 0.9) +
  geom_errorbar(aes(ymin = density_mean - density_se, ymax = density_mean + density_se), width = .2,
                 position = position_dodge(.9)) +
  labs(x = "", y = expression(Larval~recruit~density~(recruits~m^-2))) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
ggsave(here("mma", "figs", "reef_recruits.png"))

kable(rf_recruits_tot)
ggsave(here("mma", "figs", "reef_recruits.png"))
```

#### Proportion of coral recruits by species
```{r}
rf_recruits_spp <- rf_recruits %>%
  filter(count > 0) %>%
  uncount(count) %>%
  group_by(size, species) %>%
  summarize(n = n()) %>%
  mutate(freq = n / sum(n))

ggplot(rf_recruits_spp, aes(x = species, y = freq)) +
  geom_col(fill = "coral2", color = "black", alpha = 0.9) +
  facet_wrap(. ~ size) +
  labs(x = "", y = "Fraction of recruits") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, h = 1, face = "italic"))
ggsave(here("mma", "figs", "reef_recruits_spp.png"))
```

### Diadema

```{r, fig.height = 2, fig.width = 3}
diadema <- read_excel(here("mma", "data", "mma_coral_benthic.xlsx"), sheet = "inverts") %>%  
  clean_names() %>%
  select(site, transect, diadema_adult, diadema_juvenile) %>%
  pivot_longer(cols = starts_with("diadema"),
               names_to = "phase",
               names_pattern = "diadema_(.*)",
               values_to = "count")

diadema_site <- diadema %>%
  group_by(site,) %>%
  summarize(density_mean = mean(count/10), # 1x10m belt
            density_se = sd(count/10)/sqrt(n())) 
  
diadema_site_ph <- diadema %>%
  group_by(site, phase) %>%
  summarize(density_mean = mean(count/10), # 1x10m belt
            density_se = sd(count/10)/sqrt(n())) 

rect_dat_diadema = data.frame(xmin = -Inf, xmax = Inf, ymin = 3, ymax = 20) # (3-20 indv/m2, Bak 1985)
ggplot(diadema_site, aes(x = site, y = density_mean)) +
  geom_col(fill = "black", position = "dodge") +
  scale_fill_manual(values = c("gray10", "gray90"), name = "Phase", labels = c("Adult", "Juvenile")) +
  geom_errorbar(aes(ymin = density_mean - density_se, ymax = density_mean + density_se), width = .2,
                 position = position_dodge(.9)) +
  geom_rect(data = rect_dat_diadema, inherit.aes = FALSE,
                              aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax), 
                              fill = 'slategray1', alpha = 0.5) +
  labs(x = "", y = expression(Density~(indv.~m^-2))) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

ggsave(here("mma", "figs", "reef_diadema.png"))
```

```{r}
diadema_sum <- diadema %>%
  group_by(site, transect) %>%
  summarize(density = sum(count)/10) %>%
  group_by(site) %>%
  summarize(dens_mean = mean(density))
  
diadema_mean <- tibble(dens_mean = mean(diadema_sum$dens_mean),
                       dens_se = sd(diadema_sum$dens_mean)/sqrt(4))
kable(diadema_mean)
```


## Fish

```{r}
# import and clean data

rf_fish <- read_excel(here("mma", "data", "mma_coral_fish.xlsx"), sheet = "fish") %>%  
  clean_names() %>%
  filter(!is.na(site)) %>% # remove any incomplete rows at the end of the data
  mutate(number = if_else(is.na(number), 1, number)) %>% # all entries with no number specified were single observations
  uncount(number) %>% # expand to replicate rows if multiple fish were recorded to look at length distributions
  mutate(biomass = as.numeric(biomass),
         phase_code = tolower(phase),
         phase = case_when(phase == "j" ~ "Juvenile",
                           phase == "i" ~ "Initial",
                           phase == "t" ~ "Terminal"),
         family_c = if_else(family %in% c("Scaridae", "Acanthuridae", "Haemulidae", "Serranidae", "Lutjanidae", "Balistidae", "Pomacentridae"), family, "Other"),
         family_com = case_when(family_c == "Scaridae" ~ "Parrotfish",
                                family_c == "Acanthuridae" ~ "Surgeonfish",
                                family_c == "Haemulidae" ~ "Grunts",
                                family_c == "Serranidae" ~ "Grouper",
                                family_c == "Lutjanidae" ~ "Snapper",
                                family_c == "Balistidae" ~ "Triggerfish",
                                family_c == "Pomacentridae" ~ "Damselfish",
                                family_c == "Other" ~ "Other")
         ) # consolidating the number of families

n_transects <- rf_fish %>%
  select(site, transect) %>%
  distinct() %>%
  group_by(site) %>%
  summarize(n = n())
```

### Total biomass by site
* Regional averages are taken from Karr et al. 2015
* Antigua averages are taken from Camacho et al. 2020 (5921 g/m2)
```{r, fig.height = 2, fig.width = 3}
belt_area <- 30*2 # 30m x 2m belt transects

# calculating total biomass in each transect (sum) -> mean at each site
rf_fish_site <- rf_fish %>% 
  group_by(site, transect) %>%
  summarize(bm_tot = sum(biomass)/1000/belt_area*10000) %>% #kg/ha
  group_by(site) %>%
  summarize(bm_mean = mean(bm_tot),
            bm_se = sd(bm_tot)/sqrt(n()))

rf_bm_sum <- rf_fish_site %>%
  summarize(bm_mean_sum = mean(bm_mean),
            bm_se_sum = sd(bm_mean)/sqrt(n()))

# graph
ggplot(rf_fish_site, aes(x = site, y = bm_mean)) +
  geom_col(fill = "cadetblue3", color = "black", alpha = 0.8) +
  geom_errorbar(aes(ymin = bm_mean - bm_se, ymax = bm_mean + bm_se), width = .2,
                 position = position_dodge(.9)) +
  geom_hline(yintercept = 1300, linetype = "dashed", color = "black") +
  annotate("text", x = 8.5, y = 1350, size = 3, hjust = 1, label=c('Caribbean mean unfished biomass')) +
  geom_hline(yintercept = 591, linetype = "dashed", color = "black") +
  annotate("text", x = 8.5, y = 640, size = 3, hjust = 1, label=c('Antigua mean biomass')) +
  labs(x = "", y = expression(Total~fish~biomass~(kg~ha^-1))) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

kable(rf_fish_site %>% select(site, bm_mean, bm_se))
kable(rf_bm_sum)
ggsave(here("mma", "figs", "reef_fish_bm.png"))
```

### Biomass by family
```{r}
rf_fish_fam_site <- rf_fish %>% 
  expand(nesting(site, transect), family_com, family_c) %>%
  left_join(rf_fish %>%
      select(site, transect, family_com, family_c, length, biomass)) %>%
  mutate(biomass = if_else(is.na(biomass), 0, biomass)) %>%
  group_by(site, transect, family_com, family_c) %>%
  summarize(bm_tot = sum(biomass)/1000/belt_area*10000) %>% #kg/ha
  group_by(site, family_com, family_c) %>%
  summarize(bm_mean = mean(bm_tot),
            bm_se = sd(bm_tot)/sqrt(n()))
```


```{r}
rf_fish_fam <- rf_fish %>% 
  expand(nesting(site, transect), family_com) %>%
  left_join(rf_fish %>%
      select(site, transect, family_com, length, biomass)) %>%
  mutate(biomass = if_else(is.na(biomass), 0, biomass)) %>%
  group_by(site, transect, family_com) %>%
  summarize(bm_tot = sum(biomass)/1000/belt_area*10000) %>% #kg/ha
  group_by(site, family_com) %>%
  summarize(bm_site = mean(bm_tot)) %>%
  group_by(family_com) %>%
  summarize(bm_mean = mean(bm_site),
            bm_se = sd(bm_site)/sqrt(n())) %>%
  mutate(family_com = factor(family_com, levels = c("Damselfish", "Grouper", "Grunts", "Parrotfish", "Snapper", "Surgeonfish", "Triggerfish", "Other")))
```


### Herbivore biomasses
* Combining scarid and acanthurids
* Antigua scarid biomass (1,189 g/100 m2) and acanthurid biomass (1,369 g/100 m2) _> sum = 2558 g/100 m2 taken from Camacho et al. 2020
```{r, fig.width = 3, fig.height = 2}
rf_fish_herb <- rf_fish %>% 
  filter(family %in% c("Scaridae", "Acanthuridae")) %>%
  mutate(biomass = if_else(is.na(biomass), 0, biomass)) %>%
  group_by(site, transect) %>%
  summarize(bm_t = sum(biomass)/1000/belt_area*10000) %>% #kg/ha
  group_by(site) %>%
  summarize(bm_mean = mean(bm_t),
            bm_se = sd(bm_t)/sqrt(n()))

ggplot(rf_fish_herb, aes(x = site, y = bm_mean)) +
  geom_col(color = "black", fill = "cadetblue3", alpha = 0.8, stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = bm_mean - bm_se, ymax = bm_mean + bm_se), width = .2, color = "gray40",
                 position = position_dodge(.9)) +
  # geom_hline(yintercept = 256, linetype = "dashed", color = "black") +
  # annotate("text", x = 8.5, y = 258, size = 3, hjust = 1, label=c('Antigua mean biomass')) +
  labs(x = "", y = expression(Herbivore~biomass~(kg~ha^-1))) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
ggsave(here("mma", "figs", "reef_herb_bm.png"))
```


### Biomass by species across all sites

```{r, fig.height = 5, fig.width = 3}
# calculate mean biomass by species for each transect -> site -> mean across sites
rf_fish_spp <- rf_fish %>% 
  expand(nesting(site, transect), species_code) %>%
  left_join(rf_fish %>%
      select(site, transect, species_code, length, biomass)) %>%
  mutate(biomass = if_else(is.na(biomass), 0, biomass)) %>%
  group_by(site, transect, species_code) %>%
  summarize(bm_tot = sum(biomass)/1000/belt_area*10000) %>% #kg/ha
  group_by(site, species_code) %>%
  summarize(bm_mean = mean(bm_tot),
            bm_se = sd(bm_tot)/sqrt(n())) %>%
  group_by(species_code) %>%
  summarize(bm = mean(bm_mean),
            bm_se = sd(bm_mean)/sqrt(n())) %>%
  rename(bm_mean = bm) %>%
  left_join(rf_fish %>% 
              select(species_code, species_name, common_name, functional_group) %>%
              distinct())

# graph
ggplot(rf_fish_spp, aes(x = reorder(species_name, bm_mean), y = bm_mean)) +
  geom_col(fill = "cadetblue3", color = "black", alpha = 0.8) +
  geom_errorbar(aes(ymin = bm_mean - bm_se, ymax = bm_mean + bm_se), width = .2,
                 position = position_dodge(.9)) +
  coord_flip() +
  labs(x = "Species", y = expression(Biomass~(kg~ha^-1))) +
  theme_bw() +
  theme(axis.text.y = element_text(face = "italic"))

ggsave(here("mma", "figs", "reef_fish_spp.png"))
```


### Fish lengths

#### Violin plots with pooled data from each site
* Each site had 5 transects except for the Barge which had 4... so pooling all fish observed across all transects per site to make violins work
```{r}
ggplot(rf_fish, aes(x = site, y = length)) +
  geom_violin(fill = "cadetblue3", alpha = 0.8) +
  labs(x = "", y = "Fish length (cm)") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
ggsave(here("mma", "figs", "reef_fish_violin.png"))
```

#### Parrotfish lengths relative to sex change
- Minimum sizes at maturity from O'Farrell et al. 2015:
      - Sp. viride, 18 cm (Winn and Bardach 1960);
      - Sp. aurofrenatum, 15 cm (Garcı´a-Cagide et al. 1994); 
      - Sc. vetula, 19 cm (Winn and Bardach 1960); 
      - and Sc. taeniopterus, 19 cm (Munro 1983).
      
```{r}
scarids <- rf_fish %>%
  filter(species_code %in% c("stop", "rbp", "qup", "strp") & !is.na(phase)) %>%
  mutate(sex = case_when(phase %in% c("Juvenile", "Initial") ~ "Female",
                         phase == "Terminal" ~ "Male"))

ggplot(scarids, aes(x = length, fill = sex)) +
  geom_histogram(stat = "count", alpha = 0.8, position = 'identity') +
  facet_wrap(vars(species_name), ncol = 4) +
  geom_vline(data = filter(scarids, species_code == "stop"), aes(xintercept = 18, color = "Minimum length at maturity"), linetype = "dashed") +
  geom_vline(data = filter(scarids, species_code == "qup"), aes(xintercept = 19, color = "Minimum length at maturity"), linetype = "dashed") +
  geom_vline(data = filter(scarids, species_code == "rbp"), aes(xintercept = 15, color = "Minimum length at maturity"), linetype = "dashed") +
  geom_vline(data = filter(scarids, species_code == "strp"), aes(xintercept = 19, color = "Minimum length at maturity"), linetype = "dashed") +
  scale_color_manual(values = c("black")) +
  scale_fill_manual(values=c("pink2", "turquoise3")) +
  theme_bw() +
  labs(x = "Fish length (cm)", y = "Number observed", fill="") +
  theme(strip.text = element_text(face = "italic"),
        legend.title = element_blank(),
        legend.position = "top")
ggsave(here("mma", "figs", "reef_scarids_phase.png"))
```

