---
title: "JB Baseline - Reefs"
author: "Molly Wilson"
date: "8/14/2022"
output:
  pdf_document: default
  html_document:
    code_folding: hide
---

# {.tabset}

```{r, message = F, warning = F, echo = F}
library(tidyverse)
library(here) 
library(readxl) 
library(janitor)
library(snakecase) # for adjusting capitalization of text within data (e.g., species names)
library(knitr) # for including tables

knitr::opts_chunk$set(message = FALSE, warning = FALSE)
```

## Benthic
```{r}
# import and clean data

rf_benthic <- read_excel(here("jumby_baseline", "data", "JBB_reefs.xlsx"), sheet = "benthic") %>%  
  clean_names() %>%
  filter(!is.na(site)) %>% # remove any incomplete rows at the end of the data
  # separate(species_code, c("species_code", "indicator")) %>% # separating variables with an underscore into two columns
  mutate(species = to_any_case(species, case = "sentence"), # this makes sure species names are correctly capitalized
         # site = to_any_case(site, case = "title"),
         # consolidating categories here for graphs:
         cat_c = case_when(category_code %in% c("LC", "SLC") ~ "Hard corals",
                                  category_code %in% c("MA") ~ "Macroalgae",
                                  category_code %in% c("TA") ~ "Turf algae",
                                  category_code %in% c("CCA", "CCA_ND") ~ "CCA",
                                  category_code %in% c("OINV", "SPON", "AINV", "CYAN", "PEY") ~ "Other competitors",
                                  category_code %in% c("SAND", "HOLE", "SG", "PAVE") ~ "Other substratum"
                                    ),
         # adding algal type for graphs about palatability:
         algal_type = case_when(type_code %in% c("BFMA", "GFMA", "RFMA") ~ "Fleshy macroalgae",
                                type_code %in% c("GCMA", "RCMA") ~ "Calcareous macroalgae",
                                type_code %in% c("TA", "TAS", "STA") ~ "Turf algae"),
         # certain substrates are not suitable for coral or algal growth, so should not detract from percent cover
         av_sub_yn = if_else(category_code %in% c("SAND", "HOLE", "SG", "PAVE"), "no", "yes")
         ) %>%
    filter(!is.na(cat_c) & !is.na(species)) # need GRB to fix some data errors, but using this for now
```

### Percent cover 

#### Percent cover by site and category
```{r}
# "expand" dataframe so that it contains all benthic categories per site/transect/meter -> join full dataset so that any meters where a species was absent it will show up as 0 (as opposed to having no entry)
rf_pc_cat_m <- rf_benthic %>%
  filter(av_sub_yn == "yes") %>% # percent cover is relative to available substrate
  expand(nesting(site, transect, meter), cat_c) %>%
  # this is where we add in our actual data to this expanded template:
  left_join(rf_benthic %>% 
              filter(av_sub_yn == "yes") %>% # only looking at what is considered available substrate (no sand, etc.)
              group_by(site, transect, meter) %>%
              mutate(n_pts = n()) %>% # showing total number of points per meter that are considered available substrate
              ungroup() %>%
              group_by(site, transect, meter, n_pts, cat_c) %>%
              summarize(pc_m = 100*n()/n_pts) %>% # n() counts the number of entries within a given group
              ungroup() %>%
              distinct() %>%
              select(-n_pts), 
            by = c("site", "transect", "meter", "cat_c")) %>%
  mutate(pc_m = if_else(is.na(pc_m), 0, pc_m))

# average these meter-level results within transect, then within sites
rf_pc_cat_site <- rf_pc_cat_m %>%
  group_by(site, transect, cat_c) %>%
  summarize(pc_t = mean(pc_m)) %>%
  ungroup() %>%
  group_by(site, cat_c) %>%
  summarize(n_test = n(),
            pc_mean = mean(pc_t),
            pc_se = sd(pc_t)/sqrt(n())
            ) %>%
  filter(cat_c != "Other substratum") %>% 
  mutate(cat_c = factor(cat_c, levels = c("Hard corals", "CCA", "Macroalgae", "Turf algae", "Other competitors")))

# quick check to make sure everything adds up to 100% at the meter and site level
test1 <- rf_pc_cat_m %>%
  group_by(site, transect, meter) %>%
  summarize(total = sum(pc_m))
test2 <- rf_pc_cat_site %>%
  group_by(site) %>%
  summarize(total = sum(pc_mean))
```


```{r, eval = F}
cat_palette <- c("coral2", "pink", "darkolivegreen", "darkkhaki", "slategray3")
ggplot(rf_pc_cat_site, 
       aes(x = cat_c, y = pc_mean, fill = cat_c)) +
  geom_col(color = "black", alpha = 0.9) +
  geom_errorbar(aes(ymin = pc_mean - pc_se, ymax = pc_mean + pc_se), width = .2,
                 position = position_dodge(.9)) +
  scale_fill_manual(values = cat_palette) +
  facet_wrap(. ~ site, ncol = 2) +
  labs(y = "Mean percent cover", x = "", fill = "") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, h = 1, face = "italic"))
```


```{r}
cat_palette <- c("coral2", "pink", "darkolivegreen", "darkkhaki", "slategray3")
ggplot(rf_pc_cat_site, 
       aes(x = "cat_c", y = pc_mean, fill = cat_c)) + 
  geom_bar(width = 1, stat = "identity", color = "black") +
  coord_polar("y", start=0) +
  scale_fill_manual(values = cat_palette) +
  facet_wrap(vars(site), nrow = 2) +
  theme_void() +
  theme(legend.position = "bottom",
        legend.title = element_blank(),
        panel.spacing = unit(1, "lines"))

kable(rf_pc_cat_site %>% select(site, category = cat_c, pc_mean, pc_se))
```


### Percent cover by species
```{r}
# calculating percent cover by species at each meter
rf_pc_spp_m <- rf_benthic %>%
  filter(av_sub_yn == "yes") %>%
  expand(nesting(site, transect, meter), nesting(species, cat_c)) %>%
  left_join(rf_benthic %>% 
              filter(av_sub_yn == "yes") %>% 
              group_by(site, transect, meter) %>%
              mutate(n_pts = n()) %>% 
              ungroup() %>%
              group_by(site, transect, meter, n_pts, species, cat_c) %>%
              summarize(pc_m = 100*n()/n_pts) %>%
              ungroup() %>%
              distinct() %>%
              select(-n_pts), 
            by = c("site", "transect", "meter", "species", "cat_c")) %>%
  mutate(pc_m = if_else(is.na(pc_m), 0, pc_m))

# calculating mean percent cover by transect -> site
rf_pc_spp_site <- rf_pc_spp_m %>%
  group_by(site, transect, species, cat_c) %>%
  summarize(pc_t = mean(pc_m)) %>%
  ungroup() %>%
  group_by(site, species, cat_c) %>%
  summarize(pc_mean = mean(pc_t),
            pc_se = sd(pc_t)/sqrt(n())
            ) %>%
  ungroup() %>%
  distinct()

# calculating mean percent cover by species across all sites
rf_pc_spp <- rf_pc_spp_site %>%
  group_by(species, cat_c) %>%
  summarize(pc = mean(pc_mean),
            pc_se = sd(pc_mean)/sqrt(n())) %>%
  rename(pc_mean = pc) 

# testing that total percent cover per transect and site add up to 100
test1 <- rf_pc_spp_m %>%
  group_by(site, transect, meter) %>%
  summarize(total = sum(pc_m))
test2 <- rf_pc_spp_site %>%
  group_by(site) %>%
  summarize(total = sum(pc_mean))
test3 <- sum(rf_pc_spp$pc_mean)
```

#### Stony coral - percent cover by species across all sites
```{r}
ggplot(rf_pc_spp %>%
         filter(cat_c == "Hard corals"),
       aes(x = reorder(species, pc_mean), y = pc_mean)) +
  geom_col(fill = "coral2", color = "black", alpha = 0.9) +
  geom_errorbar(aes(ymin = pc_mean - pc_se, ymax = pc_mean + pc_se), width = .2,
                 position = position_dodge(.9)) +
  coord_flip() +
  labs(y = "Mean percent cover", x = "", fill = "") +
  theme_bw() +
  theme(axis.text.y = element_text(face = "italic"))

kable(rf_pc_spp %>% filter(cat_c == "Hard corals") %>% select(species, pc_mean, pc_se))
```

### Macroalgal canopy heights
* No macroalgae present at Pierce Shoals W
* Macroalgae only present on one transect at Pasture Point, hence no standard error can be calculated
```{r}
ma_height <- rf_benthic %>%
  filter(category_code == "MA") %>%
  group_by(site, transect) %>%
  summarize(height = mean(height)) %>%
  filter(!is.na(height)) %>%
  group_by(site) %>%
  summarize(height_mean = mean(height),
            height_se = sd(height)/sqrt(n()))

ggplot(ma_height, aes(x = site, y = height_mean)) +
  geom_col(fill = "darkolivegreen", color = "black", alpha = 0.9) +
  geom_errorbar(aes(ymin = height_mean - height_se, ymax = height_mean + height_se), width = .2,
                 position = position_dodge(.9)) +
  labs(x = "", y = "Macroalgal canopy height (cm)") +
  theme_bw()
```

### Need to add:
* recruits
* urchins

## Fish

```{r}
# import and clean data

rf_fish <- read_excel(here("jumby_baseline", "data", "JBB_reefs.xlsx"), sheet = "fish") %>%  
  clean_names() %>%
  filter(!is.na(site)) %>% # remove any incomplete rows at the end of the data
  mutate(number = if_else(is.na(number), 1, number)) %>% # all entries with no number specified were single observations
  uncount(number) %>% # expand to replicate rows if multiple fish were recorded to look at length distributions
  mutate(biomass = as.numeric(biomass),
         phase_code = tolower(phase),
         phase = case_when(phase == "j" ~ "Juvenile",
                           phase == "i" ~ "Initial",
                           phase == "t" ~ "Terminal"),
         family_c = if_else(family %in% c("Scaridae", "Acanthuridae", "Haemulidae", "Serranidae", "Lutjanidae", "Balistidae", "Pomacentridae"), family, "Other")
         ) # consolidating the number of families
```

### Fish biomass

#### Total biomass at each site
* Regional averages are taken from Karr et al. 2015
```{r}
# calculating total biomass in each transect (sum) -> mean at each site
rf_fish_site <- rf_fish %>% 
  group_by(site, transect) %>%
  summarize(bm_tot = sum(biomass)/1000/120*10000) %>% #kg/ha
  group_by(site) %>%
  summarize(bm_mean = mean(bm_tot),
            bm_se = sd(bm_tot)/sqrt(n()))
```


```{r}
# graph
ggplot(rf_fish_site, aes(x = site, y = bm_mean)) +
  geom_col(fill = "cadetblue3", color = "black", alpha = 0.8) +
  geom_errorbar(aes(ymin = bm_mean - bm_se, ymax = bm_mean + bm_se), width = .2,
                 position = position_dodge(.9)) +
  geom_hline(yintercept = 1300, linetype = "dashed", color = "black") +
  annotate("text", x = 5.5, y = 1350, size = 3, hjust = 1, label=c('Caribbean mean unfished biomass')) +
  labs(x = "", y = expression(Total~fish~biomass~(kg~ha^-1))) +
  theme_bw()

kable(rf_fish_site %>% select(site, bm_mean, bm_se))
```

### Biomass by site and family
* Need to reorder to put "other" last
* Are these good for families to feature?
* Should I switch to common family names, or both?
```{r}
rf_fish_fam_site <- rf_fish %>% 
  expand(nesting(site, transect), family_c) %>%
  left_join(rf_fish %>%
      select(site, transect, family_c, length, biomass)) %>%
  mutate(biomass = if_else(is.na(biomass), 0, biomass)) %>%
  group_by(site, transect, family_c) %>%
  summarize(bm_tot = sum(biomass)/1000/120*10000) %>% #kg/ha
  group_by(site, family_c) %>%
  summarize(bm_mean = mean(bm_tot),
            bm_se = sd(bm_tot)/sqrt(n()))
```


```{r, fig.height = 7, fig.width = 4}
# graph
ggplot(rf_fish_fam_site, aes(x = family_c, y = bm_mean)) +
  geom_col(fill = "cadetblue3", color = "black", alpha = 0.9) +
  geom_errorbar(aes(ymin = bm_mean - bm_se, ymax = bm_mean + bm_se), width = .2,
                 position = position_dodge(.9)) +
  facet_wrap(. ~ site, ncol = 1) +
  labs(x = "", y = expression(Biomass~(kg~ha^-1))) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, h = 1))

kable(rf_fish_fam_site %>% select(site, family = family_c, bm_mean, bm_se))
```

### Herbivore biomasses

#### Scarid biomass by site
*  - Eastern Caribbean average parrotfish biomass for NTRs from Steneck et al. 2018 was ~ 1550 g/100m2, or 155 kg/ha. Note that this is potentially misleading because I'm sure a lot of these NTRs do experience some fishing. But mean fished parrotfish biomasses for the same region was ~ 750 g/100m2
```{r}
ggplot(rf_fish_fam_site %>%
        filter(family_c == "Scaridae"),
       aes(x = site, y = bm_mean)) +
  geom_col(color = "black", fill = "cadetblue3", alpha = 0.8, stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = bm_mean - bm_se, ymax = bm_mean + bm_se), width = .2,
                 position = position_dodge(.9)) +
  geom_hline(yintercept = 155, linetype = "dashed", color = "black") +
  annotate("text", x = 4.9, y = 165, size = 3, hjust = 1, label=c('Eastern Caribbean mean unfished biomass')) +
  labs(x = "", y = expression(Parrotfish~biomass~(kg~ha^-1))) +
  theme_bw()
```

#### Acanthurid biomass by site
```{r}
ggplot(rf_fish_fam_site %>%
        filter(family_c == "Acanthuridae"),
       aes(x = site, y = bm_mean)) +
  geom_col(color = "black", fill = "cadetblue3", alpha = 0.8, stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = bm_mean - bm_se, ymax = bm_mean + bm_se), width = .2,
                 position = position_dodge(.9)) +
  labs(x = "", y = expression(Surgeonfish~biomass~(kg~ha^-1))) +
  theme_bw()
```

### Fish lengths
* These were calculated by taking the mean density of fish of each size at each size (averaged across transects). Because we have distribution data within each transect, we can't make density plots/violin plots straight from the raw data (i.e., some sites could have more transects)
* Open to any suggestions about how we want to communicate size data!

#### Length distributions of all fish by site
```{r}
rf_fish_length <- rf_fish %>%
  expand(nesting(site, transect), length) %>%
  left_join(rf_fish %>%
        group_by(site, transect, length) %>%
        summarize(count = n())) %>%
  mutate(count = as.numeric(count),
         count = if_else(is.na(count), 0, count),
         density = count/120) %>% # indv./m2
  group_by(site, length) %>%
  summarize(density_mean = mean(density),
            density_se = sd(density)/sqrt(n()))

ggplot(rf_fish_length, aes(x = length, y = density_mean)) +
  geom_col(fill = "deepskyblue4", alpha = 0.9, color = "black") +
  facet_wrap(. ~ site) +
  labs(x = "Fish length (cm)", y = expression(Density~(indv.~m^-2))) +
  theme_bw()
```

#### Length distributions of scarids at each site

```{r}
rf_fish_length_scarids <- rf_fish %>%
  expand(nesting(site, transect), length) %>%
  left_join(rf_fish %>%
        filter(family == "Scaridae") %>%
        group_by(site, transect, length) %>%
        summarize(count = n())) %>%
  mutate(count = as.numeric(count),
         count = if_else(is.na(count), 0, count),
         density = count/120) %>% # invd/m2
  group_by(site, length) %>%
  summarize(density_mean = mean(density),
            density_se = sd(density)/sqrt(n()))

ggplot(rf_fish_length_scarids, aes(x = length, y = density_mean)) +
  geom_col(fill = "deepskyblue4", alpha = 0.9, color = "black") +
  facet_wrap(. ~ site) +
  labs(x = "Fish length (cm)", y = expression(Scarid~density~(indv.~m^-2))) +
  theme_bw()
```

