---
title: "Jumby Bay Baseline Assessment"
author: "Molly Wilson"
date: "8/14/2022"
output: 
  html_document:
    code_folding: hide
---

# {.tabset}

```{r, message = F, warning = F, echo = F}
library(tidyverse)
library(here) 
library(readxl) 
library(janitor)
library(snakecase) # for adjusting capitalization of text within data (e.g., species names)
library(knitr) # for including tables

knitr::opts_chunk$set(message = FALSE, warning = FALSE)
```

## Benthic
# need to update autofills on original sheets
```{r, run = F}
# import and clean data

rf_benthic <- read_excel(here("jumby_baseline", "data", "JB_reefs.xlsx"), sheet = "benthic") %>%  
  clean_names() %>%
  filter(!is.na(site)) %>% # remove any incomplete rows at the end of the data
  # separate(species_code, c("species_code", "indicator")) %>% # separating variables with an underscore into two columns
  mutate(species = to_any_case(species, case = "sentence"), # this makes sure species names are correctly capitalized
         site = to_any_case(site, case = "title"),
         # consolidating categories here for graphs:
         cat_c = case_when(category_code %in% c("LC", "SLC") ~ "Hard corals",
                                  category_code %in% c("MA") ~ "Macroalgae",
                                  category_code %in% c("TA") ~ "Turf algae",
                                  category_code %in% c("CCA", "CCA_ND") ~ "CCA",
                                  category_code %in% c("OINV", "SPON", "AINV", "CYAN", "PEY") ~ "Other competitors",
                                  category_code %in% c("SAND", "HOLE", "SG", "PAVE") ~ "Other substratum"
                                    ),
         # adding algal type for graphs about palatability:
         algal_type = case_when(type_code %in% c("BFMA", "GFMA", "RFMA") ~ "Fleshy macroalgae",
                                type_code %in% c("GCMA", "RCMA") ~ "Calcareous macroalgae",
                                type_code %in% c("TA", "TAS", "STA") ~ "Turf algae"),
         # certain substrates are not suitable for coral or algal growth, so should not detract from percent cover
         av_sub_yn = if_else(category_code %in% c("SAND", "HOLE", "SG", "PAVE"), "no", "yes")
         ) %>%
    filter(!is.na(cat_c) & !is.na(species)) # need GRB to fix some data errors, but using this for now
```

## Fish

```{r}
# import and clean data

rf_fish <- read_excel(here("jumby_baseline", "data", "JB_reefs.xlsx"), sheet = "fish") %>%  
  clean_names() %>%
  filter(!is.na(site)) %>% # remove any incomplete rows at the end of the data
  mutate(number = if_else(is.na(number), 1, number)) %>% # all entries with no number specified were single observations
  uncount(number) %>% # expand to replicate rows if multiple fish were recorded to look at length distributions
  mutate(biomass = as.numeric(biomass),
         phase_code = tolower(phase),
         phase = case_when(phase == "j" ~ "Juvenile",
                           phase == "i" ~ "Initial",
                           phase == "t" ~ "Terminal"),
         family_c = if_else(family %in% c("Scaridae", "Acanthuridae", "Haemulidae", "Serranidae", "Lutjanidae", "Balistidae", "Pomacentridae"), family, "Other")
         ) # consolidating the number of families
```

### Total biomass by site
* Regional averages are taken from Karr et al. 2015
```{r}
# calculating total biomass in each transect (sum) -> mean at each site
rf_fish_site <- rf_fish %>% 
  group_by(site, transect) %>%
  summarize(bm_tot = sum(biomass)/1000/120*10000) %>% #kg/ha
  group_by(site) %>%
  summarize(bm_mean = mean(bm_tot),
            bm_se = sd(bm_tot)/sqrt(n()))

# graph
ggplot(rf_fish_site, aes(x = site, y = bm_mean)) +
  geom_col(fill = "cadetblue3", color = "black", alpha = 0.8) +
  geom_errorbar(aes(ymin = bm_mean - bm_se, ymax = bm_mean + bm_se), width = .2,
                 position = position_dodge(.9)) +
  geom_hline(yintercept = 1300, linetype = "dashed", color = "black") +
  annotate("text", x = 5.5, y = 1350, size = 3, hjust = 1, label=c('Caribbean mean unfished biomass')) +
  labs(x = "", y = expression(Total~fish~biomass~(kg~ha^-1))) +
  theme_bw()

kable(rf_fish_site %>% select(site, bm_mean, bm_se))
```

### Biomass by site and family
```{r}
rf_fish_fam_site <- rf_fish %>% 
  expand(nesting(site, transect), family_c) %>%
  left_join(rf_fish %>%
      select(site, transect, family_c, length, biomass)) %>%
  mutate(biomass = if_else(is.na(biomass), 0, biomass)) %>%
  group_by(site, transect, family_c) %>%
  summarize(bm_tot = sum(biomass)/1000/120*10000) %>% #kg/ha
  group_by(site, family_c) %>%
  summarize(bm_mean = mean(bm_tot),
            bm_se = sd(bm_tot)/sqrt(n()))

# graph
ggplot(rf_fish_fam_site, aes(x = family_c, y = bm_mean)) +
  geom_col(fill = "cadetblue3", color = "black", alpha = 0.9) +
  geom_errorbar(aes(ymin = bm_mean - bm_se, ymax = bm_mean + bm_se), width = .2,
                 position = position_dodge(.9)) +
  facet_wrap(. ~ site, ncol = 2) +
  labs(x = "Site", y = "Biomass (kg/ha)") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, h = 1))

kable(rf_fish_fam_site %>% select(site, family = family_c, bm_mean, bm_se))
```

### Herbivore biomass
*  - Eastern Caribbean average parrotfish biomass for NTRs from Steneck et al. 2018 was ~ 1550 g/100m2, or 155 kg/ha. Note that this is potentially misleading because I'm sure a lot of these NTRs do experience some fishing. But mean fished parrotfish biomasses for the same region was ~ 750 g/100m2
```{r}
ggplot(rf_fish_fam_site %>%
        filter(family_c == "Scaridae"),
       aes(x = site, y = bm_mean)) +
  geom_col(color = "black", fill = "cadetblue3", alpha = 0.8, stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = bm_mean - bm_se, ymax = bm_mean + bm_se), width = .2,
                 position = position_dodge(.9)) +
  geom_hline(yintercept = 155, linetype = "dashed", color = "black") +
  annotate("text", x = 8.5, y = 165, size = 3, hjust = 1, label=c('Eastern Caribbean mean unfished biomass')) +
  labs(x = "", y = expression(Parrotfish~biomass~(kg~ha^-1))) +
  theme_bw()

ggplot(rf_fish_fam_site %>%
        filter(family_c == "Acanthuridae"),
       aes(x = site, y = bm_mean)) +
  geom_col(color = "black", fill = "cadetblue3", alpha = 0.8, stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = bm_mean - bm_se, ymax = bm_mean + bm_se), width = .2,
                 position = position_dodge(.9)) +
  labs(x = "", y = expression(Surgeonfish~biomass~(kg~ha^-1))) +
  theme_bw()
```