---
title: "AGRRA 2017-2025 - cleaned"
author: "Molly Wilson"
date: "2026-04-17"
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
    theme: cerulean
    highlight: haddock
---

# Setup

```{r setup, message = F, warning = F, echo = F}
# Install packages
library(tidyverse)
library(here) # for accessing files within directory
library(readxl) # for reading indv. sheets within excel files
library(kableExtra) # for knitting tables
library(purrr) # for joining multiple sheets successively
library(janitor) # for cleaning variable names
library(scales) # for plotting percentages
library(rstatix) # for calculating significance tables
# library(ggbeeswarm) # for jittering in change plots
library(boot) # for bootstrapping CIs
library(forcats) # for wrangling data for forest plots
library(broom) # for visualizing variances
library(patchwork) # for combining plots
library(ggradar) # for radar plots
library(ggpubr) # for adding significances into plots

# Define settings
knitr::opts_chunk$set(message = FALSE, warning = FALSE)

# Define functions
fix_sites <- function(site) {
  site %>%
    str_replace("^Site \\d+.*", str_extract(., "^Site \\d+")) %>%
    recode(
      "Little Bird Patch" = "Little Bird Backreef"
    )
}

fix_names <- function(x) { # fixing issues with AGRRA col names (e.g. tAPALmean -> t_apa_lmean)
  x %>%
    gsub(" ", "_", .) %>%                        # spaces → underscores
    gsub("([a-z])([A-Z])", "\\1_\\2", .) %>%     # split camelCase
    gsub("^t([A-Z]{4})(avg|med|std)$", "t_\\1_\\2", ., perl = TRUE) %>% # tACANavg → t_ACAN_avg
    gsub("(.*?)(avg|med|std)$", "\\1_\\2", ., perl = TRUE) %>%          # ensure _ before suffix
     gsub("^t([A-Z]{4})(large|small|total)$", "t_\\1_\\2", ., perl = TRUE) %>% 
    gsub("(.*?)(large|small|total)$", "\\1_\\2", ., perl = TRUE) %>%
    tolower()
}

clean_excel_coral <- function(file_path) {
  sheets <- excel_sheets(file_path)
  sheets_to_join <- setdiff(sheets, c("TermsOfUse", "Metadata", "Overall"))  # exclude

  sheets_list <- map(sheets_to_join, ~ read_excel(file_path, sheet = .x) %>%
                       rename_with(~ gsub("^(t)([A-Za-z]{4})(avg|med|std)$", "\\1_\\2_\\3", .x, ignore.case = TRUE)) %>%
                       rename_with(~ gsub("([A-Za-z]{4})(avg|med|std)$", "\\1_\\2", .x, ignore.case = TRUE)) %>%
                       clean_names())

  reduce(sheets_list, left_join, by = c("survey_id", "code", "name", "nt"))
}

# Define ggplot themes
theme_pub <- function(base_size = 12, base_family = "") {
  theme_bw(base_size = base_size, base_family = base_family) +
    theme(
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(),
      panel.border = element_rect(color = "black", fill = NA),
      axis.line = element_line(color = "black"),
      axis.title = element_text(size = base_size * 1.2),
      axis.text = element_text(size = base_size),
      legend.title = element_text(size = base_size * 1.1),
      legend.text = element_text(size = base_size * 0.9),
      strip.background = element_rect(fill = "white", color = "black"),
      strip.text = element_text(size = base_size * 1.1, face = "bold")
    )
}

# fig directory
fig_dir <- here("agrra_monitoring", "figs")
```

# Site metadata

```{r}
site_depth <- read_excel(here("agrra_monitoring", "data_raw", "ATG_2017-2022", "Calculated", "Site", "Site.xlsx"), sheet = "Data") %>% 
  clean_names() %>%
  filter(format(date, "%Y") >= 2017) %>%
  select(code, mean_depth)
  
sites_hist <- read_excel(here("agrra_monitoring", "data_raw", "ATG_2017-2022", "Raw", "MSAccess-Metadata.xlsx"), sheet = "Survey") %>% 
  clean_names() %>%
  mutate(site = fix_sites(name)) %>%
  select(site, t_benth_hist = benthic_transects, t_fish_hist = fish_transects)

sites_meta <- read_excel(here("agrra_monitoring", "data_raw", "ATG_NEMMA_2025", "Raw", "Metadata.xlsx"), sheet = "Survey") %>% 
  mutate(zone = "NEMMA") %>%
  bind_rows(read_excel(here("agrra_monitoring", "data_raw", "ATG_NPA_2025", "Raw", "Metadata.xlsx"), sheet = "Survey") %>%
              mutate(zone = "NDNP")) %>%
  clean_names() %>%
  mutate(site = fix_sites(name),
         code = str_replace_all(code, "\\s+", "")) %>%
  select(site, code, zone, latitude, longitude, t_benth_2025 = benthic_transects, t_fish_2025 = fish_transects) %>%
  left_join(sites_hist, by = "site") %>%
  left_join(site_depth, by = "code") %>%
  distinct() %>%
  filter(!is.na(t_benth_hist)) %>% # removing sites with no historical data
  mutate(code = if_else(site == "Green Island 2", "A05-03", code))

write.csv(sites_meta, here("agrra_monitoring", "data_outputs", "sites_meta.csv"))
```

# Benthic

## Point cover and algal heights

Data overview: 

- Need to work with transect-level data to get accurate standard deviations for dif. TA groupings
- 2017-2022: "BenthicPointCoverScaledByTransect.xlsx"
  - scaled to not include non-reef substrate
  - cover is in percent format (X/100)
  - TAS and TAM included but not STA
  - some transect numbers include surveyor initials
- 2025: "BenthicCoverScaledByTransect.xlsx"
  - scaled to not include non-reef substrate)
  - cover is in proportion format (X/1)
  - TAS/TAM/STA only included within tTA sheet (on "overall" sheet TA includes all forms of turf)
  - Little Bird Patch = Little Bird Backreef
- Here "TAS" will refer to TAS + TAM (+ STA in 2025)

```{r}
# define variables to include (removing CYAN for analyses)
benthic_var <- c("lc", "cca", "ta", "tas", "fma", "cma", "ainv", "peys")

benthic_cover_trans <- 
  # historical data
  read_excel(here("agrra_monitoring", "data_raw", "ATG_2017-2022", "Calculated", "BenthicPointCoverScaled", "BenthicPointCoverScaledByTransect.xlsx"), sheet = "Data") %>% 
  clean_names() %>%
  mutate(year = year(date),
         trans = as.character(trans), # to bind with 2025 NPA data that includes initials
         t_tas = tas + tam) %>% 
  select(site, year, trans, benthic_var) %>%
  # bind 2025 data
  bind_rows(
    bind_rows(
      # 2025 NEMMA
      read_excel(here("agrra_monitoring", "data_raw", "ATG_NEMMA_2025", "Calculated", "BenthicCoverScaledByTransect.xlsx"), sheet = "Overall") %>%
        clean_names() %>%
        mutate(`transect_name` = as.character(`transect_name`)) %>%  # to bind with 2025 NPA data with initials
        # incorporate TA sheet
        left_join(read_excel(here("agrra_monitoring", "data_raw", "ATG_NEMMA_2025", "Calculated", "BenthicCoverScaledByTransect.xlsx"), sheet = "tTA") %>%
                    clean_names() %>%
                    select(transect_id, ta, sta, tas, tam)),
      # 2025 NDNP
      read_excel(here("agrra_monitoring", "data_raw", "ATG_NPA_2025", "Calculated", "BenthicCoverScaledByTransect.xlsx"), sheet = "Overall") %>%
        clean_names() %>%
        # incorporate TA sheet
        left_join(read_excel(here("agrra_monitoring", "data_raw", "ATG_NPA_2025", "Calculated", "BenthicCoverScaledByTransect.xlsx"), sheet = "tTA") %>%
                    clean_names() %>%
                    select(transect_id, ta, sta, tas, tam))
    ) %>%
      mutate(year = 2025,
             tas_sum = tas + sta + tam) %>%
      select(site = survey_name, year, trans = transect_name, lc = t_coral, cca = t_cca, ta, tas = tas_sum, fma = t_fma, cma = t_cma, ainv = t_ainv, peys = t_pey)
  ) %>%
  mutate(site = fix_sites(site)) %>% 
  filter(year != 2022, # 2022 was incomplete data collection year for NPA
         site %in% sites_meta$site) %>% 
  left_join(sites_meta %>% select(site, zone), by = "site") %>%
  # generating proportions with subset of benthic_var
  rowwise() %>%
  mutate(total = sum(c_across(all_of(benthic_var)), na.rm = TRUE)) %>%
  mutate(across(all_of(benthic_var), ~ (.x / total))) %>% # proportion format
  mutate(total = sum(c_across(all_of(benthic_var)), na.rm = TRUE)) %>% # test for total = 1
  ungroup()

benthic_cover_site <- benthic_cover_trans %>%
  group_by(site, zone, year) %>%
  rename_with(~ str_remove(.x, "t_")) %>%
  summarise(across(c(benthic_var),
                   list(mean = ~ mean(.x, na.rm = TRUE)),
                   .names = "{.col}_{.fn}")) %>%
  ungroup() %>%
  select(site, zone, year, ends_with("_mean")) %>%
  pivot_longer(cols = ends_with("_mean"),
    names_to = "var",
    values_to = "value") %>%
  mutate(metric = "benthic_cover",
         var = str_remove(var, "_mean"),
         unit = "proportion") 

benthic_cover_year <- benthic_cover_site %>%
  group_by(year, zone, metric, unit, var) %>%
  summarise(mean = mean(value, na.rm = TRUE),
            sd   = sd(value, na.rm = TRUE),
            .groups = "drop")
```

## Algal height
```{r}
algal_height_site <- read_excel(here("agrra_monitoring", "data_raw", "ATG_2017-2022", "Calculated", "AlgaePointCoverScaled", "AlgaePointCoverScaledBySite.xlsx"), sheet = "Data") %>%
  rename_with(fix_names) %>%
  mutate(year = year(date)) %>%
  select(year, site, value = fh_avg) %>%
  bind_rows(read_excel(here("agrra_monitoring", "data_raw", "ATG_NEMMA_2025", "Calculated", "BenthicAlgalHeight.xlsx"), sheet = "Scaled") %>% 
              bind_rows(read_excel(here("agrra_monitoring", "data_raw", "ATG_NPA_2025", "Calculated", "BenthicAlgalHeight.xlsx"), sheet = "Scaled")) %>% 
              rename_with(fix_names) %>%
              mutate(year = 2025) %>%
              select(year, site = name, value = 't_fma-h')) %>%
  mutate(site = fix_sites(site),
         metric = "height",
         var = "fma_height",
         unit = "cm") %>%
  filter(site %in% sites_meta$site & year != 2022) %>%
  left_join(sites_meta %>% select(site, zone))
```


## Algal composition
```{r}
# metadata for genera - family matching
fma_key <- read_excel(here("agrra_monitoring", "data_raw", "ATG_NEMMA_2025", "Calculated", "BenthicCoverScaled.xlsx"), sheet = "Metadata") %>% 
  clean_names() %>%
  fill(scaled_benthic_cover, column_header) %>%
  filter(str_starts(scaled_benthic_cover, "tFMA")) %>%
  select(code = column_header, genera = definition) %>%
  mutate(genera = if_else(code == "BFMA", "Other BFMA",
                          if_else(code == "GFMA", "Other GFMA",
                                  if_else(code == "RFMA", "Other RFMA",
                                          if_else(code == "FMA", "Other FMA", genera)))),
         code = tolower(code)) %>%
  rename(genus = genera)

# currently 2025 only
fma_gen_site <- read_excel(here("agrra_monitoring", "data_raw", "ATG_NEMMA_2025", "Calculated", "BenthicCoverScaled.xlsx"), sheet = "tFMA") %>%
  bind_rows(read_excel(here("agrra_monitoring", "data_raw", "ATG_NPA_2025", "Calculated", "BenthicCoverScaled.xlsx"), sheet = "tFMA")) %>%
  rename_with(fix_names) %>%
  mutate(year = 2025,
         site = fix_sites(name)) %>%
  left_join(sites_meta %>% select(site, zone)) %>%
  select(year, zone, site, contains("avg"), fma_tot_avg = t_fma_avg) %>% # otherwise tFMA and (other) FMA columns get duplicated as FMA
  rename_with(~ str_remove(.x, "_avg")) %>%
  pivot_longer(cols = -c(year, zone, site, fma_tot),
               names_to = "code",
               values_to = "percent_cover") %>%
  filter(fma_tot > 0) %>%
  mutate(prop_cover = percent_cover / fma_tot) %>%
  left_join(fma_key, by = "code") %>%
  select(year, zone, site, code, genus, percent_cover, prop_cover) %>%
  filter(!is.na(genus))

fma_gen_year <- fma_gen_site %>%
  group_by(year, zone, genus) %>%
  summarize(mean_perc_cov = mean(percent_cover),
            sd_perc_cov = sd(percent_cover),
            mean_prop_cov = mean(prop_cover),
            sd_prop_cov = sd(prop_cover)) %>%
  filter(mean_perc_cov > 0)

kable(fma_gen_year %>% filter(genus == "Dictyota"))
```

## Coral composition

Data overview:

- Due to high UNKN in 2019 NDNP data, family-level assessments will only include NEMMA sites
- Extremely complicated and inconsistent data structures to work with - wrote functions for cleaning headers and created meta file to link codes to species and families consistently across years

```{r}
# metadata for species - family matching
coral_meta <- read_excel(here("agrra_monitoring", "data_raw", "ATG_NEMMA_2025", "Calculated", "BenthicCoralCoverScaled.xlsx"), sheet = "Metadata") %>% 
  clean_names() %>%
  fill(scaled_coral_cover, column_header) 

coral_meta_fam <- coral_meta %>%
  filter(str_starts(column_header, "t") | column_header == "UNKN") %>%
  select(family_code = column_header, family = definition) %>%
  mutate(family = gsub("\\s*\\([^\\)]+\\)", "", family)) %>% # removing fluff
  distinct() # removing UNKN duplicate

coral_meta_spp <- coral_meta %>%
  filter(str_starts(scaled_coral_cover, "t") | scaled_coral_cover == "UNKN") %>%
  select(species_code = column_header, family_code = scaled_coral_cover, species = definition) %>%
  left_join(coral_meta_fam, by = "family_code") %>%
  mutate(genus = sub(" .*", "", species),
         family_code = str_remove(family_code, "^t"),
         across(c(species, family), ~ str_replace(., "Unknown sp. of Live Coral", "Unknown spp.")),
         family = if_else(family == "Scleractinia incertae sedis", "Unknown spp.", family))

# historical data
coral_spp_hist <- read_excel(here("agrra_monitoring", "data_raw", "ATG_2017-2022", "Calculated", "CoralCoverSpeciesScaled", "CoralCoverSpeciesScaledBySite.xlsx"), sheet = "Data") %>% 
  rename_with(fix_names) %>%
  mutate(year = year(date)) %>%
  filter(year %in% c(2017, 2019)) %>% # 2022 was incomplete NPA year
  pivot_longer(
    cols = matches("^[a-z]{4}_(avg|std)$"),   # all columns like acer_ave, apal_std, etc.
    names_to = c("species_code", ".value"),        # split into species and measurement type
    names_sep = "_"
  ) %>%
  mutate(across(c(avg, std), ~ .x / 100)) %>% # to match 2025 format (fraction of 1)
  select(site, year, species_code, avg, std)

# 2025 data
coral_spp_2025 <- clean_excel_coral(here("agrra_monitoring", "data_raw", "ATG_NPA_2025", "Calculated", "BenthicCoralCoverScaled.xlsx")) %>%
  bind_rows(clean_excel_coral(here("agrra_monitoring", "data_raw", "ATG_NEMMA_2025", "Calculated", "BenthicCoralCoverScaled.xlsx"))) %>%
  mutate(year = 2025) %>%
  pivot_longer(
    cols = matches("^[a-z]{4}_(avg|std)$"),   # all columns like acer_ave, apal_std, etc.
    names_to = c("species_code", ".value"),        # split into species and measurement type
    names_sep = "_"
  ) %>%
  select(site = name, year, species_code, avg, std)

# select primary families to simplify
coral_fams <- c("Acroporidae", "Faviidae", "Merulinidae", "Milleporidae", "Montastraeidae",  "Poritidae", "Siderastreidae")

# merge for complete df
coral_spp_site <- bind_rows(coral_spp_hist, coral_spp_2025) %>%
  mutate(site = fix_sites(site),
         species_code = toupper(species_code)) %>%
  left_join(sites_meta %>% select(site, zone)) %>%
  left_join(coral_meta_spp, by = "species_code")  %>%
  group_by(site) %>%
  filter(n_distinct(year) > 1 & # only keeping sites with over time comparisons
           family %in% coral_fams) %>% 
  ungroup()

coral_fam_site <- coral_spp_site %>%
  group_by(year, zone, site, family) %>%
  summarize(value = sum(avg, na.rm = TRUE)) %>%
  ungroup() %>%
  mutate(metric = "coral_fam_cover",
         unit = "proportion") %>%
  select(site, zone, year, var = family, value, metric, unit)

coral_fam_year <- coral_fam_site %>%
  group_by(year, zone, var) %>%
  summarize(mean = mean(value, na.rm = TRUE), 
         sd = sd(value, na.rm = TRUE)) %>%
  ungroup() %>%
  group_by(year, zone) %>%
  mutate(rel_mean = mean / sum(mean, na.rm = TRUE)) %>%
  ungroup()
kable(coral_fam_year)
```

## Diadema

Data overview:
- 2025: MotileInverts.xlsx
- 2017-2019: MotileInvertsBySite.xlsx

```{r}
diadema <- read_excel(here("agrra_monitoring", "data_raw", "ATG_2017-2022", "Calculated", "MotileInverts", "MotileInvertsBySite.xlsx"), sheet = "Data") %>% 
  rename_with(fix_names) %>%
  mutate(year = year(date)) %>%
  select(site, year, matches("^dt_|^da_|^dj_")) %>%
  rbind(read_excel(here("agrra_monitoring", "data_raw", "ATG_NEMMA_2025", "Calculated", "MotileInverts.xlsx"), sheet = "Overall") %>% 
          rbind(read_excel(here("agrra_monitoring", "data_raw", "ATG_NPA_2025", "Calculated", "MotileInverts.xlsx"), sheet = "Overall")) %>%
          rename_with(fix_names) %>%
          mutate(year = 2025) %>%
          select(site = name, year, matches("^dt_|^da_|^dj_"), -ends_with("_med"))) %>%
  filter(year != 2022) %>% # incomplete year for NPA
  mutate(site = fix_sites(site)) %>%
  filter(site %in% sites_meta$site) %>%
  left_join(sites_meta %>% select(site, zone)) %>%
  group_by(site) %>%
  filter(n_distinct(year) > 1) %>% # only keeping sites with over time comparisons
  ungroup()

diadema_site <- diadema %>%
  select(zone, site, year, dt_avg, da_avg, dj_avg) %>%  # keep density columns
  pivot_longer(
    cols = c(dt_avg, da_avg, dj_avg),
    names_to = "var",
    values_to = "value"
  ) %>%
  mutate(metric = "diadema_density",
         unit = "g_m2",
         var = recode(var,
                        dt_avg = "total",
                        da_avg = "adult",
                        dj_avg = "juvenile"))

diadema_year <- diadema_site %>%
  group_by(zone, year, metric, unit, var) %>%
  summarise(mean = mean(value, na.rm = TRUE),
            sd   = sd(value, na.rm = TRUE),
            .groups = "drop")

diadema_tests <- diadema_site %>%
  mutate(period = if_else(year == 2025, "y2025", "hist")) %>%
  select(site, zone, var, period, value) %>%
  pivot_wider(names_from = period,
              values_from = value) %>%
  group_by(zone, var) %>%
  summarise(p = wilcox.test(hist, y2025, paired = TRUE, exact = FALSE)$p.value,
    .groups = "drop")
```

# Fish

## Size and densities

Data overview:

- 2017-19: MSAccess-FishRaw.xlsx
    - requires merging transect numbers
- 2025: Metadata.xlsx
    - has different id numbers depending on phase

```{r}
# raw historical data

fish_tax_hist <- read_excel(here("agrra_monitoring", "data_raw", "ATG_2017-2022", "Raw", "MSAccess-Metadata.xlsx"), sheet = "FishTaxonomy", skip = 1, col_names = TRUE) %>% # deals with merged headers
  clean_names() %>%
  mutate(genus = if_else(is.na(genus) | grepl("/", genus), family, genus),
         species = if_else(is.na(species), "spp.", species)) %>%
  unite(species, genus, species, sep = " ", remove = TRUE)

fish_raw_hist <- read_excel(here("agrra_monitoring", "data_raw", "ATG_2017-2022", "Raw", "MSAccess-FishRaw.xlsx"), sheet = "Counts Raw") %>%
  clean_names() %>%
  left_join(read_excel(here("agrra_monitoring", "data_raw", "ATG_2017-2022", "Raw", "MSAccess-Metadata.xlsx"), sheet = "Transect") %>%
              clean_names() %>%
              rename(transect = id, date = surveyed) %>%
              left_join(read_excel(here("agrra_monitoring", "data_raw", "ATG_2017-2022", "Raw", "MSAccess-Metadata.xlsx"), sheet = "Survey") %>%
                          clean_names() %>%
                          select(survey = id, site = name))) %>%
  mutate(year = year(date),
         site = fix_sites(site),
         size_class = str_replace_all(size_class, "-", "_"),
         # convert bin names like "0_5" -> 2.5
         length_mid = str_extract(size_class, "\\d+_?\\d*"),
         length_mid = ifelse(str_detect(length_mid, "_"),
                             sapply(str_split(length_mid, "_"), function(x) mean(as.numeric(x))),
                             as.numeric(length_mid))) %>%
  filter(year != 2022, # 2022 was incomplete data collection year for NPA
         site %in% sites_meta$site) %>% 
  left_join(sites_meta %>% select(site, zone), by = "site") %>%
  left_join(fish_tax_hist, by = c("taxonomy" = "id")) %>%
  mutate(phase = case_when(terminal_phase == TRUE ~ "Terminal",
                           TRUE                  ~ "Initial")) %>%
  select(year, site, zone, trans = transect, size_class, length_mid, common_name, common_family, family, species, phase, count)

# raw 2025 data

fish_tax_2025 <- read_excel(here("agrra_monitoring", "data_raw", "ATG_NEMMA_2025", "Raw", "Metadata.xlsx"), sheet = "FishTaxonomy", skip = 1, col_names = TRUE) %>% # deals with merged headers
  clean_names() %>%
    mutate(genus = if_else(is.na(genus) | grepl("/", genus), family, genus),
         species = if_else(is.na(species), "spp.", species)) %>%
  unite(species, genus, species, sep = " ", remove = TRUE)

fish_raw_2025 <- read_excel(here("agrra_monitoring", "data_raw", "ATG_NEMMA_2025", "Raw", "FishRaw.xlsx"), sheet = "Counts Raw") %>%
  clean_names() %>%
  left_join(read_excel(here("agrra_monitoring", "data_raw", "ATG_NEMMA_2025", "Raw", "Metadata.xlsx"), sheet = "Transect") %>%
              clean_names() %>%
              select(transect = id, survey, date = surveyed) %>%
              left_join(read_excel(here("agrra_monitoring", "data_raw", "ATG_NEMMA_2025", "Raw", "Metadata.xlsx"), sheet = "Survey") %>%
                          clean_names() %>%
                          select(survey = id, site = name))) %>%
  bind_rows(read_excel(here("agrra_monitoring", "data_raw", "ATG_NPA_2025", "Raw", "FishRaw.xlsx"), sheet = "Counts Raw") %>%
              clean_names() %>%
              left_join(read_excel(here("agrra_monitoring", "data_raw", "ATG_NPA_2025", "Raw", "Metadata.xlsx"), sheet = "Transect") %>%
                          clean_names() %>%
                          select(transect = id, survey, date = surveyed) %>%
                          left_join(read_excel(here("agrra_monitoring", "data_raw", "ATG_NPA_2025", "Raw", "Metadata.xlsx"), sheet = "Survey") %>%
                                      clean_names() %>%
                                      select(survey = id, site = name)))) %>%
  clean_names() %>%
  mutate(year = year(date),
         site = fix_sites(site),
         size_class = str_replace_all(size_class, " - ", "_"),
         size_class = str_replace_all(size_class, "cm", "")) %>%
  left_join(sites_meta %>% select(site, zone), by = "site") %>%
  left_join(fish_tax_2025, by = c("taxonomy" = "id")) %>%
  mutate(phase = if_else(terminal_phase == "Yes", "Terminal", "Initial")) %>%
  select(year, site, zone, trans = transect, size_class, common_name, common_family, family, species, phase, count) 

# merge all years

fish_raw <- bind_rows(fish_raw_hist, fish_raw_2025) %>%
  filter(year %in% c(2017, 2019, 2025)) %>%
  mutate(baseline = if_else(year %in% c(2017, 2019), "baseline", "2025")) %>%
  group_by(site, year) %>%
  filter(n_distinct(trans) >= 8) %>%
  ungroup() %>%
  group_by(site) %>%
  filter(n_distinct(year) > 1) %>%
  ungroup()

fish_density_size_spp_trans <- fish_raw %>%
  expand(nesting(year, site, zone, trans, size_class), species) %>%
  left_join(fish_raw %>%
              group_by(year, site, zone, trans, size_class, species) %>%
              summarize(density_trans = sum(count)/60)) %>% # indv./m2 (60m2 belt transects)
  mutate(density_trans = replace_na(density_trans, 0)) %>%
  left_join(fish_tax_2025 %>% select(family, species), by = "species")

# fish_size_spp_site <- fish_density_size_spp_trans %>%
#   group_by(year, site, zone, size_class, species) %>%
#   summarize(density_site = mean(density_trans)) %>%
# need to make this compatible with sum_df_site if using

# subsets of interest for models

scarid_lrg_density_site <- fish_density_size_spp_trans %>%
  filter(family == "Scaridae" & 
           size_class %in% c("21_30", "31_40")) %>% # no 40s identified
  group_by(site, zone, year) %>%
  summarize(value = mean(density_trans)) %>%
  mutate(metric = "fish_density",
         var = "scarids_21_40",
         unit = "indv/m2") %>%
  select(site, zone, year, metric, var, value, unit)

acanthurid_lrg_density_site <- fish_density_size_spp_trans %>%
  filter(family == "Acanthuridae" &
           size_class %in% c("21_30", "31_40")) %>%
  group_by(site, zone, year) %>%
  summarize(value = mean(density_trans)) %>%
  mutate(metric = "fish_density",
         var = "acanthurids_21_40",
         unit = "indv/m2") %>%
  select(site, zone, year, metric, var, value, unit)
  
```


## Biomass

- Currently by family only, but could calculate biomass from raw data to get at the species level

Data overview:

- 2017: FishBiomassBySite.xlsx
   - "Data" sheet has transect-level data by family and group
   - "Key" has family definitions
   - Ignoring "group" classifications here as it isn't consistent with 2025
- 2025: FishBiomass.xlsx
   - Family sheets have transect-level data by species
   - Metadata has species/family definitions (needed substantial reformatting)
- Converting all data to kg/100m2 - convert back to g/100m2

```{r}
# create compatible family key by merging metadata
fish_fam_key <- full_join(
  read_excel(here("agrra_monitoring", "data_raw", "ATG_2017-2022", "Calculated", "FishBiomass", "FishBiomassBySite.xlsx"), sheet = "Key") %>% 
    clean_names() %>%
    slice(21:40) %>%
    select(code_com = 1,
           definition = 2) %>%
    mutate(family = str_remove(word(definition, 1), ":"),
           family = if_else(family == "Wrasse", "Wrasses", 
                            if_else(code_com == "PUFF", "Pufferfishes", family))) %>%
    select(-definition), 
  read_excel(here("agrra_monitoring", "data_raw", "ATG_NEMMA_2025", "Calculated", "FishBiomass.xlsx"), sheet = "Metadata") %>% 
    clean_names() %>%
    slice(23:42) %>%
    select(code_lat = 2,
           definition = 3) %>%
    mutate(family = str_remove(word(definition, 1), ":")) %>%
    select(-definition), 
  by = "family") %>%
  mutate(code_lat = tolower(gsub("^t", "", code_lat)),
         code_com = tolower(code_com)) %>%
  select(code_com, code_lat, family) %>%
  bind_rows(
    tibble(code_com = "t",
           code_lat = "all",
           family   = "Total")) %>%
  pivot_longer(cols = c(code_com, code_lat),
               names_to = "code_type",
               values_to = "code")

fish_bm_fam_site <- bind_rows(read_excel(here("agrra_monitoring", "data_raw", "ATG_2017-2022",
                  "Calculated", "FishBiomass", "FishBiomassBySite.xlsx"),
             sheet = "Data") %>%
    rename_with(fix_names) %>%
    mutate(code_type = "code_com",
             year = year(date)) %>%
    filter(year %in% c(2017, 2019)),
  # 2025 data
  bind_rows(read_excel(here("agrra_monitoring", "data_raw", "ATG_NEMMA_2025",
                    "Calculated", "FishBiomass.xlsx"),
               sheet = "Overall"),
        read_excel(here("agrra_monitoring", "data_raw", "ATG_NPA_2025",
                    "Calculated", "FishBiomass.xlsx"),
               sheet = "Overall")) %>%
    rename_with(fix_names) %>%
    mutate(site = name,
           year = 2025,
           code_type = "code_lat") %>%
    rename_with(~ gsub("^t_", "", .x), .cols = contains("avg"))) %>%
  # wrangle data
  mutate(site = fix_sites(site)) %>%
  ## min n transect = 8 in both years
  group_by(site) %>%
  filter(all(nt >= 8)) %>%
  ungroup() %>%
  select(site, year, code_type, contains("avg")) %>%
  pivot_longer(cols = contains("avg"),
               names_to = "family_stat",
               values_to = "value") %>%
  separate(family_stat,
           into = c("code", "stat"),
           sep = "_(?=[^_]+$)") %>%
  # merge with family key
  left_join(fish_fam_key, by = c("code_type", "code")) %>%
  left_join(sites_meta %>% select(site, zone)) %>%
  mutate(
    #value = value/1000, # converting from g to kg
         metric = "fish_biomass",
         unit = "g_100m2") %>%
  filter(site %in% sites_meta$site,
         year != 2022, # incomplete NDNP year
         !is.na(family)) %>% # historical groupings excluded from key
  select(site, zone, year, metric, var = family, value, unit)

herb_bm_site <- fish_bm_fam_site %>%
  filter(var %in% c("Parrotfishes", "Surgeonfishes")) %>%
  group_by(site, zone, year, metric, unit) %>%
  summarise(value = sum(value, na.rm = TRUE),
    .groups = "drop") %>%
  mutate(var = "Herbivores") %>%
  select(site, zone, year, metric, var, value, unit)

fish_bm_site <- rbind(fish_bm_fam_site, herb_bm_site) %>%
  mutate(var = recode(var, 
                      "Parrotfishes" = "Scarids",
                      "Surgeonfishes" = "Acanthurids")) %>%
  filter(var %in% c("Total", "Scarids", "Acanthurids", "Herbivores"))

fish_bm_year <- fish_bm_site %>%
  group_by(year, zone, metric, var, unit) %>%
  summarize(mean = mean(value),
            sd = sd(value))
```

## Average fish weight

Calculating AFW = biomass/density by site to compare with Valles and Oxenford 2014

Data overview:

- fish_bm_site: biomass by site for total, herbivores, scarids, and acanthurids
- densities:
  - 2017-19: FishDensityBySite.xlsx
    - sheet = "Data" 
    - T = total, PARR = scarids, SURG = acanthurids
  - 2025: FishDensity.xlsx
    - sheet = "Overall"
    = ALL = total, SCAR = scarids, ACAN = acanthurids
    
```{r}
fish_dens_fam_site <- bind_rows(
  # historical data
  read_excel(here("agrra_monitoring", "data_raw", "ATG_2017-2022", "Calculated", "FishDensity", "FishDensityBySite.xlsx"), sheet = "Data") %>%
    rename_with(fix_names) %>%
    mutate(year = year(date),
           code_type = "code_com"),
  # 2025 data
  bind_rows(read_excel(here("agrra_monitoring", "data_raw", "ATG_NEMMA_2025", "Calculated", "FishDensity.xlsx"), sheet = "Overall"),
            read_excel(here("agrra_monitoring", "data_raw", "ATG_NPA_2025", "Calculated", "FishDensity.xlsx"), sheet = "Overall")) %>%
    rename_with(fix_names) %>%
    mutate(site = name,
           year = 2025,
           code_type = "code_lat") %>%
    rename_with(~ gsub("^t_", "", .x), .cols = contains("avg"))
) %>%
  group_by(site) %>%
  filter(all(nt >= 8)) %>%
  ungroup() %>%
  mutate(site = fix_sites(site)) %>%
  select(site, year, code_type, contains("avg")) %>%
  pivot_longer(cols = contains("avg"),
               names_to = "family_stat",
               values_to = "value") %>%
  separate(family_stat,
           into = c("code", "stat"),
           sep = "_(?=[^_]+$)") %>%
  # merge with family key
  left_join(fish_fam_key, by = c("code_type", "code")) %>%
  left_join(sites_meta %>% select(site, zone)) %>%
  mutate(metric = "fish_density",
         unit = "indv_100m2") %>%
  filter(site %in% sites_meta$site,
         year != 2022, # incomplete NDNP year
         !is.na(family)) %>% # historical groupings excluded from key
  select(site, zone, year, metric, var = family, value, unit)

herb_dens_site <- fish_dens_fam_site %>%
  filter(var %in% c("Parrotfishes", "Surgeonfishes")) %>%
  group_by(site, zone, year, metric, unit) %>%
  summarise(value = sum(value, na.rm = TRUE),
    .groups = "drop") %>%
  mutate(var = "Herbivores") %>%
  select(site, zone, year, metric, var, value, unit)
  
fish_dens_site <- bind_rows(fish_dens_fam_site, herb_dens_site) %>%
  mutate(var = recode(var, 
                      "Parrotfishes" = "Scarids",
                      "Surgeonfishes" = "Acanthurids")) %>%
  filter(var %in% c("Total", "Scarids", "Acanthurids", "Herbivores"))
  
fish_afw_site <- bind_rows(fish_dens_site, fish_bm_site) %>%
  select(site, zone, year, var, metric, value) %>%
  pivot_wider(names_from = metric,
    values_from = value) %>%
  mutate(ave_fish_weight = fish_biomass / fish_density,
         metric = "ave_fish_weight",
         unit = "g_fish") %>%
  select(site, zone, year, metric, var, value = ave_fish_weight, unit)

fish_afw_year <- fish_afw_site %>%
  group_by(year, zone, metric, var, unit) %>%
  summarize(mean = mean(value),
            sd = sd(value))
```
    

# Summary dfs
```{r}
sum_site_yr <- bind_rows(benthic_cover_site, algal_height_site, coral_fam_site, diadema_site, fish_bm_site, fish_afw_site, scarid_lrg_density_site, acanthurid_lrg_density_site) %>%
  mutate(var = tolower(var))

# create explicit paired site-level dataset
sum_site_pairs <- sum_site_yr %>%
  mutate(period = if_else(year == 2025, "2025", "hist")) %>%
  select(site, zone, metric, var, unit, period, value) %>%
  pivot_wider(
    names_from = period,
    values_from = value
  ) %>%
  filter(!is.na(hist) & !is.na(`2025`)) %>%
  mutate(delta = `2025` - hist)

options(scipen = 999) # avoid scientific notation in printing values

# create zone-level summary
# sum_zone <- sum_site_pairs %>%
#   group_by(zone, metric, var, unit) %>%
#   summarise(mean_hist = mean(hist, na.rm = TRUE),
#             sd_hist   = sd(hist, na.rm = TRUE),
#             mean_2025 = mean(`2025`, na.rm = TRUE),
#             sd_2025   = sd(`2025`, na.rm = TRUE),
#             pct_change = ((mean(`2025`, na.rm = TRUE) -
#                              mean(hist, na.rm = TRUE)) /
#                             mean(hist, na.rm = TRUE)) * 100,
#             delta = mean(delta, na.rm = TRUE),
#             CI = list(
#               quantile(
#                 replicate(
#                   5000,
#                   mean(sample(delta,
#                               replace = TRUE),
#                        na.rm = TRUE)),
#                 probs = c(0.025, 0.975),
#                 na.rm = TRUE)),
#             p = suppressWarnings(
#               wilcox.test(
#                 hist,
#                 `2025`,
#                 paired = TRUE,
#                 exact = FALSE
#               )$p.value
#             ),
#             .groups = "drop") %>%
#   mutate(metric_pub = str_to_sentence(gsub("_", " ", metric)),
#          CI_low  = map_dbl(CI, 1),
#          CI_high = map_dbl(CI, 2),
#          var_pub = recode(var,
#                           lc   = "Live coral",
#                           cca  = "CCA",
#                           ta   = "Turf algae",
#                           tas  = "Turf-sediment complex",
#                           fma  = "Fleshy macroalgae",
#                           cma  = "Calc. macroalgae",
#                           ainv = "Aggressive inverts",
#                           peys = "Peysonnelids",
#                           fma_height = "Fleshy macroalgal height",
#                           acanthurids_21_40 = "Acanthurids > 20cm",
#                           scarids_21_40 = "Scarids > 20cm",
#                           .default = str_to_title(var)
#          )) %>%
#   select(-CI) %>%
#   rstatix::add_significance(p.col = "p") %>%
#   rename(p_label = p.signif) %>%
#   select(zone, metric, metric_pub, var, var_pub, unit, everything())

sum_zone <- sum_site_pairs %>%
  rename(site_delta = delta) %>%
  group_by(zone, metric, var, unit) %>%
  summarise(
    mean_hist = mean(hist, na.rm = TRUE),
    sd_hist   = sd(hist, na.rm = TRUE),
    mean_2025 = mean(`2025`, na.rm = TRUE),
    sd_2025   = sd(`2025`, na.rm = TRUE),

    pct_change = ((mean_2025 - mean_hist) / mean_hist) * 100,

    CI = list({
      d <- site_delta[!is.na(site_delta)]

      if (length(d) < 2) {
        c(NA_real_, NA_real_)
      } else {
        quantile(
          replicate(
            5000,
            mean(sample(d, size = length(d), replace = TRUE))
          ),
          probs = c(0.025, 0.975),
          names = FALSE
        )
      }
    }),

    delta = mean(site_delta, na.rm = TRUE),

    p = suppressWarnings(
      wilcox.test(hist, `2025`, paired = TRUE, exact = FALSE)$p.value
    ),

    n_sites = sum(!is.na(site_delta)),
    .groups = "drop"
  ) %>%
  mutate(
    metric_pub = str_to_sentence(gsub("_", " ", metric)),
    CI_low  = map_dbl(CI, 1),
    CI_high = map_dbl(CI, 2),
    var_pub = recode(
      var,
      lc   = "Live coral",
      cca  = "CCA",
      ta   = "Turf algae",
      tas  = "Turf-sediment complex",
      fma  = "Fleshy macroalgae",
      cma  = "Calc. macroalgae",
      ainv = "Aggressive inverts",
      peys = "Peysonnelids",
      fma_height = "Fleshy macroalgal height",
      acanthurids_21_40 = "Acanthurids > 20cm",
      scarids_21_40 = "Scarids > 20cm",
      .default = str_to_title(var)
    )
  ) %>%
  select(-CI) %>%
  rstatix::add_significance(p.col = "p") %>%
  rename(p_label = p.signif) %>%
  select(zone, metric, metric_pub, var, var_pub, unit, everything())

sum_zone_pub <- sum_zone %>% 
  unite(variable, metric_pub, var_pub, sep = " - ", remove = TRUE) %>%
  select(Zone = zone, 
         Variable = variable, 
         Unit = unit, 
         `Historical mean` = mean_hist, 
         `Historical SD` = sd_hist, 
         `2025 mean` = mean_2025, 
         `2025 SD` = sd_2025, 
         `Pct. change` = pct_change, 
         Delta = delta, 
         `CI low` = CI_low, 
         `CI high` = CI_high, 
         p, 
         Significance = p_label)

kable(sum_zone_pub)
write.csv(sum_zone_pub, here("agrra_monitoring", "data_outputs", "sum_zone.csv"))
```

# Models
```{r}
delta_herb_fma <- sum_site_pairs %>%
  filter((metric == "fish_biomass" & var == "herbivores") | 
           (metric == "ave_fish_weight" & var == "scarids") |
           var %in% c("scarids_21_40", "acanthurids_21_40", "fma", "fma_height")) %>%
  mutate(var = paste(metric, var, sep = "_")) %>%
  select(site, zone, var, `2025`, delta) %>% # not including historical numbers here anymore
  pivot_wider(
    names_from = var,
    values_from = c(`2025`, delta)
  ) %>%
  rename_with(~ str_replace(.x, "^([^_]+)_[^_]+_(.*)$", "\\1_\\2")) %>%
  filter(!is.na(`2025_biomass_herbivores`)) # sites without fish data / low transect numbers


# spearman_herb2025_fmadelta <- cor.test(delta_herb_fma$`2025_fish_biomass_herbivores`,
#          delta_herb_fma$site_delta_benthic_cover_fma,
#          method = "spearman")
# spearman_herb2025_fmadelta
# 
# spearman_afw2025_fmadelta <- cor.test(delta_herb_fma$`2025_ave_fish_weight_parrotfishes`,
#          delta_herb_fma$site_delta_benthic_cover_fma,
#          method = "spearman")
# spearman_afw2025_fmadelta

library(dplyr)
library(purrr)
library(tibble)

responses <- c("2025_cover_fma", "delta_cover_fma")

predictors <- c(
  "2025_biomass_herbivores",
  "2025_fish_weight_scarids",
  "2025_density_scarids_21_40",
  "2025_density_acanthurids_21_40"
)

spearman_results <- map_dfr(responses, function(resp) {
  map_dfr(predictors, function(pred) {
    
    df <- delta_herb_fma[, c(resp, pred)] %>%
      na.omit()
    
    test <- cor.test(df[[resp]], df[[pred]], method = "spearman")
    
    tibble(
      response = resp,
      predictor = pred,
      rho = unname(test$estimate),
      p_value = test$p.value,
      n = nrow(df)
    )
  })
}) %>%
  mutate(
    rho = round(rho, 2),
    p_value = round(p_value, 3)
  ) %>%
  mutate(
    response = recode(
      response,
      "2025_cover_fma" = "2025 FMA cover",
      "delta_cover_fma" = "Change in FMA cover"
    ),
    predictor = recode(
      predictor,
      "2025_biomass_herbivores" = "Herbivore biomass",
      "2025_fish_weight_scarids" = "Scarid mean fish weight",
      "2025_density_scarids_21_40" = "Scarid density (21–40 cm)",
      "2025_density_acanthurids_21_40" = "Acanthurid density (21–40 cm)"
    )
  )

spearman_results

spearman_wide <- spearman_results %>%
  mutate(result = paste0("ρ = ", rho, ", p = ", p_value)) %>%
  select(predictor, response, result) %>%
  pivot_wider(names_from = response, values_from = result)

spearman_wide
kable(spearman_wide)
write.csv(spearman_wide, here("agrra_monitoring", "data_outputs", "fma_herb_spearman.csv"))
```


# Plots

### Benthic bar
```{r}
benthic_barplot <- benthic_cover_year %>%
  filter(var %in% c(benthic_var)) %>%
  # create "other" category for plot
  mutate(var = if_else(var %in% c("cma", "peys", "ainv"),
                          "other",
                          var)) %>%
  # recalculate means with new "other" category
  group_by(year, zone, var) %>%
  summarise(mean = sum(mean),
            .groups = "drop") %>%
  # labels
  mutate(
    var = recode(
      var,
      lc    = "Live coral",
      cca   = "CCA",
      ta    = "Turf algae",
      tas   = "Turf + sediment",
      fma   = "Fleshy macroalgae",
      other = "Other"
    ),
    var = factor(
      var,
      levels = c(
        "Other",
        "Live coral",
        "CCA",
        "Turf algae",
        "Turf + sediment",
        "Fleshy macroalgae")),
    year = factor(year, levels = c(2025, 2019, 2017)))

ggplot(benthic_barplot,
       aes(x = mean,
           y = factor(year),
           fill = var)) +
  geom_col(width = 0.7) +
  facet_wrap(~zone,
             scales = "free_y") +
  scale_fill_manual(values = c(
    "Live coral" = "#fc8d59",
    "CCA" = "pink",
    "Turf algae" = "#d9ef8b",
    "Turf + sediment" = "tan",
    "Fleshy macroalgae" = "#1b9e77",
    "Other" = "gray80")) +
  scale_x_continuous(labels = scales::percent_format(accuracy = 1)) +
  labs(x = "Percent cover",
       y = NULL,
       fill = NULL) +
  guides(fill = guide_legend(
    nrow = 1,
    reverse = TRUE,
    byrow = TRUE)) +
  theme_minimal(base_size = 12) +
  theme(panel.grid.major.y = element_blank(),
        panel.grid.minor = element_blank(),
        strip.text = element_text(face = "bold"),
        axis.title.x = element_text(
          margin = margin(t = 12)),
        legend.position = "bottom",
        legend.text = element_text(size = 10),
        legend.key.width = unit(0.8, "cm"),
        legend.box = "horizontal")

# ggsave(here(fig_dir, "benthic_sum.png"), width = 8, height = 5)
```

### Benthic delta

```{r}
p_delta_pc <- ggplot(sum_zone %>%
                       filter(metric == "benthic_cover",
                              var %in% benthic_var) %>%
                       mutate(delta_pct = delta * 100,
                              CI_low_pct = CI_low * 100,
                              CI_high_pct = CI_high * 100),
                     aes(x = delta_pct, y = forcats::fct_rev(var_pub))) +
  geom_vline(xintercept = 0,
             linetype = "dashed",
             color = "grey50") +
  geom_errorbarh(aes(xmin = CI_low_pct,
                     xmax = CI_high_pct),
                 height = 0.2) +
  geom_point(size = 3) +
  facet_wrap(~ zone) +
  labs(x = "Mean change (percent cover)",
       y = NULL) +
  theme_bw() +
  theme(strip.background = element_blank(),
        strip.text = element_text(face = "bold"),
        panel.grid.major.y = element_blank())


p_delta_fmah <- ggplot(sum_zone %>%
                         filter(var == "fma_height"),
                       aes(x = delta, y = var_pub)) +
  geom_vline(xintercept = 0,
             linetype = "dashed",
             color = "grey50") +
  geom_errorbarh(aes(xmin = CI_low,
                     xmax = CI_high),
                 height = 0.2) +
  geom_point(size = 3) +
  facet_wrap(~ zone) +
  labs(x = "Mean change (cm)",
       y = NULL) +
  theme_bw() +
  theme(strip.background = element_blank(),
        strip.text = element_text(face = "bold"),
        panel.grid.major.y = element_blank())

p_delta_pc / p_delta_fmah + plot_layout(heights = c(3, 1))
ggsave(here(fig_dir, "benthic_delta_panel.png"), width = 6, height = 4)
```

### Coral composition
```{r}
p_coral_delta <- ggplot(sum_zone %>%
                       filter(metric == "coral_fam_cover" & zone == "NEMMA") %>%
                       mutate(delta_pct = delta * 100,
                              CI_low_pct = CI_low * 100,
                              CI_high_pct = CI_high * 100),
                     aes(x = delta_pct, y = forcats::fct_rev(var_pub))) +
  geom_vline(xintercept = 0,
             linetype = "dashed",
             color = "grey50") +
  geom_errorbarh(aes(xmin = CI_low_pct,
                     xmax = CI_high_pct),
                 height = 0.2) +
  geom_point(size = 2) +
  labs(x = "Mean change (% cover)",
       y = NULL) +
  theme_bw() +
  theme(strip.background = element_blank(),
        strip.text = element_text(face = "bold"),
        panel.grid.major.y = element_blank())

p_coral_heat <- ggplot(coral_fam_site %>%
         filter(zone == "NEMMA") %>%
         left_join(sites_meta %>% select(site, code)),
       aes(code, forcats::fct_rev(var), fill = value)) +
  geom_tile() +
  facet_wrap(~year) +
  scale_fill_viridis_c(labels = percent,
                       name = "Cover") +
  labs(x = "Site",
       y = NULL) +
  theme_minimal(base_size = 12) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 7),
        axis.text.y = element_blank(),
        legend.position = "right")

(p_coral_delta | p_coral_heat) + plot_layout(widths = c(1, 2))
ggsave(here(fig_dir, "coral_fam_comp.png"), width = 8, height = 3.5)
```

### Fish biomass line
```{r}
make_lp_summary <- function(summary_df,
                            sig_df,
                            metric_name,
                            family_name,
                            ylab,
                            hline_value = NULL) {

  plot_df <- summary_df %>%
    filter(metric == metric_name,
           var == family_name) %>%
    group_by(zone, year) %>%
    summarise(
      mean = mean(value, na.rm = TRUE),
      sd   = sd(value, na.rm = TRUE),
      .groups = "drop"
    ) %>%
    mutate(year = as.numeric(year))

  sig_df_var <- sig_df %>%
    filter(metric == metric_name,
           var == family_name)

  plot_df <- plot_df %>%
    left_join(sig_df_var, by = "zone")

  p <- ggplot(plot_df,
              aes(x = year,
                  y = mean,
                  group = zone,
                  color = zone)) +
    geom_point(size = 3) +
    geom_line(size = 1) +
    scale_color_manual(values = c("black", "cadetblue")) +
    scale_x_continuous(breaks = seq(2017, 2025, by = 2),  # odd years from 2017 to 2025
                       labels = seq(2017, 2025, by = 2)) +
    geom_errorbar(aes(ymin = mean - sd,
          ymax = mean + sd),
      width = 0.2) +
    geom_text(data = plot_df %>% filter(year == max(year)),
      aes(label = p_label),
      hjust = -0.3,
      size = 5,
      show.legend = FALSE) +
    labs(x = "",
      y = ylab,
      color = NULL) +
    theme_pub() +
    theme(legend.position = "bottom")

  if (!is.null(hline_value)) {
    p <- p +
      geom_hline(
        yintercept = hline_value,
        linetype = "dashed",
        color = "black"
      )
  }

  return(p)
}

p_bm_tot <- make_lp_summary(
  summary_df = sum_site_yr,
  sig_df = sum_zone,
  metric_name = "fish_biomass",
  family_name = "total",
  ylab = expression("Total fish biomass (" * g ~ 100 ~ m^{-2} * ")"),
  hline_value = 1306 * 10 # Karr et al 2015 - 1300 kg/ha - converting to g/100m2
)

p_bm_herb <- make_lp_summary(
  summary_df = sum_site_yr,
  sig_df = sum_zone,
  metric_name = "fish_biomass",
  family_name = "herbivores",
  ylab = expression("Herbivore biomass (" * g ~ 100 ~ m^{-2} * ")"),
  hline_value = 29.2*100 # Edwards et al 2014 - 29.2 g/m2
)

p_bm_scar <- make_lp_summary(
  summary_df = sum_site_yr,
  sig_df = sum_zone,
  metric_name = "fish_biomass",
  family_name = "scarids",
  ylab = expression("Scarid biomass (" * g ~ 100 ~ m^{-2} * ")"),
  hline_value = 1989 # Valles & Oxenford 2014 - 29.2 g/m2
)

p_afw_scar <- make_lp_summary(
  summary_df = sum_site_yr,
  sig_df = sum_zone,
  metric_name = "ave_fish_weight",
  family_name = "scarids",
  ylab = expression("Average scarid weight (" * g ~ indv^{-1} * ")"),
  hline_value = 215 # Valles & Oxenford 2014 
)

p_bm_tot / p_bm_herb / p_afw_scar + plot_layout(heights = c(10,10,10)) +
  plot_layout(guides = "collect") &
  theme(legend.position = "top")
# (p_bm_tot | p_bm_herb) /
# (p_bm_scar | p_afw_scar) +
#   plot_layout(guides = "collect") &
#   theme(legend.position = "top")
ggsave(here(fig_dir, "fish_bm_change.png"), width = 5, height = 10)
```
