---
title: "AGRRA 2017-2025 - cleaned"
author: "Molly Wilson"
date: "2026-04-17"
output: html_document
---

# Setup

```{r setup, message = F, warning = F, echo = F}
# Install packages
library(tidyverse)
library(here) # for accessing files within directory
library(readxl) # for reading indv. sheets within excel files
library(purrr) # for joining multiple sheets successively
library(janitor) # for cleaning variable names
library(scales) # for plotting percentages
library(rstatix) # for calculating significance tables
library(ggbeeswarm) # for jittering in change plots
library(boot) # for bootstrapping CIs
library(forcats) # for wrangling data for forest plots
library(broom) # for visualizing variances
library(patchwork) # for combining plots
library(ggpubr) # for adding significances into plots

# Define settings
knitr::opts_chunk$set(message = FALSE, warning = FALSE)

# Define functions
fix_names <- function(x) { # fixing issues with AGRRA col names (e.g. tAPALmean -> t_apa_lmean)
  x %>%
    gsub(" ", "_", .) %>%                        # spaces → underscores
    gsub("([a-z])([A-Z])", "\\1_\\2", .) %>%     # split camelCase
    gsub("^t([A-Z]{4})(avg|med|std)$", "t_\\1_\\2", ., perl = TRUE) %>% # tACANavg → t_ACAN_avg
    gsub("(.*?)(avg|med|std)$", "\\1_\\2", ., perl = TRUE) %>%          # ensure _ before suffix
     gsub("^t([A-Z]{4})(large|small|total)$", "t_\\1_\\2", ., perl = TRUE) %>% 
    gsub("(.*?)(large|small|total)$", "\\1_\\2", ., perl = TRUE) %>%
    tolower()
}

se <- function(x) sd(x, na.rm = TRUE) / sqrt(sum(!is.na(x)))

# Define ggplot themes
theme_pub <- function(base_size = 14, base_family = "") {
  theme_bw(base_size = base_size, base_family = base_family) +
    theme(
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(),
      panel.border = element_rect(color = "black", fill = NA),
      axis.line = element_line(color = "black"),
      axis.title = element_text(size = base_size * 1.2),
      axis.text = element_text(size = base_size),
      legend.title = element_text(size = base_size * 1.1),
      legend.text = element_text(size = base_size * 0.9),
      strip.background = element_rect(fill = "white", color = "black"),
      strip.text = element_text(size = base_size * 1.1, face = "bold")
    )
}

# fig directory
fig_dir <- here("agrra_monitoring", "figs")
```

# Site metadata

```{r}
sites_hist <- read_excel(here("agrra_monitoring", "data_raw", "ATG_2017-2022", "Raw", "MSAccess-Metadata.xlsx"), sheet = "Survey") %>% 
  clean_names() %>%
  mutate(name = if_else(str_detect(name, "^Site \\d+"), 
                        str_extract(name, "^Site \\d+"), 
                        name)) %>%
  select(site = name, t_benth_hist = benthic_transects, t_fish_hist = fish_transects)

sites_meta <- read_excel(here("agrra_monitoring", "data_raw", "ATG_NEMMA_2025", "Raw", "Metadata.xlsx"), sheet = "Survey") %>% 
  mutate(site_cat = "NEMMA") %>%
  bind_rows(read_excel(here("agrra_monitoring", "data_raw", "ATG_NPA_2025", "Raw", "Metadata.xlsx"), sheet = "Survey") %>%
              mutate(site_cat = "NDNP")) %>%
  clean_names() %>%
  mutate(name = if_else(str_detect(name, "^Site \\d+"), 
                        str_extract(name, "^Site \\d+"), 
                        name),
         name = if_else(name == "Little Bird Patch", "Little Bird Backreef", name)) %>%
  select(site = name, site_cat, latitude, longitude, t_benth_2025 = benthic_transects, t_fish_2025 = fish_transects) %>%
  left_join(sites_hist, by = "site") %>%
  distinct() %>%
  filter(!is.na(t_benth_hist)) # removing sites with no historical data

write.csv(sites_meta, here("agrra_monitoring", "data_outputs", "sites_meta.csv"))
```

# Benthic data

Data overview: 

- Need to work with transect-level data to get accurate standard deviations for dif. TA groupings
- 2017-2022: "BenthicPointCoverScaledByTransect.xlsx"
  - scaled to not include non-reef substrate
  - cover is in percent format (X/100)
  - TAS and TAM included but not STA
  - some transect numbers include surveyor initials
- 2025: "BenthicCoverScaledByTransect.xlsx"
  - scaled to not include non-reef substrate)
  - cover is in proportion format (X/1)
  - TAS/TAM/STA only included within tTA sheet (on "overall" sheet TA includes all forms of turf)
  - Little Bird Patch = Little Bird Backreef
- Here "TAS" will refer to TAS + TAM (+ STA in 2025)

```{r}
# define variables to include (will remove CYAN and PEYS from some analyses)
benthic_var <- c("lc", "cca", "ta", "tas", "fma", "cma", "cyan", "ainv", "peys")

benthic_cover_trans <- 
  # historical data
  read_excel(here("agrra_monitoring", "data_raw", "ATG_2017-2022", "Calculated", "BenthicPointCoverScaled", "BenthicPointCoverScaledByTransect.xlsx"), sheet = "Data") %>% 
  clean_names() %>%
  mutate(year = year(date),
         trans = as.character(trans), # to bind with 2025 NPA data that includes initials
         t_tas = tas + tam) %>% 
  select(site, year, trans, benthic_var) %>%
  mutate(across(benthic_var, ~ . / 100)) %>% # convert to proportion format
  # bind 2025 data
  bind_rows(
    bind_rows(
      # NEMMA
      read_excel(here("agrra_monitoring", "data_raw", "ATG_NEMMA_2025", "Calculated", "BenthicCoverScaledByTransect.xlsx"), sheet = "Overall") %>%
        clean_names() %>%
        mutate(`transect_name` = as.character(`transect_name`)) %>%  # to bind with 2025 NPA data with initials
        # incorporate TA sheet
        left_join(read_excel(here("agrra_monitoring", "data_raw", "ATG_NEMMA_2025", "Calculated", "BenthicCoverScaledByTransect.xlsx"), sheet = "tTA") %>%
                    clean_names() %>%
                    select(transect_id, ta, sta, tas, tam)),
      # NDNP
      read_excel(here("agrra_monitoring", "data_raw", "ATG_NPA_2025", "Calculated", "BenthicCoverScaledByTransect.xlsx"), sheet = "Overall") %>%
        clean_names() %>%
        # incorporate TA sheet
        left_join(read_excel(here("agrra_monitoring", "data_raw", "ATG_NPA_2025", "Calculated", "BenthicCoverScaledByTransect.xlsx"), sheet = "tTA") %>%
                    clean_names() %>%
                    select(transect_id, ta, sta, tas, tam))) %>%
      mutate(year = 2025,
             survey_name = if_else(survey_name == "Little Bird Patch", "Little Bird Backreef", survey_name),
             tas_sum = tas + sta + tam) %>%
      select(site = survey_name, year, trans = transect_name, lc = t_coral, cca = t_cca, ta, tas = tas_sum, fma = t_fma, cma = t_cma, cyan = t_cyan, ainv = t_ainv, peys = t_pey)) %>%
  mutate(site = if_else(str_detect(site, "^Site \\d+"), 
                        str_extract(site, "^Site \\d+"), 
                        site)) %>% # dealing with NPA sites that have site numbers with different name variations afterwards
  filter(year != 2022, # 2022 was incomplete data collection year for NPA
         site %in% sites_meta$site) %>% 
  left_join(sites_meta %>% select(site, site_cat), by = "site")

# algal height 
algal_height_hist <- read_excel(here("agrra_monitoring", "data_raw", "ATG_2017-2022", "Calculated", "AlgaePointCoverScaled", "AlgaePointCoverScaledBySite.xlsx"), sheet = "Data") %>%
  rename_with(fix_names) %>%
  mutate(year = year(date)) %>%
  select(year, site, fma_height_mean = fh_avg)
  
algal_height_2025 <- read_excel(here("agrra_monitoring", "data_raw", "ATG_NEMMA_2025", "Calculated", "BenthicAlgalHeight.xlsx"), sheet = "Scaled") %>% 
  bind_rows(read_excel(here("agrra_monitoring", "data_raw", "ATG_NPA_2025", "Calculated", "BenthicAlgalHeight.xlsx"), sheet = "Scaled")) %>% 
  rename_with(fix_names) %>%
  mutate(year = 2025,
         site = if_else(name == "Little Bird Patch", "Little Bird Backreef", name)) %>%
  select(year, site, fma_height_mean = 't_fma-h')

algal_height_site <- bind_rows(algal_height_hist, algal_height_2025) %>%
  mutate(site = if_else(str_detect(site, "^Site \\d+"), 
                        str_extract(site, "^Site \\d+"), 
                        site), # dealing with NPA sites that have site numbers with different name variations afterwards
         site_cat = if_else(str_detect(site, "^Site \\d+"), "NDNP",
                            if_else(str_detect(site, "Redonda"), "Redonda",
                                    "NEMMA"))) %>%
  filter(year != 2022 & site_cat %in% c("NEMMA", "NDNP"))

benthic_cover_site <- benthic_cover_trans %>%
  group_by(site, site_cat, year) %>%
  rename_with(~ str_remove(.x, "t_")) %>%
  summarise(across(c(benthic_var),
                   list(mean = ~ mean(.x, na.rm = TRUE),
                        sd   = ~ sd(.x)),
                   .names = "{.col}_{.fn}")) %>%
  ungroup()
```

