---
title: "AGRRA 2017-2025 - cleaned"
author: "Molly Wilson"
date: "2026-04-17"
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
    theme: cerulean
    highlight: haddock
---

# Setup

```{r setup, message = F, warning = F, echo = F}
# Install packages
library(tidyverse)
library(here) # for accessing files within directory
library(readxl) # for reading indv. sheets within excel files
library(kableExtra) # for knitting tables
library(purrr) # for joining multiple sheets successively
library(janitor) # for cleaning variable names
library(scales) # for plotting percentages
library(rstatix) # for calculating significance tables
library(ggbeeswarm) # for jittering in change plots
library(boot) # for bootstrapping CIs
library(forcats) # for wrangling data for forest plots
library(broom) # for visualizing variances
library(patchwork) # for combining plots
library(ggradar) # for radar plots
library(ggpubr) # for adding significances into plots

# Define settings
knitr::opts_chunk$set(message = FALSE, warning = FALSE)

# Define functions
fix_names <- function(x) { # fixing issues with AGRRA col names (e.g. tAPALmean -> t_apa_lmean)
  x %>%
    gsub(" ", "_", .) %>%                        # spaces → underscores
    gsub("([a-z])([A-Z])", "\\1_\\2", .) %>%     # split camelCase
    gsub("^t([A-Z]{4})(avg|med|std)$", "t_\\1_\\2", ., perl = TRUE) %>% # tACANavg → t_ACAN_avg
    gsub("(.*?)(avg|med|std)$", "\\1_\\2", ., perl = TRUE) %>%          # ensure _ before suffix
     gsub("^t([A-Z]{4})(large|small|total)$", "t_\\1_\\2", ., perl = TRUE) %>% 
    gsub("(.*?)(large|small|total)$", "\\1_\\2", ., perl = TRUE) %>%
    tolower()
}

se <- function(x) sd(x, na.rm = TRUE) / sqrt(sum(!is.na(x)))

# Define ggplot themes
theme_pub <- function(base_size = 12, base_family = "") {
  theme_bw(base_size = base_size, base_family = base_family) +
    theme(
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(),
      panel.border = element_rect(color = "black", fill = NA),
      axis.line = element_line(color = "black"),
      axis.title = element_text(size = base_size * 1.2),
      axis.text = element_text(size = base_size),
      legend.title = element_text(size = base_size * 1.1),
      legend.text = element_text(size = base_size * 0.9),
      strip.background = element_rect(fill = "white", color = "black"),
      strip.text = element_text(size = base_size * 1.1, face = "bold")
    )
}

# fig directory
fig_dir <- here("agrra_monitoring", "figs")
```

# Site metadata

```{r}
site_depth <- read_excel(here("agrra_monitoring", "data_raw", "ATG_2017-2022", "Calculated", "Site", "Site.xlsx"), sheet = "Data") %>% 
  clean_names() %>%
  filter(format(date, "%Y") >= 2017) %>%
  select(code, mean_depth)
  
sites_hist <- read_excel(here("agrra_monitoring", "data_raw", "ATG_2017-2022", "Raw", "MSAccess-Metadata.xlsx"), sheet = "Survey") %>% 
  clean_names() %>%
  mutate(name = if_else(str_detect(name, "^Site \\d+"), # making NDNP site names consistent
                        str_extract(name, "^Site \\d+"), 
                        name)) %>%
  select(site = name, t_benth_hist = benthic_transects, t_fish_hist = fish_transects)

sites_meta <- read_excel(here("agrra_monitoring", "data_raw", "ATG_NEMMA_2025", "Raw", "Metadata.xlsx"), sheet = "Survey") %>% 
  mutate(zone = "NEMMA") %>%
  bind_rows(read_excel(here("agrra_monitoring", "data_raw", "ATG_NPA_2025", "Raw", "Metadata.xlsx"), sheet = "Survey") %>%
              mutate(zone = "NDNP")) %>%
  clean_names() %>%
  mutate(name = if_else(str_detect(name, "^Site \\d+"), 
                        str_extract(name, "^Site \\d+"), 
                        name),
         name = if_else(name == "Little Bird Patch", "Little Bird Backreef", name),
         code = str_replace_all(code, "\\s+", "")) %>%
  select(site = name, code, zone, latitude, longitude, t_benth_2025 = benthic_transects, t_fish_2025 = fish_transects) %>%
  left_join(sites_hist, by = "site") %>%
  left_join(site_depth, by = "code") %>%
  distinct() %>%
  filter(!is.na(t_benth_hist)) # removing sites with no historical data

write.csv(sites_meta, here("agrra_monitoring", "data_outputs", "sites_meta.csv"))
```

# Benthic

## Point cover and algal heights

Data overview: 

- Need to work with transect-level data to get accurate standard deviations for dif. TA groupings
- 2017-2022: "BenthicPointCoverScaledByTransect.xlsx"
  - scaled to not include non-reef substrate
  - cover is in percent format (X/100)
  - TAS and TAM included but not STA
  - some transect numbers include surveyor initials
- 2025: "BenthicCoverScaledByTransect.xlsx"
  - scaled to not include non-reef substrate)
  - cover is in proportion format (X/1)
  - TAS/TAM/STA only included within tTA sheet (on "overall" sheet TA includes all forms of turf)
  - Little Bird Patch = Little Bird Backreef
- Here "TAS" will refer to TAS + TAM (+ STA in 2025)

```{r}
# define variables to include (removing CYAN for analyses)
benthic_var <- c("lc", "cca", "ta", "tas", "fma", "cma", "ainv", "peys")

benthic_cover_trans <- 
  # historical data
  read_excel(here("agrra_monitoring", "data_raw", "ATG_2017-2022", "Calculated", "BenthicPointCoverScaled", "BenthicPointCoverScaledByTransect.xlsx"), sheet = "Data") %>% 
  clean_names() %>%
  mutate(year = year(date),
         trans = as.character(trans), # to bind with 2025 NPA data that includes initials
         t_tas = tas + tam) %>% 
  select(site, year, trans, benthic_var) %>%
  # bind 2025 data
  bind_rows(
    bind_rows(
      # 2025 NEMMA
      read_excel(here("agrra_monitoring", "data_raw", "ATG_NEMMA_2025", "Calculated", "BenthicCoverScaledByTransect.xlsx"), sheet = "Overall") %>%
        clean_names() %>%
        mutate(`transect_name` = as.character(`transect_name`)) %>%  # to bind with 2025 NPA data with initials
        # incorporate TA sheet
        left_join(read_excel(here("agrra_monitoring", "data_raw", "ATG_NEMMA_2025", "Calculated", "BenthicCoverScaledByTransect.xlsx"), sheet = "tTA") %>%
                    clean_names() %>%
                    select(transect_id, ta, sta, tas, tam)),
      # 2025 NDNP
      read_excel(here("agrra_monitoring", "data_raw", "ATG_NPA_2025", "Calculated", "BenthicCoverScaledByTransect.xlsx"), sheet = "Overall") %>%
        clean_names() %>%
        # incorporate TA sheet
        left_join(read_excel(here("agrra_monitoring", "data_raw", "ATG_NPA_2025", "Calculated", "BenthicCoverScaledByTransect.xlsx"), sheet = "tTA") %>%
                    clean_names() %>%
                    select(transect_id, ta, sta, tas, tam))) %>%
      mutate(year = 2025,
             survey_name = if_else(survey_name == "Little Bird Patch", "Little Bird Backreef", survey_name),
             tas_sum = tas + sta + tam) %>%
      select(site = survey_name, year, trans = transect_name, lc = t_coral, cca = t_cca, ta, tas = tas_sum, fma = t_fma, cma = t_cma, ainv = t_ainv, peys = t_pey)) %>%
  mutate(site = if_else(str_detect(site, "^Site \\d+"), 
                        str_extract(site, "^Site \\d+"), 
                        site)) %>% # dealing with NPA site name inconsistencies
  filter(year != 2022, # 2022 was incomplete data collection year for NPA
         site %in% sites_meta$site) %>% 
  left_join(sites_meta %>% select(site, zone), by = "site") %>%
  # generating proportions with subset of benthic_var
  rowwise() %>%
  mutate(total = sum(c_across(all_of(benthic_var)), na.rm = TRUE)) %>%
  mutate(across(all_of(benthic_var), ~ (.x / total))) %>% # proportion format
  mutate(total = sum(c_across(all_of(benthic_var)), na.rm = TRUE)) %>% # test for total = 1
  ungroup()

benthic_cover_site <- benthic_cover_trans %>%
  group_by(site, zone, year) %>%
  rename_with(~ str_remove(.x, "t_")) %>%
  summarise(across(c(benthic_var),
                   list(mean = ~ mean(.x, na.rm = TRUE)),
                   .names = "{.col}_{.fn}")) %>%
  ungroup() %>%
  select(site, zone, year, ends_with("_mean")) %>%
  pivot_longer(cols = ends_with("_mean"),
    names_to = "var",
    values_to = "value") %>%
  mutate(metric = "benthic_cover",
         var = str_remove(var, "_mean"),
         unit = "proportion") 
# 
# benthic_cover_year <- benthic_cover_site %>%
#   group_by(year, zone, metric, unit, var) %>%
#   summarise(mean = mean(value, na.rm = TRUE),
#             sd   = sd(value, na.rm = TRUE),
#             .groups = "drop")
# 
# benthic_ttests <- benthic_cover_site %>%
#   group_by(zone, metric, unit, var) %>%
#   wilcox_test(value ~ year,
#               paired = TRUE)
# kable(benthic_ttests)
```

## Algal height
```{r}
algal_height_site <- read_excel(here("agrra_monitoring", "data_raw", "ATG_2017-2022", "Calculated", "AlgaePointCoverScaled", "AlgaePointCoverScaledBySite.xlsx"), sheet = "Data") %>%
  rename_with(fix_names) %>%
  mutate(year = year(date)) %>%
  select(year, site, value = fh_avg) %>%
  bind_rows(read_excel(here("agrra_monitoring", "data_raw", "ATG_NEMMA_2025", "Calculated", "BenthicAlgalHeight.xlsx"), sheet = "Scaled") %>% 
              bind_rows(read_excel(here("agrra_monitoring", "data_raw", "ATG_NPA_2025", "Calculated", "BenthicAlgalHeight.xlsx"), sheet = "Scaled")) %>% 
              rename_with(fix_names) %>%
              mutate(year = 2025,
                     site = if_else(name == "Little Bird Patch", "Little Bird Backreef", name)) %>%
              select(year, site, value = 't_fma-h')) %>%
  mutate(site = if_else(str_detect(site, "^Site \\d+"), 
                        str_extract(site, "^Site \\d+"), 
                        site), # dealing with NPA sites name inconsistencies
         metric = "fma_height",
         var = "height",
         unit = "cm") %>%
  filter(site %in% sites_meta$site & year != 2022) %>%
  left_join(sites_meta %>% select(site, zone))
```




## Algal composition
```{r}
# metadata for genera - family matching
fma_key <- read_excel(here("agrra_monitoring", "data_raw", "ATG_NEMMA_2025", "Calculated", "BenthicCoverScaled.xlsx"), sheet = "Metadata") %>% 
  clean_names() %>%
  fill(scaled_benthic_cover, column_header) %>%
  filter(str_starts(scaled_benthic_cover, "tFMA")) %>%
  select(code = column_header, genera = definition) %>%
  mutate(genera = if_else(code == "BFMA", "Other BFMA",
                          if_else(code == "GFMA", "Other GFMA",
                                  if_else(code == "RFMA", "Other RFMA",
                                          if_else(code == "FMA", "Other FMA", genera)))),
         code = tolower(code)) %>%
  rename(genus = genera)

# currently 2025 only
fma_gen_site <- read_excel(here("agrra_monitoring", "data_raw", "ATG_NEMMA_2025", "Calculated", "BenthicCoverScaled.xlsx"), sheet = "tFMA") %>% 
  mutate(zone = "NEMMA") %>%
  bind_rows(read_excel(here("agrra_monitoring", "data_raw", "ATG_NPA_2025", "Calculated", "BenthicCoverScaled.xlsx"), sheet = "tFMA") %>% 
              mutate(zone = "NDNP")) %>%
  rename_with(fix_names) %>%
  mutate(year = 2025,
         site = if_else(name == "Little Bird Patch", "Little Bird Backreef", name)) %>%
  select(year, zone, site, contains("avg"), fma_tot_avg = t_fma_avg) %>% # otherwise tFMA and (other) FMA columns get duplicated as FMA
  rename_with(~ str_remove(.x, "_avg")) %>%
  pivot_longer(cols = -c(year, zone, site, fma_tot),
               names_to = "code",
               values_to = "percent_cover") %>%
  filter(fma_tot > 0) %>%
  mutate(prop_cover = percent_cover / fma_tot) %>%
  left_join(fma_key, by = "code") %>%
  select(year, zone, site, code, genus, percent_cover, prop_cover) %>%
  filter(!is.na(genus))

fma_gen_year <- fma_gen_site %>%
  group_by(year, zone, genus) %>%
  summarize(mean_perc_cov = mean(percent_cover),
            sd_perc_cov = sd(percent_cover),
            mean_prop_cov = mean(prop_cover),
            sd_prop_cov = sd(prop_cover)) %>%
  filter(mean_perc_cov > 0)

kable(fma_gen_year %>% filter(code == "dict"))
```

## Coral composition

Data overview:

- Due to high UNKN in 2019 NDNP data, family-level assessments will only include NEMMA sites
- Extremely complicated and inconsistent data structures to work with - wrote functions for cleaning headers and created meta file to link codes to species and families consistently across years

```{r}
# metadata for species - family matching
coral_meta <- read_excel(here("agrra_monitoring", "data_raw", "ATG_NEMMA_2025", "Calculated", "BenthicCoralCoverScaled.xlsx"), sheet = "Metadata") %>% 
  clean_names() %>%
  fill(scaled_coral_cover, column_header) 

coral_meta_fam <- coral_meta %>%
  filter(str_starts(column_header, "t") | column_header == "UNKN") %>%
  select(family_code = column_header, family = definition) %>%
  mutate(family = gsub("\\s*\\([^\\)]+\\)", "", family)) %>% # removing fluff
  distinct() # removing UNKN duplicate

coral_meta_spp <- coral_meta %>%
  filter(str_starts(scaled_coral_cover, "t") | scaled_coral_cover == "UNKN") %>%
  select(species_code = column_header, family_code = scaled_coral_cover, species = definition) %>%
  left_join(coral_meta_fam, by = "family_code") %>%
  mutate(genus = sub(" .*", "", species),
         family_code = str_remove(family_code, "^t"),
         across(c(species, family), ~ str_replace(., "Unknown sp. of Live Coral", "Unknown spp.")),
         family = if_else(family == "Scleractinia incertae sedis", "Unknown spp.", family))

# historical data
coral_spp_hist <- read_excel(here("agrra_monitoring", "data_raw", "ATG_2017-2022", "Calculated", "CoralCoverSpeciesScaled", "CoralCoverSpeciesScaledBySite.xlsx"), sheet = "Data") %>% 
  rename_with(fix_names) %>%
  mutate(year = year(date)) %>%
  filter(year %in% c(2017, 2019)) %>% # 2022 was incomplete NPA year
  pivot_longer(
    cols = matches("^[a-z]{4}_(avg|std)$"),   # all columns like acer_ave, apal_std, etc.
    names_to = c("species_code", ".value"),        # split into species and measurement type
    names_sep = "_"
  ) %>%
  mutate(across(c(avg, std), ~ .x / 100)) %>% # to match 2025 format (fraction of 1)
  select(site, year, species_code, avg, std) 

# 2025 data
file_coral_nemma <- here("agrra_monitoring", "data_raw", "ATG_NEMMA_2025", "Calculated", "BenthicCoralCoverScaled.xlsx")
file_coral_npa <- here("agrra_monitoring", "data_raw", "ATG_NPA_2025", "Calculated", "BenthicCoralCoverScaled.xlsx")

clean_excel_coral <- function(file_path) {
  sheets <- excel_sheets(file_path)
  sheets_to_join <- setdiff(sheets, c("TermsOfUse", "Metadata", "Overall"))  # exclude

  sheets_list <- map(sheets_to_join, ~ read_excel(file_path, sheet = .x) %>%
                       rename_with(~ gsub("^(t)([A-Za-z]{4})(avg|med|std)$", "\\1_\\2_\\3", .x, ignore.case = TRUE)) %>%
                       rename_with(~ gsub("([A-Za-z]{4})(avg|med|std)$", "\\1_\\2", .x, ignore.case = TRUE)) %>%
                       clean_names())

  reduce(sheets_list, left_join, by = c("survey_id", "code", "name", "nt"))
}

coral_spp_2025 <- clean_excel_coral(file_coral_npa) %>%
  bind_rows(clean_excel_coral(file_coral_nemma)) %>%
  mutate(year = 2025) %>%
  rename(site = name) %>%
  pivot_longer(
    cols = matches("^[a-z]{4}_(avg|std)$"),   # all columns like acer_ave, apal_std, etc.
    names_to = c("species_code", ".value"),        # split into species and measurement type
    names_sep = "_"
  ) %>%
  select(site, year, species_code, avg, std)

# merge for complete df
coral_spp_site <- bind_rows(coral_spp_hist, coral_spp_2025) %>%
  mutate(zone = if_else(str_detect(site, "^Site \\d+"), "NDNP",
                            if_else(str_detect(site, "Redonda"), "Redonda",
                                    "NEMMA")),
         site = if_else(str_detect(site, "^Site \\d+"), 
                        str_extract(site, "^Site \\d+"), 
                        site)) %>%
  mutate(species_code = toupper(species_code),
         site = if_else(site == "Little Bird Patch", "Little Bird Backreef", site)) %>%
  left_join(coral_meta_spp, by = "species_code")  %>%
  group_by(site) %>%
  filter(n_distinct(year) > 1) %>% # only keeping sites with over time comparisons
  ungroup()

coral_fam_site <- coral_spp_site %>%
  group_by(year, zone, site, family) %>%
  summarize(value = sum(avg, na.rm = TRUE)) %>%
  ungroup()

coral_fam_year <- coral_fam_site %>%
  group_by(year, zone, site) %>%
  complete( # adding 0s for when nothing was present
    family = unique(coral_spp_site$family),
    fill = list(avg = 0)
  ) %>%
  ungroup() %>%
  group_by(year, zone, family) %>%
  summarize(mean = mean(value, na.rm = TRUE), 
         sd = sd(value, na.rm = TRUE)) %>%
  ungroup() %>%
  group_by(year, zone) %>%
  mutate(rel_mean = mean / sum(mean, na.rm = TRUE)) %>%
  ungroup()

chk <- coral_fam_year %>%
  group_by(year, zone) %>%
  summarize(tot = sum(rel_mean, na.rm = TRUE))
kable(coral_fam_year)

coral_fam_ttests <- coral_fam_site %>%
  group_by(zone, family) %>%
  wilcox_test(value ~ year,
              paired = TRUE)
kable(coral_fam_ttests)

# select primary families for plots
coral_fams <- c("Acroporidae", "Faviidae", "Merulinidae", "Milleporidae", "Montastraeidae",  "Poritidae", "Siderastreidae") #"Unknown spp."
```

### Delta plots
```{r}
coral_delta_table <- coral_fam_site %>%
  filter(zone == "NEMMA", year %in% c(2017, 2025), family %in% coral_fams) %>%
  select(year, site, family, value) %>%
  # Pivot so each row is a site × family, with 2017 and 2025 in separate columns
  tidyr::pivot_wider(
    names_from = year,
    values_from = value,
    names_prefix = "yr_"
  ) %>%
  # Keep only sites that have data for both years
  filter(!is.na(yr_2017) & !is.na(yr_2025)) %>%
  # Calculate delta per site
  mutate(delta = yr_2025 - yr_2017) %>%
  group_by(family) %>%
  summarise(
    mean_delta = mean(delta, na.rm = TRUE),
    ci = list(if(length(unique(delta)) > 1) boot_mean_ci(delta) else c(NA, NA)),
    wilcox_p = if(length(unique(delta)) > 1) wilcox.test(delta, mu = 0)$p.value else NA,
    .groups = "drop"
  ) %>%
  mutate(
    ci_lower = sapply(ci, `[[`, 1),
    ci_upper = sapply(ci, `[[`, 2)
  ) %>%
  select(-ci) %>%
  mutate(
    mean_delta = round(mean_delta, 3),
    ci_lower = round(ci_lower, 3),
    ci_upper = round(ci_upper, 3),
    wilcox_p = signif(wilcox_p, 3)
  ) %>%
  arrange(family)

p_delta_coral <- ggplot(coral_delta_table,
                        aes(x = mean_delta, y = forcats::fct_rev(family))) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "grey50") +
  geom_errorbarh(aes(xmin = ci_lower, xmax = ci_upper), height = 0.2) +
  geom_point(size = 3) +
  labs(
    x = "Mean change (% cover)",
    y = NULL,
    title = NULL
  ) +
  # scale_x_continuous(labels = label_percent(accuracy = 0.1)) +
  scale_x_continuous(labels = function(x) sprintf("%.1f", x * 100)) +
  theme_pub() +
  theme(panel.grid.major.x = element_line(color = "grey95"))

p_delta_coral
```


### Heat plots
```{r}
ggplot(coral_fam_site %>%
         filter(zone == "NEMMA" & family %in% c(coral_fams)) %>%
         left_join(sites_meta %>% select(site, code)),
       aes(code, family, fill = value)) +
  geom_tile() +
  facet_wrap(~year) +
  scale_fill_viridis_c(labels = percent,
                       name = "Percent cover") +
  labs(x = NULL,
       y = NULL,
       fill = "Percent cover") +
  theme_minimal(base_size = 12) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
  
ggsave(here(fig_dir, "coral_fam_heat.png"), width = 8, height = 5)
```

## Coral recruits

Data overview:

- Historical: CoralRecruitsBySite.xlsx
  - family level
  - remove "DICH" code (not in 2025 classifications, not present in either dataset)
  - missing many families present in 2025 (e.g., ACRO, MERU)
- 2025: BenthicCoralRecruis.xlxs
  - genus and species level
  - missing site-level standard deviation in data

```{r}
# using coral_meta_spp key generated in prev. section

# historical data

recruits_hist_sized <- read_excel(here("agrra_monitoring", "data_raw", "ATG_2017-2022", "Calculated", "CoralRecruits", "CoralRecruitsBySite.xlsx"), sheet = "Small") %>% 
  mutate (size = "small") %>%
  rbind(read_excel(here("agrra_monitoring", "data_raw", "ATG_2017-2022", "Calculated", "CoralRecruits", "CoralRecruitsBySite.xlsx"), sheet = "Large") %>% 
          mutate (size = "large")) %>%
  rename_with(fix_names) %>%
  mutate(year = year(date)) %>%
  filter(year %in% c(2017, 2019)) %>% # 2022 was incomplete NPA year
  pivot_longer(
    cols = matches(".*_(avg|std)$"),   # all columns ending in avg or std
    names_to = c("family_code", ".value"),        # split into species and measurement type
    names_sep = "_"
  ) %>%
  select(site, year, family_code, size, avg) 

recruits_hist_tot <- recruits_hist_sized %>%
  group_by(site, year, family_code) %>%
  summarize(
    avg = sum(avg, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(size = "total") %>%
  select(site, year, family_code, size, avg) 
recruits_hist <- bind_rows(recruits_hist_tot, recruits_hist_sized)

# 2025 data

recruits_2025 <- read_excel(here("agrra_monitoring", "data_raw", "ATG_NEMMA_2025", "Calculated", "BenthicCoralRecruits.xlsx"), sheet = "Overall") %>%
  bind_rows(read_excel(here("agrra_monitoring", "data_raw", "ATG_NPA_2025", "Calculated", "BenthicCoralRecruits.xlsx"), sheet = "Overall")) %>%
  rename_with(fix_names) %>%
  mutate(year = 2025) %>%
  rename(site = name) %>%
  select(year, site, matches("_(small|large|total)$")) %>%
  pivot_longer(
    cols = matches("_(small|large|total)$"),
    names_to = c("family_code", "size"),
    names_pattern = "^(.*?)_(small|large|total)$",
    values_to = "avg"
  ) %>%
  select(site, year, family_code, size, avg) %>%
  mutate(family_code = str_remove(family_code, "^t_"))
# no SD for 2025

test <- recruits_2025 %>%
  distinct(family_code, size) %>%
  anti_join(
    recruits_hist %>%
      distinct(family_code, size),
    by = c("family_code", "size")
  )

# merge for complete df
recruits_fam_site <- bind_rows(recruits_hist, recruits_2025) %>%
  mutate(zone = if_else(str_detect(site, "^Site \\d+"), "NDNP",
                            if_else(str_detect(site, "Redonda"), "Redonda",
                                    "NEMMA")),
         site = if_else(str_detect(site, "^Site \\d+"), 
                        str_extract(site, "^Site \\d+"), 
                        site),
         site = if_else(site == "Little Bird Patch", "Little Bird Backreef", site),
         family_code = toupper(family_code)) %>%
  filter(family_code != "DICH") %>% # in historical data only, none present  
  # only keeping sites with over time comparisons
  group_by(site) %>%
  filter(n_distinct(year) > 1) %>% 
  ungroup() %>%
  # adding 0s for missing data
  complete(nesting(site, year, zone), 
           family_code,
           size,
           fill = list(avg = 0)) %>%
  distinct() %>%
  # joining family names
  left_join(coral_meta_spp %>%
              select(family_code, family) %>%
              distinct(), 
            by = "family_code") %>%
  mutate(family = if_else(family_code == "ALL", "All species", 
                          if_else(family_code == "CARB", "Unknown spp.", family)))

recruits_fam_year <- recruits_fam_site %>%
  group_by(year, zone, family_code, family, size) %>%
  summarize(mean = mean(avg), # 0s already included for absent families
            sd = sd(avg)) %>%
  group_by(year, zone) %>%
  mutate(rel_mean = mean / sum(mean))
# kable(recruits_fam_year)
# 
# recruits_fam_ttests <- recruits_fam_site %>%
#   group_by(zone, family, family_code, size) %>%
#   wilcox_test(avg ~ year,
#               paired = TRUE)
# kable(recruits_fam_ttests)
```

## Diadema

Data overview:
- 2025: MotileInverts.xlsx
- 2017-2019: MotileInvertsBySite.xlsx

```{r}
diadema <- read_excel(here("agrra_monitoring", "data_raw", "ATG_2017-2022", "Calculated", "MotileInverts", "MotileInvertsBySite.xlsx"), sheet = "Data") %>% 
  rename_with(fix_names) %>%
  mutate(year = year(date)) %>%
  select(site, year, matches("^dt_|^da_|^dj_")) %>%
  rbind(read_excel(here("agrra_monitoring", "data_raw", "ATG_NEMMA_2025", "Calculated", "MotileInverts.xlsx"), sheet = "Overall") %>% 
          rbind(read_excel(here("agrra_monitoring", "data_raw", "ATG_NPA_2025", "Calculated", "MotileInverts.xlsx"), sheet = "Overall")) %>%
          rename_with(fix_names) %>%
          mutate(year = 2025,
                 name = if_else(name == "Little Bird Patch", "Little Bird Backreef", name)) %>%
          select(site = name, year, matches("^dt_|^da_|^dj_"), -ends_with("_med"))) %>%
  filter(year != 2022) %>% # incomplete year for NPA
  mutate(site = if_else(str_detect(site, "^Site \\d+"), 
                        str_extract(site, "^Site \\d+"), 
                        site)) %>%
  filter(site %in% sites_meta$site) %>%
  left_join(sites_meta %>% select(site, zone)) %>%
  group_by(site) %>%
  filter(n_distinct(year) > 1) %>% # only keeping sites with over time comparisons
  ungroup()

diadema_site <- diadema %>%
  select(zone, site, year, dt_avg, da_avg, dj_avg) %>%  # keep density columns
  pivot_longer(
    cols = c(dt_avg, da_avg, dj_avg),
    names_to = "var",
    values_to = "value"
  ) %>%
  mutate(metric = "diadema_density",
         unit = "g_m2",
         var = recode(var,
                        dt_avg = "total",
                        da_avg = "adult",
                        dj_avg = "juvenile"))

diadema_year <- diadema_site %>%
  group_by(zone, year, metric, unit, var) %>%
  summarise(mean = mean(value, na.rm = TRUE),
            sd   = sd(value, na.rm = TRUE),
            .groups = "drop")

diadema_tests <- diadema_site %>%
  mutate(period = if_else(year == 2025, "y2025", "hist")) %>%
  select(site, zone, var, period, value) %>%
  pivot_wider(names_from = period,
              values_from = value) %>%
  group_by(zone, var) %>%
  summarise(p = wilcox.test(hist, y2025, paired = TRUE, exact = FALSE)$p.value,
    .groups = "drop")
```

# Fish

## Biomass

Data overview:

- 2017: FishBiomassByTransect.xlsx
   - "Data" sheet has transect-level data by family and group
   - Ignoring "group" classifications here as it isn't consistent with 2025
- 2025: FishBiomassByTransect.xlsx
   - "Overall" sheet has transect-level data by family
   - Family sheets have transect-level data by species
- Converting all data to kg/100m2

```{r}
# create compatible family key by merging metadata
fish_fam_key <- full_join(
  read_excel(here("agrra_monitoring", "data_raw", "ATG_2017-2022", "Calculated", "FishBiomass", "FishBiomassBySite.xlsx"), sheet = "Key") %>% 
    clean_names() %>%
    slice(21:40) %>%
    select(code_com = 1,
           definition = 2) %>%
    mutate(family = str_remove(word(definition, 1), ":"),
           family = if_else(family == "Wrasse", "Wrasses", 
                            if_else(code_com == "PUFF", "Pufferfishes", family))) %>%
    select(-definition), 
  read_excel(here("agrra_monitoring", "data_raw", "ATG_NEMMA_2025", "Calculated", "FishBiomass.xlsx"), sheet = "Metadata") %>% 
    clean_names() %>%
    slice(23:42) %>%
    select(code_lat = 2,
           definition = 3) %>%
    mutate(family = str_remove(word(definition, 1), ":")) %>%
    select(-definition), 
  by = "family") %>%
  mutate(code_lat = tolower(gsub("^t", "", code_lat)),
         code_com = tolower(code_com)) %>%
  select(code_com, code_lat, family) %>%
  bind_rows(
    tibble(code_com = "t",
           code_lat = "all",
           family   = "Total")) %>%
  pivot_longer(cols = c(code_com, code_lat),
               names_to = "code_type",
               values_to = "code")

fish_bm_fam_site <- bind_rows(
  # historical data
  read_excel(here("agrra_monitoring", "data_raw", "ATG_2017-2022",
                  "Calculated", "FishBiomass", "FishBiomassBySite.xlsx"),
             sheet = "Data") %>%
    rename_with(fix_names) %>%
    mutate(code_type = "code_com",
             year = year(date)),
  # 2025 data
  rbind(read_excel(here("agrra_monitoring", "data_raw", "ATG_NEMMA_2025",
                    "Calculated", "FishBiomass.xlsx"),
               sheet = "Overall"),
        read_excel(here("agrra_monitoring", "data_raw", "ATG_NPA_2025",
                    "Calculated", "FishBiomass.xlsx"),
               sheet = "Overall")) %>%
    rename_with(fix_names) %>%
    mutate(year = 2025,
           site = if_else(name == "Little Bird Patch",
                          "Little Bird Backreef",
                          name),
           code_type = "code_lat") %>%
    rename_with(~ gsub("^t_", "", .x), .cols = contains("avg"))
) %>%
  # wrangle data
  ## min n transect = 8 in both years
  group_by(site) %>%
  filter(all(nt >= 8)) %>%
  ungroup() %>%
  mutate(site = if_else(str_detect(site, "^Site \\d+"), str_extract(site, "^Site \\d+"), site),
         if_else(name == "Little Bird Patch", "Little Bird Backreef", name)) %>%
  select(site, year, code_type, contains("avg")) %>%
  pivot_longer(cols = contains("avg"),
               names_to = "family_stat",
               values_to = "value") %>%
  separate(family_stat,
           into = c("code", "stat"),
           sep = "_(?=[^_]+$)") %>%
  # merge with family key
  left_join(fish_fam_key, by = c("code_type", "code")) %>%
  left_join(sites_meta %>% select(site, zone)) %>%
  mutate(value = value/1000, # converting from g to kg
         metric = "fish_biomass",
         unit = "kg_100m2") %>%
  filter(site %in% sites_meta$site,
         year != 2022, # incomplete NDNP year
         !is.na(family)) %>% # historical groupings excluded from key
  select(site, zone, year, metric, var = family, value, unit)

herb_bm_site <- fish_bm_fam_site %>%
  filter(var %in% c("Parrotfishes", "Surgeonfishes")) %>%
  group_by(site, zone, year, metric, unit) %>%
  summarise(value = sum(value, na.rm = TRUE),
    .groups = "drop") %>%
  mutate(var = "Herbivores") %>%
  select(site, zone, year, metric, var, value, unit)

fish_bm_site <- rbind(fish_bm_fam_site, herb_bm_site) %>%
  filter(var %in% c("Total", "Herbivores"))

fish_bm_year <- fish_bm_site %>%
  group_by(year, zone, metric, var, unit) %>%
  summarize(mean = mean(value),
            sd = sd(value))
```



# Summary dfs
```{r}
sum_site_yr <- bind_rows(benthic_cover_site, algal_height_site, diadema_site, fish_bm_site) %>%
  mutate(var = tolower(var))

# create explicit paired site-level dataset
sum_site_pairs <- sum_site_yr %>%
  mutate(period = if_else(year == 2025, "2025", "hist")) %>%
  select(site, zone, metric, var, unit, period, value) %>%
  pivot_wider(
    names_from = period,
    values_from = value
  ) %>%
  filter(!is.na(hist) & !is.na(`2025`)) %>%
  mutate(site_delta = `2025` - hist)

options(scipen = 999) # avoid scientific notation in printing values

# create zone-level summary
sum_zone <- sum_site_pairs %>%
  group_by(zone, metric, var, unit) %>%
  summarise(

    mean_hist = mean(hist, na.rm = TRUE),
    sd_hist   = sd(hist, na.rm = TRUE),

    mean_2025 = mean(`2025`, na.rm = TRUE),
    sd_2025   = sd(`2025`, na.rm = TRUE),

    pct_change = ((mean(`2025`, na.rm = TRUE) -
                     mean(hist, na.rm = TRUE)) /
                     mean(hist, na.rm = TRUE)) * 100,

    delta = mean(site_delta, na.rm = TRUE),

    CI = list(
      quantile(
        replicate(
          5000,
          mean(sample(site_delta,
                      replace = TRUE),
               na.rm = TRUE)
        ),
        probs = c(0.025, 0.975),
        na.rm = TRUE
      )
    ),

    p = suppressWarnings(
      wilcox.test(
        hist,
        `2025`,
        paired = TRUE,
        exact = FALSE
      )$p.value
    ),

    .groups = "drop"
  ) %>%
  
  mutate(
    CI_low  = map_dbl(CI, 1),
    CI_high = map_dbl(CI, 2)
  ) %>%
  
  select(-CI) %>%
  
  rstatix::add_significance(p.col = "p") %>%
  rename(p_label = p.signif)

kable(sum_zone)
```

# Plots

### Benthic bar
```{r}
benthic_barplot <- benthic_cover_year %>%
  filter(var %in% c(benthic_var)) %>%
  # create "other" category for plot
  mutate(var = if_else(var %in% c("cma", "peys", "ainv"),
                          "other",
                          var)) %>%
  # recalculate means with new "other" category
  group_by(year, zone, var) %>%
  summarise(mean = sum(mean),
            .groups = "drop") %>%
  # labels
  mutate(
    var = recode(
      var,
      lc    = "Live coral",
      cca   = "CCA",
      ta    = "Turf algae",
      tas   = "Turf + sediment",
      fma   = "Fleshy macroalgae",
      other = "Other"
    ),
    var = factor(
      var,
      levels = c(
        "Other",
        "Live coral",
        "CCA",
        "Turf algae",
        "Turf + sediment",
        "Fleshy macroalgae")),
    year = factor(year, levels = c(2025, 2019, 2017)))

ggplot(benthic_barplot,
       aes(x = mean,
           y = factor(year),
           fill = var)) +
  geom_col(width = 0.7) +
  facet_wrap(~zone,
             scales = "free_y") +
  scale_fill_manual(values = c(
    "Live coral" = "#fc8d59",
    "CCA" = "pink",
    "Turf algae" = "#d9ef8b",
    "Turf + sediment" = "tan",
    "Fleshy macroalgae" = "#1b9e77",
    "Other" = "gray80")) +
  scale_x_continuous(labels = scales::percent_format(accuracy = 1)) +
  labs(x = "Percent cover",
       y = NULL,
       fill = NULL) +
  guides(fill = guide_legend(
    nrow = 1,
    reverse = TRUE,
    byrow = TRUE)) +
  theme_minimal(base_size = 12) +
  theme(panel.grid.major.y = element_blank(),
        panel.grid.minor = element_blank(),
        strip.text = element_text(face = "bold"),
        axis.title.x = element_text(
          margin = margin(t = 12)),
        legend.position = "bottom",
        legend.text = element_text(size = 10),
        legend.key.width = unit(0.8, "cm"),
        legend.box = "horizontal")

# ggsave(here(fig_dir, "benthic_sum.png"), width = 8, height = 5)
```

### Benthic delta

```{r}
p_delta_pc <- ggplot(sum_zone %>%
                       filter(metric == "benthic_cover",
                              var %in% benthic_var) %>%
                       mutate(delta_pct = delta * 100,
                              CI_low_pct = CI_low * 100,
                              CI_high_pct = CI_high * 100,
                              var = recode(var,
                                           lc   = "Live coral",
                                           cca  = "CCA",
                                           ta   = "Turf algae",
                                           tas  = "Turf + sediment",
                                           fma  = "Fleshy macroalgae",
                                           cma  = "Calc. macroalgae",
                                           ainv = "Aggressive inverts",
                                           peys = "Peysonnelids")),
                     aes(x = delta_pct, y = var)) +
  geom_vline(xintercept = 0,
             linetype = "dashed",
             color = "grey50") +
  geom_errorbarh(aes(xmin = CI_low_pct,
                     xmax = CI_high_pct),
                 height = 0.2) +
  geom_point(size = 3) +
  facet_wrap(~ zone) +
  labs(x = "Mean change (percent cover)",
       y = NULL) +
  theme_bw() +
  theme(strip.background = element_blank(),
        strip.text = element_text(face = "bold"),
        panel.grid.major.y = element_blank())


p_delta_fmah <- ggplot(sum_zone %>%
                         filter(metric == "fma_height") %>%
                         mutate(var = "Fleshy macroalgal height"),
                       aes(x = delta, y = var)) +
  geom_vline(xintercept = 0,
             linetype = "dashed",
             color = "grey50") +
  geom_errorbarh(aes(xmin = CI_low,
                     xmax = CI_high),
                 height = 0.2) +
  geom_point(size = 3) +
  facet_wrap(~ zone) +
  labs(x = "Mean change (cm)",
       y = NULL) +
  theme_bw() +
  theme(strip.background = element_blank(),
        strip.text = element_text(face = "bold"),
        panel.grid.major.y = element_blank())

p_delta_pc / p_delta_fmah + plot_layout(heights = c(3, 1))
ggsave(here(fig_dir, "benthic_delta_panel.png"), width = 6, height = 4)
```

### Fish biomass line
```{r}
make_lp_summary <- function(summary_df,
                            sig_df,
                            metric_name,
                            family_name,
                            ylab,
                            hline_value = NULL) {

  plot_df <- summary_df %>%
    filter(metric == metric_name,
           var == family_name) %>%
    group_by(zone, year) %>%
    summarise(
      mean = mean(value, na.rm = TRUE),
      sd   = sd(value, na.rm = TRUE),
      .groups = "drop"
    ) %>%
    mutate(year = as.numeric(year))

  sig_df_var <- sig_df %>%
    filter(metric == metric_name,
           var == family_name)

  plot_df <- plot_df %>%
    left_join(sig_df_var, by = "zone")

  p <- ggplot(plot_df,
              aes(x = year,
                  y = mean,
                  group = zone,
                  color = zone)) +
    geom_point(size = 3) +
    geom_line(size = 1) +
    scale_color_manual(values = c("black", "cadetblue")) +
    scale_x_continuous(breaks = seq(2017, 2025, by = 2),  # odd years from 2017 to 2025
                       labels = seq(2017, 2025, by = 2)) +
    geom_errorbar(aes(ymin = mean - sd,
          ymax = mean + sd),
      width = 0.2) +
    geom_text(data = plot_df %>% filter(year == max(year)),
      aes(label = p_label),
      hjust = -0.3,
      size = 5,
      show.legend = FALSE) +
    labs(x = "",
      y = ylab,
      color = NULL) +
    theme_pub() +
    theme(legend.position = "bottom")

  if (!is.null(hline_value)) {
    p <- p +
      geom_hline(
        yintercept = hline_value,
        linetype = "dashed",
        color = "black"
      )
  }

  return(p)
}

p_bm_tot <- make_lp_summary(
  summary_df = sum_site_yr,
  sig_df = sum_zone,
  metric_name = "fish_biomass",
  family_name = "total",
  ylab = expression("Total fish biomass (" * kg ~ 100 ~ m^{-2} * ")"),
  hline_value = 1306 / 100 # Karr et al 2015 - 1300 kg/ha - converting to g/100m2
)

p_bm_herb <- make_lp_summary(
  summary_df = sum_site_yr,
  sig_df = sum_zone,
  metric_name = "fish_biomass",
  family_name = "herbivores",
  ylab = expression("Herbivore biomass (" * kg ~ 100 ~ m^{-2} * ")"),
  hline_value = 29.2*100 / 1000 # Edwards et al 2014 - 29.2 g/m2 - converting to g/100m2
)

p_bm_tot / p_bm_herb + plot_layout(heights = c(10,10))
ggsave(here(fig_dir, "fish_bm_change.png"), width = 6, height = 8)
```
